---
title: "Methods and data analysis for activity assays in cell cultures"
author: "Leonard Blaschek"
date: '2022-09-02'
output: 
  pdf_document:
    latex_engine: lualatex
    fig_caption: yes
    fig_height: 6
    includes:
      in_header: rmd_temp.tex
sansfont: IBM Plex Sans
monofont: Inconsolata
keep_tex: yes
citation_package: biblatex
bibliography: /home/leonard/Documents/Bibliography/Zotero_lib.bib
csl: elife.csl
link-citations: true
linkcolor: KU
urlcolor: KU
citecolor: KU
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(showtext)

#### import IBMPlexSans ####
font_add(
  "IBMPlexSans",
  regular = "/OFL_fonts/IBMPlexSans-Light.otf",
  italic = "/OFL_fonts/IBMPlexSans-LightItalic.otf",
  bold = "/OFL_fonts/IBMPlexSans-SemiBold.otf",
  bolditalic = "/OFL_fonts/IBMPlexSans-SemiBoldItalic.otf"
)
showtext_auto()

#### generating plot theme ####
theme_leo <- function(base_size = 11,
                      base_family = "IBMPlexSans") {
  theme_minimal(
    base_size = base_size,
    base_family = base_family
  ) %+replace%
    theme(
      strip.text = element_text(hjust = 0, face = "italic"),
      axis.ticks = element_line(
        size = 0.25,
        lineend = "square",
        color = "black"
      ),
      axis.text.x = element_text(
        colour = "black", # flipped coords
        margin = margin(1, 1, 1, 1)
      ),
      axis.text.y = element_text(
        colour = "black",
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1)
      ),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      # panel.border = element_rect(fill = NA, color = "black", size = 0.25),
      panel.spacing = unit(1.5, "mm"),
      legend.position = "bottom",
      legend.text = element_text(size = rel(0.8)),
      legend.key.height = unit(4, "mm"),
      complete = TRUE
    )
}

ggtext_size <- 8 / (14 / 5)
```

# Protocol

## Stock solutions

```{r, echo=FALSE, message=FALSE}
options(knitr.kable.NA = '')
stock_table <- read_csv("/home/leonard/Dropbox/2021_lac_manuscript/Tables/substrates.csv") %>% 
  select(-Supplier, - `Working concentration sections`)

kable(stock_table,
  "latex",
  caption = "Stock solutions, working concentrations and absorbance extrema of LAC substrates.",
  col.names = c(
    "Compound",
    "Abbrev.",
    "Product No.",
    "Stock",
    "Solvent",
    "Working",
    "$\\lambda$\\textsubscript{max}",
    "$\\lambda$\\textsubscript{min}"
  ),
  escape = F,
  booktabs = T,
  linesep = ""
) %>%
  kable_styling(latex_options = c(
    "striped",
    "HOLD_position",
    "scale_down"
  )) %>% 
  footnote(
    general = "All compounds purchased from Sigma-Aldrich",
    # alphabet = c("\\\\cite{Espin1997}", "\\\\cite{Richardson2000}"),
    general_title = "",
    threeparttable = T,
    escape = F
  )
```

## Harvesting and washing of cells

### Harvesting

1. Harvest 10 ml of culture per day; this is enough for roughly 30 wells (*e.g*. 8 substrates for pH 4--7). At later stages (>5 d) of the non-induced cultures, harvesting 5 ml or even 3 ml is enough due to the density of the culture.
2. Spin down the cells (2 min @ 3000 *g*, or whatever pellets the cells properly), discard the supernatant
3. Resuspend the cells in 70% ethanol to a total volume of ~2 ml and transfer to a 2 ml eppendorf
4. Spin down the eppendorf, discard the supernatant and fill up with 70% ethanol; vortex well
5. Store at -20°C

### Washing prior to the activity assay

1. Take the required amount of cells from the freezer (I calculate 60 ml per well, plus margins)
2. Wash cells twice in 70% ethanol and twice in MilliQ
3. Store in the fridge (I washed cells for one week of experiments in advance, longer storage reduces activity)

## Starting the assay

### Plate layout

My usually used plate layout looked as you can see below, testing one replicate for one substrate across the whole differentiation timeline at pH 4--7. This can of course be changed freely.

```{r, echo = FALSE, include = TRUE, warning=FALSE, fig.height=4, fig.cap="Plate layout of a typical LAC acitivty assay."}
plate <- tibble(
  row = rep(LETTERS[1:8], length.out = 96, each = 12),
  column = rep(1:12, length.out = 96, each = 1),
  sample = rep(c("nTE", "TE"), each = 12, length.out = 96),
  pH = rep(c(4:7), each = 24, length.out = 96),
  day = rep(c(0:9, 14, 21), length.out = 96),
  substrate = ordered("ABTS",
    levels = c("ABTS", "DAF", "MBTH", "PYGL", "SGZ")
  ),
  concentration = 5,
  replicate = 5
) %>%
  mutate(sample = case_when(
    row %in% c("B", "D", "F", "H") & column == 1 ~ "neg. ctrl.",
    TRUE ~ sample
  ))

ggplot(plate) +
  geom_tile(aes(x = column, y = reorder(row, desc(row)), fill = substrate, alpha = pH),
    colour = rgb(1, 1, 1, 1),
    # fill = "#66CCCC",
    size = 2
  ) +
  geom_text(
    data = plate %>% filter(sample != "neg. ctrl."),
    aes(
      x = column, y = reorder(row, desc(row)),
      label = paste(sample, "\nDPI:", day)
    ),
    size = 2, family = "IBMPlexSans"
  ) +
  geom_text(
    data = plate %>% filter(sample == "neg. ctrl."),
    aes(
      x = column, y = reorder(row, desc(row)),
      label = "NEGATIVE\nCONTROL"
    ),
    size = 2, family = "IBMPlexSans"
  ) +
  # labs(title = paste("Replicate", unique(plate$replicate))) +
  scale_x_continuous(breaks = c(1:12), position = "top", expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_fill_manual(
    values = c(
      "#66CCCC",
      "#6699FF",
      "#D35877",
      "#FFCC66",
      "#FF99CC"
    ),
    drop = FALSE
  ) +
  theme_leo() +
  theme(
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = "right"
  ) +
  coord_fixed()
```

### Reaction mix

The final reaction mix contains:

* 100 µl 0.2M buffer (half strength, *i.e.* 100 µl 0.1M buffer, for SGZ, otherwise it precipitates)
* 60 µl washed cells (H~2~O for the negative control)
* 40 µl of appropriately diluted substrate (see table) --- added after the first two measurements

### Hidex program

The used program of the Hidex plate reader should be named something like "TE_activity". It measures the whole spectrum every ten minutes and includes a pause after 10 min (two measurements), to add the substrate. The average of the first two measurements is used to estimate cell density (detailed below).

# References
