---
title: "Methods and data analysis for activity assays in cell cultures"
author: "Leonard Blaschek"
date: '2022-09-02'
output: 
  pdf_document:
    latex_engine: lualatex
    fig_caption: yes
    fig_height: 6
    includes:
      in_header: rmd_temp.tex
sansfont: IBM Plex Sans
monofont: Inconsolata
monofontoptions: 'Scale=0.7'
keep_tex: yes
citation_package: biblatex
bibliography: /home/leonard/Documents/Bibliography/Zotero_lib.bib
csl: elife.csl
link-citations: true
linkcolor: KU
urlcolor: KU
citecolor: KU
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(broom)
library(ggtext)
library(kableExtra)
library(showtext)

#### import IBMPlexSans ####
font_add(
  "IBMPlexSans",
  regular = "/OFL_fonts/IBMPlexSans-Light.otf",
  italic = "/OFL_fonts/IBMPlexSans-LightItalic.otf",
  bold = "/OFL_fonts/IBMPlexSans-SemiBold.otf",
  bolditalic = "/OFL_fonts/IBMPlexSans-SemiBoldItalic.otf"
)
showtext_auto()

#### generating plot theme ####
theme_leo <- function(base_size = 6,
                      base_family = "IBMPlexSans") {
  theme_minimal(
    base_size = base_size,
    base_family = base_family
  ) %+replace%
    theme(
      axis.title.x = element_text(size = 6),
      axis.title.y = element_text(angle = 90, size = 6),
      strip.text = element_text(size = 6, hjust = 0, face = "italic"),
      axis.text.x = element_text(
        size = 6,
        colour = "black", # flipped coords
        margin = margin(1, 1, 1, 1)
      ),
      axis.text.y = element_text(
        size = 6,
        colour = "black",
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1)
      ),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      # panel.border = element_rect(fill = NA, color = "black", size = 0.25),
      panel.spacing = unit(1.5, "mm"),
      legend.position = "bottom",
      legend.text = element_text(size = rel(0.8)),
      legend.key.height = unit(4, "mm"),
      complete = TRUE
    )
}

ggtext_size <- 8 / (14 / 5)

pal_dawn <- c(
  "#434c52ff",
  "#46636fff",
  "#6088a0ff",
  "#ab6c82ff",
  "#d8737fff",
  "#eaa05fff",
  "#ffd175ff"
)

#### machine dependent paths ####
datapath <- ifelse(dir.exists("/data/"), "/data/",
  ifelse(dir.exists("/run/media/leonard/data/grsync/data/"), "/run/media/leonard/data/grsync/data/",
    ifelse(dir.exists("/run/media/leonard/data/"), "/run/media/leonard/data/", "/home/leonard/Documents/Uni/")
  )
)
```

# Experimental procedure

## Stock solutions

```{r, echo=FALSE, message=FALSE}
options(knitr.kable.NA = "\\textcolor{Gray}{\\textit{not tested}}")
stock_table <- read_csv("/home/leonard/Dropbox/2021_lac_manuscript/Tables/substrates.csv") %>%
  select(-Supplier, -`Working concentration sections`)

kable(stock_table,
  "latex",
  caption = "Stock solutions, working concentrations and absorbance extrema of LAC substrates.",
  col.names = c(
    "Compound",
    "Abbrev.",
    "Product No.",
    "Stock",
    "Solvent",
    "Working",
    "$\\lambda$\\textsubscript{max}",
    "$\\lambda$\\textsubscript{min}"
  ),
  escape = F,
  booktabs = T,
  linesep = ""
) %>%
  kable_styling(latex_options = c(
    "striped",
    "HOLD_position",
    "scale_down"
  )) %>%
  footnote(
    general = "All compounds purchased from Sigma-Aldrich",
    # alphabet = c("\\\\cite{Espin1997}", "\\\\cite{Richardson2000}"),
    general_title = "",
    threeparttable = T,
    escape = F
  )
```

## Harvesting and washing of cells

### Harvesting

1. Harvest 10 ml of culture per day; this is enough for roughly 30 wells (*e.g*. 8 substrates for pH 4--7). At later stages (>5 d) of the non-induced cultures, harvesting 5 ml or even 3 ml is enough due to the density of the culture.
2. Spin down the cells (2 min @ 3000 *g*, or whatever pellets the cells properly), discard the supernatant
3. Resuspend the cells in 70% ethanol to a total volume of ~2 ml and transfer to a 2 ml eppendorf
4. Spin down the eppendorf, discard the supernatant and fill up with 70% ethanol; vortex well
5. Store at -20°C

### Washing prior to the activity assay

1. Take the required amount of cells from the freezer (I calculate 60 ml per well, plus margins)
2. Wash cells twice in 70% ethanol and twice in MilliQ
3. Store in the fridge (I washed cells for one week of experiments in advance, longer storage reduces activity)

## Starting the assay

### Plate layout

My usually used plate layout looked as you can see below, testing one replicate for one substrate across the whole differentiation timeline at pH 4--7. This can of course be changed freely.

```{r, echo = FALSE, include = TRUE, warning=FALSE, fig.height=4, fig.cap="Plate layout of a typical LAC acitivty assay."}
plate <- tibble(
  row = rep(LETTERS[1:8], length.out = 96, each = 12),
  column = rep(1:12, length.out = 96, each = 1),
  sample = rep(c("nTE", "TE"), each = 12, length.out = 96),
  pH = rep(c(4:7), each = 24, length.out = 96),
  day = rep(c(0:9, 14, 21), length.out = 96),
  substrate = ordered("ABTS",
    levels = c("ABTS", "DAF", "MBTH", "PYGL", "SGZ")
  ),
  concentration = 5,
  replicate = 5
) %>%
  mutate(sample = case_when(
    row %in% c("B", "D", "F", "H") & column == 1 ~ "neg. ctrl.",
    TRUE ~ sample
  ))

ggplot(plate) +
  geom_tile(aes(x = column, y = reorder(row, desc(row)), fill = substrate, alpha = pH),
    colour = rgb(1, 1, 1, 1),
    # fill = "#66CCCC",
    size = 2
  ) +
  geom_text(
    data = plate %>% filter(sample != "neg. ctrl."),
    aes(
      x = column, y = reorder(row, desc(row)),
      label = paste(sample, "\nDPI:", day)
    ),
    size = 2, family = "IBMPlexSans"
  ) +
  geom_text(
    data = plate %>% filter(sample == "neg. ctrl."),
    aes(
      x = column, y = reorder(row, desc(row)),
      label = "NEGATIVE\nCONTROL"
    ),
    size = 2, family = "IBMPlexSans"
  ) +
  # labs(title = paste("Replicate", unique(plate$replicate))) +
  scale_x_continuous(breaks = c(1:12), position = "top", expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_fill_manual(
    values = c(
      "#66CCCC",
      "#6699FF",
      "#D35877",
      "#FFCC66",
      "#FF99CC"
    ),
    drop = FALSE
  ) +
  theme_leo() +
  theme(
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    legend.position = "right"
  ) +
  coord_fixed()
```

### Reaction mix

The final reaction mix contains:

* 100 µl 0.2M buffer (half strength -- *i.e.* 100 µl 0.1M buffer -- for SGZ, otherwise it precipitates)
* 60 µl washed cells (H~2~O for the negative control)
* 40 µl of appropriately diluted substrate (see table) --- added after the first two measurements

### Hidex program

The used program of the Hidex plate reader should be named something like "TE_activity". It measures the whole spectrum every ten minutes and includes a pause after 10 min (two measurements), to add the substrate. The average of the first two measurements is used to estimate cell density (detailed below).

# Data analysis

I'll include the basic R code here, I assume you won't need it but I guess it can't hurt.

## Load data

For each experiment, I have the Hidex output in one .csv file and the plate layout in another .csv file. They're matched by date and start time (morning or evening). How you do it doesn't matter of course, as long as you know which sample/substrate is in each well in the end.

### Load plate layouts

```{r}
plate_files <-
  list.files(
    path = paste0(datapath, "PhD/Laccase_activity/2021-04_TE_activity/csv_files/plates/"),
    pattern = "*.csv$",
    recursive = FALSE,
    full.names = TRUE
  )

read_plus <- function(flnm) {
  read_csv(flnm, show_col_types = FALSE) %>%
    mutate(
      filename = basename(flnm)
    ) %>%
    separate(filename, into = c("date", "started"), sep = "_") %>%
    mutate(
      Well = paste0(row, column),
      started = str_remove(started, fixed(".csv")),
      date = ymd(date),
      pH = as.character(pH)
    ) %>%
    select(-row, -column)
}

TE_plates <- map_dfr(plate_files, read_plus)

head(TE_plates)
```

### Load absorbance data

Defining the absorbance maxima and minima (for background subtraction).

```{r}
abs_max <- tibble(
  substrate = c("ABTS", "DAF", "MBTH", "PYGL", "SGZ"),
  abs_max = c(420, 610, 505, 450, 530),
  abs_min = c(480, 930, 700, 600, 600)
)
```

Loading the .csv files with the Hidex data

```{r}
spec_files <-
  list.files(
    path = paste0(datapath, "PhD/Laccase_activity/2021-04_TE_activity/csv_files/absorbance/"),
    pattern = "*.csv$",
    recursive = FALSE,
    full.names = TRUE
  )

read_vroom <- function(flnm) {
  read_csv(flnm, show_col_types = FALSE) %>%
    mutate(
      filename = basename(flnm)
    ) %>%
    separate(filename, into = c("author", "assay", "date", "time"), sep = "_") %>%
    select(-author, -assay) %>%
    mutate(
      time = str_remove(time, fixed(".csv")),
      date = ymd(date),
      started = ymd_hms(paste(date, time, sep = "_")),
      started = case_when(
        hour(started) < 14 ~ "morning",
        TRUE ~ "evening"
      )
    ) %>%
    select(-time) %>%
    pivot_longer(c(`220`:`1000`), names_to = "wavelength", values_to = "measurement") %>%
    filter(wavelength %in% c(abs_max$abs_max, abs_max$abs_min))
}

TE_spec <- map_dfr(spec_files, read_vroom) %>%
  left_join(TE_plates, by = c("date", "started", "Well")) %>%
  drop_na()

head(TE_spec)
```

Since I didn't use the autocorrection for the length of the lightpath, I adjust the absorbance manually to yield results equivalent to measurements taken with a 1 cm lightpath. Lastly, I adjust the absorbance for cell density (measured as OD~600~ before adding the substrate) and zero the measurements at the absorbance directly after substrate addition.

```{r}
TE_data <- TE_spec %>%
  mutate(
    wavelength = as.numeric(wavelength),
    absorbance = measurement / 0.52, # experimental path length in the plate
    time = (`Cycle #` * 10) - 10 # 10 minutes between measurements
  ) %>%
  filter(time < 390) %>%
  group_by(date, replicate, substrate, pH, wavelength, time) %>%
  mutate("blank" = absorbance[sample == "neg. ctrl."]) %>%
  group_by(date, sample, substrate, pH, replicate, day) %>%
  mutate(
    density = mean(absorbance[wavelength == 600 & `Cycle #` %in% c(1, 2)]),
    adjusted_absorbance = case_when(
      sample == "neg. ctrl." ~ absorbance, # do not adjust the negative control
      TRUE ~ (absorbance - blank) / density # adjust rest for density
    ),
    # zero absorbance at first measurement after substrate addition
    zeroed_absorbance = adjusted_absorbance - adjusted_absorbance[`Cycle #` == 3]
  ) %>%
  select(-`Cycle #`, -`Time (s)`)

rm(TE_spec)
# gc()

head(TE_data)
```

## Visualise data

### Plot corrected absorbance over time

Here I subtract the absorbance spectrum of the negative control (substrate, buffer, H~2~O) at each time point to account for auto-oxidation.

```{r, message=FALSE}
kinetics <- TE_data %>%
  left_join(abs_max) %>%
  group_by(
    date, sample, substrate, pH, replicate, day, time, density
  ) %>%
  mutate(corrected_absorbance = zeroed_absorbance - zeroed_absorbance[wavelength == abs_min]) %>%
  summarise(
    corrected_absorbance = corrected_absorbance[wavelength == abs_max],
    adjusted_absorbance = adjusted_absorbance[wavelength == abs_max],
    absorbance = absorbance[wavelength == abs_max]
  )
```

Then we plot the corrected absorbance over time.

```{r, message=FALSE, echo=FALSE}
ggplot(
  kinetics %>% filter(sample == "TE"),
  aes(
    x = time,
    y = corrected_absorbance,
    colour = pH
  )
) +
  stat_smooth(
    geom = "line",
    aes(group = interaction(replicate, pH)),
    alpha = 0.5,
    size = 0.2,
    se = F
  ) +
  scale_colour_manual(values = pal_dawn[c(1, 3, 5, 6)]) +
  scale_x_continuous(breaks = c(100, 300)) +
  labs(
    x = "Time [min]",
    y = "Corrected absorbance"
  ) +
  theme_leo() +
  theme(
    legend.position = "bottom",
    axis.ticks.x = element_line(
      size = 0.25,
      lineend = "square",
      color = "black"
    )
  ) +
  facet_grid(substrate ~ day,
    scales = "free_y"
  )
```

### Identify linear phase

From these figures, I identify the pH-dependent near-linear phase of oxidation by eye.

```{r, message=FALSE, echo=FALSE}
lm_bounds <- tribble(
  ~substrate, ~pH, ~lm_start, ~lm_stop,
  "ABTS", 4, 40, 160,
  "ABTS", 5, 40, 160,
  "ABTS", 6, 100, 400,
  "ABTS", 7, 100, 400,
  "DAF", 4, 40, 160,
  "DAF", 5, 40, 160,
  "DAF", 6, 20, 100,
  "DAF", 7, 100, 300,
  "MBTH", 4, 100, 300,
  "MBTH", 5, 100, 300,
  "MBTH", 6, 100, 300,
  "MBTH", 7, 20, 120,
  "PYGL", 4, 100, 400,
  "PYGL", 5, 100, 400,
  "PYGL", 6, 100, 400,
  "PYGL", 7, 100, 300,
  "SGZ", 4, 200, 400,
  "SGZ", 5, 200, 400,
  "SGZ", 6, 200, 400,
  "SGZ", 7, 200, 400,
) %>%
  mutate(pH = as.character(pH))

kable(lm_bounds,
  "latex",
  caption = "Linear phases of substrate oxidation.",
  col.names = c(
    "Substrate",
    "pH",
    "Start [min]",
    "End [min]"
  ),
  escape = F,
  booktabs = T,
  linesep = ""
) %>%
  kable_styling(latex_options = c(
    "striped",
    "HOLD_position"
    # "scale_down"
  )) 
```


```{r, message=FALSE, echo=FALSE}
kinetics <- kinetics %>%
  left_join(lm_bounds)

ggplot(
  kinetics %>% filter(sample == "TE"),
  aes(
    x = time,
    y = corrected_absorbance,
    colour = pH
  )
) +
  stat_smooth(
    geom = "line",
    aes(group = interaction(replicate, pH)),
    alpha = 0.5,
    size = 0.2,
    se = F
  ) +
  geom_smooth(
    data = filter(kinetics, sample == "TE" & time %in% c(lm_start:lm_stop)),
    aes(group = interaction(replicate, pH)),
    method = "lm",
    size = 0.5,
    se = F
  ) +
  scale_colour_manual(values = pal_dawn[c(1, 3, 5, 6)]) +
  scale_x_continuous(breaks = c(100, 300)) +
  labs(
    x = "Time [min]",
    y = "Corrected absorbance"
  ) +
  theme_leo() +
  theme(
    legend.position = "bottom",
    axis.ticks.x = element_line(
      size = 0.25,
      lineend = "square",
      color = "black"
    )
  ) +
  facet_grid(substrate ~ day,
    scales = "free_y"
  )
```

## Plot linear phase slope

### Calculate slopes

Now I calculate the slopes of linear regressions of the (near-)linear phases. I save the R^2^, but I would avoid filtering by it (*i.e.* using only replicates with R^2^ over a certain value), because the R^2^ is always super low when activity is minimal (*i.e.* a horizontal line). Negative slopes are set to 0.

```{r, message=FALSE}
kinetics_nested <- kinetics %>%
  filter(time >= lm_start & time <= lm_stop) %>%
  select(-absorbance, -adjusted_absorbance) %>%
  nest(data = c(time, corrected_absorbance)) %>%
  mutate(
    fit = map(data, ~ lm(corrected_absorbance ~ time, data = .x)),
    tidied = map(fit, tidy),
    model = map(fit, glance),
    adj.r.squared = unlist(map(model, ~ .x[["adj.r.squared"]]))
  ) %>%
  select(-model) %>%
  unnest(tidied) %>%
  filter(term == "time") %>%
  mutate(pH = as.factor(pH)) %>%
  group_by(substrate, pH, replicate, sample) %>%
  mutate(
    scaled_estimate = estimate / max(estimate),
    scaled_estimate = case_when(
      scaled_estimate < 0 ~ 0,
      TRUE ~ scaled_estimate
    )
  )
```

### Plot slopes with SD

```{r, message=FALSE, echo=FALSE}

kinetics_avg <- kinetics_nested %>%
  filter(sample != "neg. ctrl.") %>%
  group_by(substrate, sample, day, pH) %>%
  summarise(
    mean_slope = mean(estimate),
    sd_slope = sd(estimate)
  )

ggplot(
  kinetics_avg,
  aes(
    x = day,
    y = mean_slope,
    colour = substrate
  )
) +
  annotate("rect",
    ymin = -Inf,
    ymax = Inf,
    xmin = 5,
    xmax = 21,
    fill = "grey95"
  ) +
  geom_ribbon(
    aes(
      group = sample,
      fill = substrate,
      ymin = mean_slope - sd_slope,
      ymax = mean_slope + sd_slope
    ),
    colour = NA,
    alpha = 0.25
  ) +
  geom_line(
    aes(linetype = sample),
    size = 0.4,
  ) +
  # scale_colour_brewer(palette = "Set1") +
  scale_colour_manual(
    aesthetics = c("colour", "fill"),
    values = c(
      "#66CCCC",
      "#6699FF",
      "#D35877",
      "#FFCC66",
      "#FF99CC"
    )
  ) +
  scale_linetype_manual(values = c(2, 1)) +
  scale_y_continuous(labels = function(x) format(x, scientific = TRUE, digits = 1)) +
  expand_limits(y = 0) +
  labs(
    y = "Substrate oxidation rate [AU min<sup>-1</sup>]",
    x = "Days past induction (dpi)"
  ) +
  theme_leo() +
  theme(
    strip.text = element_blank(),
    legend.position = "top",
    axis.title.y = element_textbox_simple(
      orientation = "left-rotated",
      halign = 0.5,
      width = unit(100, "mm")
    )
  ) +
  facet_wrap(pH ~ substrate, scales = "free_y")
```
