---
title: "piecewise SEM of IRX phenotypes"
author: "Leonard Blaschek"
date: "26 March 2019"
output: pdf_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(showtext)
library(ggthemes)
library(nlme)
library(piecewiseSEM)
library(Hmisc)
library(zoo)
library(ggridges)
library(ggrepel)
library(tidyverse)

font_add(
  "Helvetica",
  regular = "/prop_fonts/01. Helvetica     [1957 - Max Miedinger]/HelveticaNeueLTStd-Lt.otf",
  italic = "/prop_fonts/01. Helvetica     [1957 - Max Miedinger]/HelveticaNeueLTStd-LtIt.otf",
  bold = "/prop_fonts/01. Helvetica     [1957 - Max Miedinger]/HelveticaNeueLTStd-Bd.otf",
  bolditalic = "/prop_fonts/01. Helvetica     [1957 - Max Miedinger]/HelveticaNeueLTStd-BdIt.otf"
)
showtext_auto()

#### generating plot theme ####
theme_leo <- function(base_size = 12,
                      base_family = "Helvetica"){
  theme_minimal(base_size = base_size,
                base_family = base_family) %+replace%
    theme(
      strip.text = element_text(hjust = 0, face = "italic"),
      axis.ticks = element_line(
        size = 0.25,
        lineend = "square",
        color = "black"
      ),
      axis.text.x = element_text(colour = "black", # flipped coords
                                 margin = margin(1, 1, 1, 1)),
      axis.text.y = element_text(
        colour = "black",
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1)
      ),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", size = 0.25),
      panel.spacing = unit(1.5, "mm"),
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text = element_text(size = rel(0.8)),
      legend.key.height = unit(4, "mm"),
      legend.key.width = unit(30, "mm"),
      # plot.margin = unit(c(0, 0, 0, 0), "cm"),   
      
      complete = TRUE
    )
}
```

![First model, including all measured parameters that are implicated by logic or the literature](/home/leonard/Documents/Uni/PhD/Courses/SEM/causal_model_1.pdf)

### Loading the Data

```{r load data, message = FALSE, include=TRUE}
library(dplyr)
library(tidyr)
library(ggplot2)
sem.data <- read.csv("file:///home/leonard/R/Output/PhD/SEM_data.csv")

sem.data$replicate <- as.factor(sem.data$replicate)
sem.data$technical <- as.factor(sem.data$technical)

# calculate height and width relative to perimeter
sem.data <- sem.data %>%
  mutate(rel.Height = Height/Perim.,
         rel.Width = Width/Perim.,
         HeightByWidth = Height/Width)
summary(sem.data)

ggplot(sem.data, aes(x = genotype, y = lig.peak)) +
  geom_violin() +
  geom_jitter() +
  theme_leo() +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.5)) +
  facet_wrap(~ cell.type, ncol = 3)

ggplot(sem.data, aes(x = genotype, y = Circ.)) +
  geom_violin() +
  geom_jitter() +
  scale_y_continuous(limits = c(0,1)) +
  theme_leo() +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.5)) +
  facet_wrap(~ cell.type, ncol = 3)
```

### Overall Correlations

```{r plot correlations, message = FALSE}
library(PerformanceAnalytics)
chart.Correlation(select(ungroup(sem.data), 
                         -genotype, -cell.type, -replicate, -technical), 
                  histogram=TRUE, pch=21)

ggplot(sem.data, aes(x = lig.peak, y = Circ., fill = cell.type)) +
  scale_fill_viridis_d() +
  scale_colour_viridis_d() +
  geom_point(shape = 21, alpha = 0.75, size = 3) +
  stat_smooth(method="loess", linetype = 2, se = FALSE, aes(colour = cell.type)) +
  theme_leo()

ggplot(sem.data, aes(x = Circ., y = HeightByWidth, fill = cell.type)) +
  scale_fill_viridis_d() +
  scale_colour_viridis_d() +
  geom_point(shape = 21, alpha = 0.75, size = 3) +
  stat_smooth(method="loess", linetype = 2, se = FALSE, aes(colour = cell.type)) +
  theme_leo()

ggplot(sem.data, aes(x = Circ., y = Plant.height, fill = cell.type)) +
  scale_fill_viridis_d() +
  geom_point(shape = 21, alpha = 0.75, size = 3) +
  stat_smooth(method="loess", colour = "black", linetype = 2, se = FALSE, aes(group = 1)) +
  theme_leo()
```

![Second model, now simplified and focused on the variables that are central to the hypothesis or implicated by direct correlation](/home/leonard/Documents/Uni/PhD/Courses/SEM/causal_model_2.pdf)

### Fit First pSEM Model

This model includes only primary metaxylem vessels and uses simple linear regressions without random effects.

```{r first psem}
first.model <- psem(
  lm(Circ. ~ 
       n_v +
       mean.OD1 +
       cellu.peak +
       lig.peak.pos +
       lig.peak,
     data = drop_na(filter(sem.data, cell.type == "PMX"))),
  lm(Plant.height ~ 
       Circ. +
       lig.peak.pos, 
     data = drop_na(filter(sem.data, cell.type == "PMX"))),
  data = drop_na(filter(sem.data, cell.type == "PMX"))
)
summary(first.model, .progressBar = FALSE)

plot(first.model, digits = 2)
```
 Here is the same model with a poisson regression for the plant height, increasing the explained variance to 70%. Note that the shown estimates are **not standardised**.
```{r first psem.2, message = FALSE, warning = FALSE}
first.model.poisson <- psem(
  lm(Circ. ~ 
       n_v +
       mean.OD1 +
       cellu.peak +
       lig.peak.pos +
       lig.peak,
     data = drop_na(filter(sem.data, cell.type == "PMX"))),
  glm(Plant.height ~ 
       Circ. +
       lig.peak.pos, 
       family = poisson(),
     data = drop_na(filter(sem.data, cell.type == "PMX"))),
  data = drop_na(filter(sem.data, cell.type == "PMX"))
)
summary(first.model.poisson, .progressBar = FALSE)

plot(first.model.poisson, digits = 6, show = "unstd")
```

![Third model, now including the composite variable "Cellular environment"](/home/leonard/Documents/Uni/PhD/Courses/SEM/causal_model_3.pdf)

### pSEM including all three vessel types

This model incorporates all measured vessel types (PX, PMX, SMX), as well as random effects for the cell type and the replicate.
The composite variable was dropped again to emphasise the negative effect of vessel neighbourhood on circularity.

```{r second psem}
#introduce neighbourhood composite
# summary(drop_na(sem.data))
betas <- summary(lm(Circ. ~ n_f + n_v,
     data = sem.data))$coefficients[-1,1]

sem.data$composite <- rowSums(sem.data[, c("n_f", "n_v")] * betas)

second.model <- psem(
  lme(Circ. ~
        # composite +
        n_v +
        # n_p +
        mean.OD1 +
        # cellu.peak +
        lig.peak.pos +
        lig.peak,
      random = ~1|cell.type/replicate,
      data = drop_na(sem.data)
      ),
  lme(Plant.height ~
        Circ. +
        # cellu.peak +
        lig.peak.pos +
        lig.peak,
      random = ~1|cell.type/replicate,
      data = drop_na(sem.data)
  ),
  data = drop_na(sem.data)
)

summary(second.model, .progressBar = FALSE)

plot(second.model, digits = 2)
```

### multigroup pSEM comparing IRX repsonses of the three vessel types

This third model is a multigroup model contrasting the IRX responses and their causes between vessel types. Lignin content is only a significant predictor for circularity in the PMX. However, lignin content of both PMX and SMX directly impacts plant height.

```{r multigroup}
third.model <-   psem(
  lme(Circ. ~
        n_p +
        n_f +
        mean.OD1 +
        Perim. +
        X1340.1603 +
        X1670.1603 +
        # cellu.peak +
        # lig.peak.pos +
        lig.peak,
      random = ~1|replicate/cell.type,
      data = drop_na(sem.data)
  ),
  # lme(Perim. ~
  #       # n_v,
  #       # n_f +
  #       mean.OD1 +
  #       # cellu.peak +
  #       # lig.peak.pos +
  #       lig.peak,
  #     random = ~1|replicate/cell.type,
  #     data = drop_na(sem.data)
  # ),
  lme(HeightByWidth ~
        # mean.OD1 +
        # lig.peak +
        Circ.,
      random = ~1|replicate/cell.type,
      data = drop_na(sem.data)
  ),
  # lme(rel.Height ~
  #       Circ.,
  #     random = ~1|replicate/cell.type,
  #     data = drop_na(sem.data)
  # ),
  data = drop_na(sem.data)
)

summary(third.model, .progressBar = FALSE)

plot(third.model, digits = 2)

# third.model.multigroup <- multigroup(third.model,
                                     # group = "cell.type"
# )
# third.model.multigroup

TM.multigroup <- function(x) {
  psem(
    lme(Circ. ~
          n_v +
          # n_f +
          mean.OD1 +
          Perim. +
          X1670.1603 +
          X1340.1603 +
          # cellu.peak +
          # lig.peak.pos +
          lig.peak,
        random = ~1|replicate,
        data = filter(drop_na(sem.data), cell.type == x)
    ),
    # lme(Perim. ~
    #       # n_v,
    #       # n_f +
    #       mean.OD1 +
    #       # cellu.peak +
    #       # lig.peak.pos +
    #       lig.peak,
    #     random = ~1|replicate,
    #     data = filter(drop_na(sem.data), cell.type == x)
    # ),
    lme(HeightByWidth ~
          # mean.OD1 +
          # lig.peak +
          # n_v +
          Circ.,
        random = ~1|replicate,
        data = filter(drop_na(sem.data), cell.type == x)
    ),
    # lme(rel.Height ~
    #       Circ.,
    #     random = ~1|replicate,
    #     data = filter(drop_na(sem.data), cell.type == x)
    # ),
    data = filter(drop_na(sem.data), cell.type == x)
  )
}
summary(TM.multigroup("PX"), .progressBar = FALSE)
plot(TM.multigroup("PX"))
summary(TM.multigroup("PMX"), .progressBar = FALSE)
plot(TM.multigroup("PMX"))
summary(TM.multigroup("SMX"), .progressBar = FALSE)
plot(TM.multigroup("SMX"))

summary(lme(Circ. ~
          n_v +
          # n_f +
          mean.OD1 +
          Perim. +
          X1340.1603 +
          # cellu.peak +
          # lig.peak.pos +
          lig.peak,
        random = ~1|replicate,
        data = filter(drop_na(sem.data), cell.type == "PMX")
    ))
```

```{r average data}
avg.data <- read.csv("SEM_data_avg.csv")
third.model <-   psem(
  lme(Circ. ~
        n_v +
        # n_f +
        mean.OD1 +
        Perim. +
        # cellu.peak +
        # lig.peak.pos +
        lig.peak,
      random = ~1|replicate/cell.type,
      data = drop_na(avg.data)
  ),
  # lme(Perim. ~
  #       # n_v,
  #       # n_f +
  #       mean.OD1 +
  #       # cellu.peak +
  #       # lig.peak.pos +
  #       lig.peak,
  #     random = ~1|replicate/cell.type,
  #     data = drop_na(sem.data)
  # ),
  # lme(HeightByWidth ~
  #       # mean.OD1 +
  #       # lig.peak +
  #       Circ.,
  #     random = ~1|replicate/cell.type,
  #     data = drop_na(avg.data)
  # ),
  # lme(rel.Height ~
  #       Circ.,
  #     random = ~1|replicate/cell.type,
  #     data = drop_na(sem.data)
  # ),
  data = drop_na(avg.data)
)

summary(third.model, .progressBar = FALSE)

plot(third.model, digits = 2)
```

```{r, corrplots, warning=FALSE, fig.showtext = TRUE, out.width = '100%', message = FALSE}
#### calculate linear regressions ####
raman_spectra <- read_csv("raman_spectra_corrected.csv", col_types = "ccccddd") 
raman_irx <- read_csv("raman_irx_shape.csv", col_types = "ccccddddddddddd")
raman_corr <- right_join(raman_spectra, select(raman_irx, genotype:cell.type, Circ.))
raman_corr_switch <- left_join(read_tsv("/home/leonard/Documents/Uni/PhD/IRX/PMX_which_SMX.tsv",
                                       col_types = "cccc"), raman_corr) %>%
  mutate(technical = str_c(technical, "switched", sep = "_"),
         cell.type = "SMX")
raman_corr <- raman_corr %>%
  anti_join(read_tsv("/home/leonard/Documents/Uni/PhD/IRX/PMX_which_SMX.tsv",
                     col_types = "cccc")) %>%
  bind_rows(raman_corr_switch)

lin.reg <- function(x) {
  corr.mat <- as.matrix(x[, c("corrected.intensity", "Circ.")])
  corr <- data.frame(rcorr(corr.mat)$r)
  corr$p <- data.frame(rcorr(corr.mat)$P)
  corr["wavenumber"] <- unique(as.numeric(x$wavenumber))
  reg <- lm(Circ. ~ corrected.intensity, data = x)
  mod <- summary(reg)["adj.r.squared"]
  return(c(corr, mod))
}

irx_raman_corr <- drop_na(raman_corr) %>%
  group_by(wavenumber, cell.type) %>%
  do(data.frame(lin.reg(.))) %>%
  select(-Circ., -p.Circ.) %>%
  rename(r = corrected.intensity,
         p = p.corrected.intensity) %>%
  drop_na() %>%
  full_join(raman_corr %>% 
              ungroup() %>% 
              group_by(cell.type, wavenumber) %>% 
              summarise(intensity.mean = mean(corrected.intensity, na.rm = TRUE))) %>%
  filter(wavenumber > 300 & wavenumber < 2000)

plot_correlations <- function(x, y, z) {
ggplot(data = filter(x, cell.type == y)) +
  geom_ridgeline_gradient(aes(
    height = max(intensity.mean) - rollmean(intensity.mean, 3, na.pad = TRUE),
    fill = r,
    y = rollmean(intensity.mean, 3, na.pad = TRUE),
    x = wavenumber
  ),
  colour = "white"
  ) +
  scale_x_reverse(limits = c(2000, 300)) +
  labs(
    title = y,
    x = "Wavenumber",
    y = "Absolute intensity"
  ) +
  scale_fill_distiller(palette = z) +
  scale_colour_distiller(palette = z) +
  theme_leo() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(4, "mm")
  ) +
  guides(fill = guide_legend(), colour = "none") +
  geom_label_repel(data = filter(x, 
                                 cell.type == y & 
                                   (r > (max(x$r[x$cell.type == y]) - 0.02) | 
                                      r < (min(x$r[x$cell.type == y]) + 0.02))),
                   aes(x = wavenumber,
                       y = intensity.mean,
                       label = wavenumber,
                       colour = r),
    nudge_y = 750,
    family = "Helvetica",
    segment.colour = "black",
    size = 2
  )
}

plot_correlations(irx_raman_corr, "PX", "PuOr")
plot_correlations(irx_raman_corr, "PMX", "PuOr")
plot_correlations(irx_raman_corr, "SMX", "PuOr")
```

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE}
raman_spectra <- read_csv("raman_spectra_corrected.csv", col_types = "ccccddd") 
raman_irx <- read_csv("raman_irx_shape.csv", col_types = "ccccddddddddddd")
raman_corr <- right_join(raman_spectra, select(raman_irx, genotype:cell.type, Circ.))
raman_corr_switch <- left_join(read_tsv("/home/leonard/Documents/Uni/PhD/IRX/PMX_which_SMX.tsv",
                                       col_types = "cccc"), raman_corr) %>%
  mutate(technical = str_c(technical, "switched", sep = "_"),
         cell.type = "SMX")
raman_corr <- raman_corr %>%
  anti_join(read_tsv("/home/leonard/Documents/Uni/PhD/IRX/PMX_which_SMX.tsv",
                     col_types = "cccc")) %>%
  bind_rows(raman_corr_switch) %>%
  group_by(genotype, cell.type, replicate, technical) %>%
  mutate(intensity.norm = corrected.intensity / corrected.intensity[wavenumber == 1603])

lin.reg <- function(x) {
  corr.mat <- as.matrix(x[, c("intensity.norm", "Circ.")])
  corr <- data.frame(rcorr(corr.mat)$r)
  corr$p <- data.frame(rcorr(corr.mat)$P)
  corr["wavenumber"] <- unique(as.numeric(x$wavenumber))
  reg <- lm(Circ. ~ intensity.norm, data = x)
  mod <- summary(reg)["adj.r.squared"]
  return(c(corr, mod))
}

irx_raman_corr <- drop_na(raman_corr) %>%
  group_by(wavenumber, cell.type) %>%
  do(data.frame(lin.reg(.))) %>%
  select(-Circ., -p.Circ.) %>%
  rename(r = intensity.norm,
         p = p.intensity.norm) %>%
  drop_na() %>%
  full_join(raman_corr %>% 
              ungroup() %>% 
              group_by(cell.type, wavenumber) %>% 
              summarise(intensity.mean = mean(intensity.norm, na.rm = TRUE))) %>%
  filter(wavenumber > 300 & wavenumber < 2000) %>%
  drop_na()

plot_correlations <- function(x, y, z) {
ggplot(data = filter(x, cell.type == y)) +
  geom_ridgeline_gradient(aes(
    height = max(intensity.mean) - rollmean(intensity.mean, 3, na.pad = TRUE),
    fill = r,
    y = rollmean(intensity.mean, 3, na.pad = TRUE),
    x = wavenumber
  ),
  colour = "white"
  ) +
  scale_x_reverse(limits = c(2000, 300)) +
  labs(
    title = y,
    x = "Wavenumber",
    y = "Intensity normalised to 1603"
  ) +
  scale_fill_distiller(palette = z) +
  scale_colour_distiller(palette = z) +
  theme_leo() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(4, "mm")
  ) +
  guides(fill = guide_legend(), colour = "none") +
  geom_label_repel(data = filter(x, 
                                 cell.type == y & 
                                   (r > (max(x$r[x$cell.type == y]) - 0.04) | 
                                      r < (min(x$r[x$cell.type == y]) + 0.04))),
                   aes(x = wavenumber,
                       y = intensity.mean,
                       label = wavenumber,
                       colour = r),
    nudge_y = 0.25,
    family = "Helvetica",
    segment.colour = "black",
    size = 2
  )
}

plot_correlations(irx_raman_corr, "PX", "PuOr")
plot_correlations(irx_raman_corr, "PMX", "PuOr")
plot_correlations(irx_raman_corr, "SMX", "PuOr")
```