---
title: "piecewise SEM of IRX phenotypes"
author: "Leonard Blaschek"
date: "26 March 2019"
output: pdf_document
always_allow_html: yes
---

```{r setup, include=FALSE, fig.width=10, fig.height=5, fig.path='figures/', dev=c('png', 'pdf')}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(ggthemes)
library(nlme)
library(piecewiseSEM)

#### generating plot theme ####
theme_leo <- function(base_size = 12,
                      base_family = "Helvetica"){
  theme_minimal(base_size = base_size,
                base_family = base_family) %+replace%
    theme(
      strip.text = element_text(hjust = 0, face = "italic"),
      axis.ticks = element_line(
        size = 0.25,
        lineend = "square",
        color = "black"
      ),
      axis.text.x = element_text(colour = "black", # flipped coords
                                 margin = margin(1, 1, 1, 1)),
      axis.text.y = element_text(
        colour = "black",
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1)
      ),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", size = 0.25),
      panel.spacing = unit(1.5, "mm"),
      legend.position = "bottom",
      legend.text = element_text(size = rel(0.8)),
      legend.key.height = unit(4, "mm"),
      legend.key.width = unit(30, "mm"),
      # plot.margin = unit(c(0, 0, 0, 0), "cm"),   
      
      complete = TRUE
    )
}
```

![First model, including all measured parameters that are implicated by logic or the literature](/home/leonard/Dropbox/SEM course/causal_model_1.pdf)

### Loading the Data


```{r load data, message = FALSE, include=TRUE}
library(dplyr)
library(tidyr)
library(ggplot2)
sem.data <- read.csv("file:///home/leonard/R/Output/PhD/SEM_data.csv") %>%
  select(-X, -X1599...cm....1.1119..cm....1)

summary(sem.data)

ggplot(sem.data, aes(x = genotype, y = Integrated.lignin.peak)) +
  geom_violin() +
  geom_jitter() +
  theme_leo() +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.5)) +
  facet_wrap(~ cell.type, ncol = 3)

ggplot(sem.data, aes(x = genotype, y = Circ.)) +
  geom_violin() +
  geom_jitter() +
  scale_y_continuous(limits = c(0,1)) +
  theme_leo() +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.5)) +
  facet_wrap(~ cell.type, ncol = 3)
```

### Overall Correlations

```{r plot correlations, message = FALSE}
library(PerformanceAnalytics)
chart.Correlation(select(ungroup(sem.data), 
                         -genotype, -cell.type, -replicate, -technical), 
                  histogram=TRUE, pch=21)

ggplot(sem.data, aes(x = Integrated.lignin.peak, y = Circ., fill = cell.type)) +
  scale_fill_viridis_d() +
  scale_colour_viridis_d() +
  geom_point(shape = 21, alpha = 0.75, size = 3) +
  stat_smooth(method="loess", linetype = 2, se = FALSE, aes(colour = cell.type)) +
  theme_leo()

ggplot(sem.data, aes(x = Circ., y = Plant.height, fill = cell.type)) +
  scale_fill_viridis_d() +
  geom_point(shape = 21, alpha = 0.75, size = 3) +
  stat_smooth(method="loess", colour = "black", linetype = 2, se = FALSE, aes(group = 1)) +
  theme_leo()
```

![Second model, now simplified and focused on the variables that are central to the hypothesis or implicated by direct correlation](/home/leonard/Dropbox/SEM course/causal_model_2.pdf)

### Fit First pSEM Model

This model includes only primary metaxylem vessels and uses simple linear regressions without random effects.

```{r first psem}
first.model <- psem(
  lm(Circ. ~ 
       n_v +
       mean.OD1 +
       Integrated.cellulose.peak +
       Lignin.peak.position +
       Integrated.lignin.peak,
     data = drop_na(filter(sem.data, cell.type == "PMX"))),
  lm(Plant.height ~ 
       Circ. +
       Lignin.peak.position, 
     data = drop_na(filter(sem.data, cell.type == "PMX"))),
  data = drop_na(filter(sem.data, cell.type == "PMX"))
)
summary(first.model, .progressBar = FALSE)

plot(first.model, digits = 2)
```
 Here is the same model with a poisson regression for the plant height, increasing the explained variance to 70%. Note that the shown estimates are **not standardised**.
```{r first psem.2, message = FALSE, warning = FALSE}
first.model.poisson <- psem(
  lm(Circ. ~ 
       n_v +
       mean.OD1 +
       Integrated.cellulose.peak +
       Lignin.peak.position +
       Integrated.lignin.peak,
     data = drop_na(filter(sem.data, cell.type == "PMX"))),
  glm(Plant.height ~ 
       Circ. +
       Lignin.peak.position, 
       family = poisson(),
     data = drop_na(filter(sem.data, cell.type == "PMX"))),
  data = drop_na(filter(sem.data, cell.type == "PMX"))
)
summary(first.model.poisson, .progressBar = FALSE)

plot(first.model.poisson, digits = 6, show = "unstd")
```

![Third model, now including the composite variable "Cellular environment"](/home/leonard/Dropbox/SEM course/causal_model_3.pdf)

### pSEM including all three vessel types

This model incorporates all measured vessel types (PX, PMX, SMX), as well as random effects for the cell type and the replicate.
The composite variable was dropped again to emphasise the neagtive effect of vessel neighbourhood on circularity.

```{r second psem}
#introduce neighbourhood composite
betas <- summary(lm(Circ. ~ n_f + n_v,
     data = sem.data))$coefficients[-1,1]

sem.data$composite <- rowSums(sem.data[, c("n_f", "n_v")] * betas)

second.model <- psem(
  lme(Circ. ~ 
        # composite +
        n_v +
        # n_p +
        # mean.OD1 +
        Integrated.cellulose.peak +
        Integrated.lignin.peak +
        Lignin.peak.position,
      random = ~1|cell.type/replicate,
      data = drop_na(sem.data)),
  lme(Plant.height ~ 
        Circ. +
        Integrated.lignin.peak +
        Integrated.cellulose.peak +
        Lignin.peak.position,
      random = ~1|cell.type/replicate,
      data = drop_na(sem.data)
  ),
  data = drop_na(sem.data)
)

summary(second.model, .progressBar = FALSE)

plot(second.model, digits = 2)
```

### multigroup pSEM comparing IRX repsonses of the three vessel types

This third model is a multigroup model contrasting the IRX responses and their causes between vessel types. Lignin content is only a significant predictor for circularity in the PMX. However, lignin content of both PMX and SMX directly impacts plant height.

```{r multigroup}
third.model <- psem(
  lme(Circ. ~ 
        n_v +
        # n_f +
        # mean.OD1 +
        Integrated.cellulose.peak +
        Integrated.lignin.peak +
        Lignin.peak.position,
      random = ~1|replicate,
      data = drop_na(sem.data)),
  lme(Plant.height ~ 
        Circ. +
        Integrated.lignin.peak +
        Integrated.cellulose.peak +
        Lignin.peak.position,
      random = ~1|replicate,
      data = drop_na(sem.data)
  ),
  data = drop_na(sem.data)
)

# summary(third.model, .progressBar = FALSE)

plot(third.model, digits = 2)

third.model.multigroup <- multigroup(third.model,
                                      group = "cell.type"
                                      )
third.model.multigroup

TM.multigroup <- function(x) {
  psem(
    lme(Circ. ~ 
          n_v +
          # n_f +
          mean.OD1 +
          # Integrated.cellulose.peak +
          Integrated.lignin.peak +
          Lignin.peak.position,
        random = ~1|replicate,
        data = filter(drop_na(sem.data), cell.type == x)),
    lme(Area ~ 
          n_v +
          # n_f +
          mean.OD1 +
          # Integrated.cellulose.peak +
          Integrated.lignin.peak +
          Lignin.peak.position,
        random = ~1|replicate,
        data = filter(drop_na(sem.data), cell.type == x)),
    lme(Perim. ~ 
          n_v +
          # n_f +
          mean.OD1 +
          # Integrated.cellulose.peak +
          Integrated.lignin.peak +
          Lignin.peak.position,
        random = ~1|replicate,
        data = filter(drop_na(sem.data), cell.type == x)),
    data = drop_na(sem.data)
  )
}

plot(TM.multigroup("PX"))
plot(TM.multigroup("PMX"))
plot(TM.multigroup("SMX"))
```


