---
title: "IRX publication figures"
author: "Leonard Blaschek"
date: "07/10/2019"
output: 
  pdf_document:
    latex_engine: lualatex
    fig_caption: yes
    fig_height: 6
    includes:
      in_header: rmd_temp.tex
sansfont: IBM Plex Sans
monofont: Inconsolata
mainfont: IBM Plex Sans
bibliography: /home/leonard/Documents/Bibliographics/zotero_lib.bib
link-citations: true
linkcolor: black
urlcolor: blue
citecolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(baseline)
library(interactions)
library(effects)
library(grid)
library(png)
library(magick)
library(agricolae)
library(ggthemes)
library(showtext)
library(zoo)
library(cowplot)
library(kableExtra)
library(tidyverse)
library(ggbeeswarm)
library(ggridges)
library(sf)
library(piecewiseSEM)
library(tukeygrps)
library(scales)
library(patchwork)


pal_ostwald_cont <- c(
  "#155DA7",
  "#0A75B9",
  # "#0C89C9",
  "#FED32F",
  # "#F8A63A",
  "#EF663A",
  "#ED4137"
)
pal_ostwald_cont2 <- c(
  "#275d95",
  "#286e9b",
  "#e8c245",
  "#d37456",
  "#d25952"
)
pal_ostwald_disc <- c(
  "#275d95",
  "#e8c245",
  "#d25952"
)
pal_ostwald_disc2 <- c(
  "#8fab1d",
  "#2d7d73",
  "#1d566f"
)
pal_ostwald_disc3 <- c(
  "#275d95",
  "#275d95",
  "#e8c245",
  "#d25952",
  "#d25952"
)
pal_flame <- c(
  "#1F265D",
  "#203F8F",
  "#404678",
  "#897B88",
  "#E7BBA2",
  "#FFEDC3",
  "#FEFEFE"
)
pal_flame_disc <- c(
  "#ffedc3",
  "#897b88",
  "#1f265d"
)
pal_ylgnbu <- c(
  "#c7e9b4",
  "#7fcdbb",
  "#41b6c4",
  "#1d91c0",
  "#225ea8",
  "#253494",
  "#081d58"
)


#### import Helvetica ####
font_add(
  "Helvetica",
  regular = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-Lt.otf",
  italic = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-LtIt.otf",
  bold = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-Bd.otf",
  bolditalic = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-BdIt.otf"
)
showtext_auto()

#### generating plot theme ####
theme_leo <- function(base_size = 6,
                      base_family = "Helvetica") {
  theme_minimal(
    base_size = base_size,
    base_family = base_family
  ) %+replace%
    theme(
      strip.text = element_text(hjust = 0, face = "italic"),
      axis.ticks = element_line(
        size = 0.125,
        lineend = "square",
        color = "black"
      ),
      axis.text.x = element_text(
        size = 6,
        colour = "black", # flipped coords
        margin = margin(1, 1, 1, 1)
      ),
      axis.text.y = element_text(
        colour = "black",
        size = 6,
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1)
      ),
      axis.title = element_text(
        colour = "black",
        size = 6
      ),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", size = 0.25),
      panel.spacing = unit(1.5, "mm"),
      legend.position = "none",
      legend.text = element_text(size = 6),
      legend.key.height = unit(4, "mm"),
      plot.title = element_text(
        size = 6,
        hjust = 0
      ),
      complete = TRUE
    )
}

scale_range <- function(x) {
  (x) / (max(x))
}

ggtext_size <- 6 / (14 / 5)
twocol <- 18 / 2.54
onehalfcol <- 14 / 2.54
onecol <- 9 / 2.54
```

## EDS data

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}

eds_data <- read_csv("/data/PhD/IRX/EDS/eds_data.csv",
  col_types = "ccnnn"
)

eds_plot <- function(num, den, lbl) {
  num <- enquo(num)
  den <- enquo(den)
  lbl <- enquo(lbl)

  p_values <- eds_data %>%
    mutate(ratio = !!num / !!den) %>%
    group_by(days) %>%
    summarise(p_value = broom::tidy(
      t.test(ratio[layer == "PCW"], ratio[layer == "SCW"])
    )[["p.value"]]) %>%
    mutate(p_value = case_when(
      p_value > 0.001 ~ round(p_value, digits = 3),
      p_value > 0.0001 ~ round(p_value, digits = 4),
      p_value > 0.00001 ~ round(p_value, digits = 5),
      p_value > 0.000001 ~ round(p_value, digits = 6),
      p_value > 0.0000001 ~ round(p_value, digits = 7),
      TRUE ~ round(p_value, digits = 8)
    ))

  eds_avg <- eds_data %>%
    mutate(min = min(!!num / !!den)) %>%
    group_by(layer, days) %>%
    summarise(
      mean = mean(!!num / !!den),
      min = min(min)
    ) %>%
    group_by(days) %>%
    summarise(
      diff = round((mean[layer == "SCW"] / mean[layer == "PCW"]) * 100, digits = 0),
      min = min(min)
    ) %>%
    left_join(p_values)

  ggplot(
    eds_data,
    aes(
      x = days,
      y = !!num / !!den,
      fill = layer
    )
  ) +
    geom_quasirandom(
      # aes(fill = cell.type, group = days),
      dodge.width = 0.6,
      shape = 21,
      # width = 0.1,
      # colour = NA,
      alpha = 0.75,
      size = 1,
      stroke = 0.2
    ) +
    geom_violin(aes(
      group = interaction(days, layer),
      # fill = cell.type
    ),
    draw_quantiles = 0.5,
    fill = "white",
    colour = "black",
    alpha = 0.85,
    width = 0.2,
    position = position_dodge(width = 0.6),
    size = 0.2,
    scale = "width"
    ) +
    geom_text(
      data = eds_avg,
      aes(
        label = paste0(diff, "%\n", "P=", p_value),
        y = 0,
        vjust = 0,
        fill = NA
      ),
      size = ggtext_size,
      family = "Helvetica",
      lineheight = 0.75,
      position = position_dodge(width = 0.6)
    ) +
    # scale_x_discrete(labels = c("PX", "PMX", "SMX")) +
    # scale_fill_manual(values = pal_flame_disc) +
    scale_fill_grey(start = 0.9, end = 0.3) +
    expand_limits(y = 0) +
    scale_y_continuous(expand = expand_scale(mult = c(0.1, 0.05))) +
    labs(x = "Days after induction") +
    theme_leo() +
    theme(
      text = element_text(family = "Helvetica"),
      legend.position = "none",
      plot.margin = unit(c(0.2, 0.2, 0.1, 0.1), "mm")
    )
}

C.Cr <- eds_plot(C, Cr) + theme(
  axis.text.x = element_blank(),
  axis.title.x = element_blank(),
  axis.ticks.x = element_blank()
) + annotate(
  geom = "segment",
  x = c(0.85, 1.15),
  xend = c(0.85, 1.15),
  y = c(6, 5.5),
  yend = c(9, 7.5),
  size = 0.2
) +
  annotate("text",
    x = c(0.85, 1.15),
    y = c(9.75, 8.25),
    label = c("PCW", "SCW"),
    family = "Helvetica",
    size = ggtext_size
  )

C.O <- eds_plot(C, O)

eds <- C.Cr / C.O

pdf("eds_plots.pdf", width = onecol * 0.7, height = onecol * 0.7)
eds
dev.off()
```

## EM drying collapse

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}
em_data <- read_csv("/data/PhD/IRX/EM/AD_collapse.csv") %>%
  pivot_longer(cols = c(PX, MX), names_to = "cell.type", values_to = "count") %>%
  mutate(
    state = ordered(state, levels = c("Fully", "Partially", "Intact")),
    cell.type = ordered(cell.type, levels = c("PX", "MX")),
    days = as.character(days),
    # days = paste(days, "days")
  ) %>%
  group_by(experiment, treatment, days, cell.type) %>%
  mutate(sum = sum(count)) %>%
  group_by(experiment, treatment, days, cell.type, state) %>%
  mutate(proportion = (count / sum) * 100)

em_data_avg <- em_data %>%
  group_by(treatment, days, cell.type, state) %>%
  summarise(
    sd = sd(proportion),
    proportion = mean(proportion)
  ) %>%
  group_by(treatment, days, cell.type) %>%
  mutate(
    ymin = case_when(
      state == "Partially" ~ proportion[state == "Intact"] +
        proportion[state == "Partially"] - sd,
      state == "Fully" ~ 0,
      state == "Intact" ~ proportion - sd
    ),
    ymax = case_when(
      state == "Partially" ~ proportion[state == "Intact"] +
        proportion[state == "Partially"] + sd,
      state == "Fully" ~ 0,
      state == "Intact" ~ proportion + sd
    )
  )

em_data_points <- em_data %>%
  select(-sum, -count) %>%
  group_by(treatment, days, cell.type, state) %>%
  pivot_wider(id_cols = -c(state, proportion), names_from = state, values_from = proportion) %>%
  mutate(
    Partially = Partially + Intact,
    Fully = Fully + Partially
  ) %>%
  pivot_longer(cols = c(Fully, Partially, Intact), names_to = "state", values_to = "proportion") %>%
  mutate(state2 = case_when(
    state == "Intact" ~ "Partially",
    state == "Partially" ~ "Fully",
    TRUE ~ "none"
  ))

em_collapse_plot <- function(cond) {
  em_collapse <- ggplot(em_data_avg %>% filter(treatment == cond), aes(
    x = days,
    y = proportion,
    # group = experiment,
    fill = state
  )) +
    geom_col(
      alpha = 0.85,
      width = 0.6
    ) +
    geom_errorbar(aes(
      ymin = ymin,
      ymax = ymax
    ),
    width = 0.1,
    size = 0.2
    ) +
    scale_fill_manual(values = rev(pal_ostwald_disc), name = "Collapse") +
    scale_y_continuous(expand = expand_scale(mult = c(0, 0.05))) +
    theme_leo() +
    labs(
      y = "Proportion",
      x = "Days after induction"
    ) +
    facet_wrap(~cell.type, nrow = 1) +
    theme(
      # axis.title.x = element_blank(),
      legend.key.height = unit(2, "mm"),
      legend.key.width = unit(2, "mm"),
      legend.margin = unit(c(0, 0, 0, 0), "mm"),
      plot.margin = unit(c(0.5, 6, 0.5, 6), "pt")
    ) +
    coord_cartesian(ylim = c(0, 100))
}

cpd_plot <- em_collapse_plot("CPD") + labs(y = "After CPD [%]")
# theme(legend.position = "bottom",
#       axis.text.x = element_blank(),
#       axis.title.x = element_blank(),
#       axis.ticks.x = element_blank())
# cpd_plot$layers[[2]]$aes_params$alpha <- 0
# cpd_plot$layers[[3]]$aes_params$alpha <- 0

ad_plot <- em_collapse_plot("AD") + labs(y = "After air drying [%]")
# theme(strip.text = element_blank())

pdf("em_collapse.pdf", width = onecol, height = onecol * 0.3)
plot_grid(
  cpd_plot,
  ad_plot,
  nrow = 1,
  rel_heights = c(1, 1)
)
dev.off()
```

## AFM ratios 

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}

afm_data <- read_csv("/data/PhD/IRX/AFM/afm_PX.csv") %>%
  tidyr::fill(everything()) %>%
  mutate(days = as.character(days)) %>%
  pivot_longer(cols = c(modulus, deformation, adhesion), names_to = "variable", values_to = "value") %>%
  group_by(cell.type, days, cell, measurement, variable) %>%
  summarise(ratio = value[layer == "SCW"] / value[layer == "PCW"]) %>%
  filter(!(variable == "adhesion" & # remove > 20-fold outlier
    cell == "Cell 4" &
    measurement == 6 &
    cell.type == "PX" &
    days == "50")) %>%
  ungroup() %>%
  mutate(cell.type = ordered(cell.type, levels = c("PX", "MX")))

rmn_violin <- function(x) {
  afm_data <- afm_data

  afm_test <- afm_data %>% filter(variable == x)
  p_values <- afm_test %>%
    group_by(cell.type) %>%
    summarise(p_value = broom::tidy(
      t.test(ratio[days == "10"], ratio[days == "50"])
    )[["p.value"]]) %>%
    mutate(p_value = case_when(
      p_value > 0.001 ~ round(p_value, digits = 3),
      p_value > 0.0001 ~ round(p_value, digits = 4),
      p_value > 0.00001 ~ round(p_value, digits = 5),
      p_value > 0.000001 ~ round(p_value, digits = 6),
      p_value > 0.0000001 ~ round(p_value, digits = 7),
      p_value > 0.000000001 ~ round(p_value, digits = 9),
      p_value > 0.0000000001 ~ round(p_value, digits = 10),
      TRUE ~ round(p_value, digits = 13)
    ))


  ggplot(
    data = afm_data %>% filter(variable == x),
    aes(x = days, y = ratio)
  ) +
    # geom_line(aes(group = interaction(cell.type, variable, cell, measurement)),
    #           size = 0.2,
    #           alpha = 0.25) +
    geom_quasirandom(
      aes(fill = cell.type, group = days),
      dodge.width = 0.6,
      shape = 21,
      # width = 0.1,
      # colour = NA,
      alpha = 0.75,
      size = 1,
      stroke = 0.2
    ) +
    # geom_boxplot(
    #   aes(group = interaction(cell.type, variable), fill = cell.type),
    #   # fill = "white",
    #   alpha = 0.5,
    #   width = 0.6,
    #   position = position_dodge2(width = 1),
    #   outlier.alpha = 0,
    #   lwd = 0.25,
    #   fatten = 1,
    #   width = 0.25
    # ) +
    geom_violin(aes(
      group = interaction(days, cell.type),
      # fill = cell.type
    ),
    draw_quantiles = 0.5,
    fill = "white",
    colour = "black",
    alpha = 0.85,
    width = 0.2,
    position = position_dodge(width = 0.6),
    size = 0.2,
    scale = "width"
    ) +
    geom_text(
      data = p_values,
      aes(label = paste(("P ="), p_value)),
      x = 0.5,
      y = 0,
      hjust = 0,
      vjust = 1,
      family = "Helvetica",
      size = ggtext_size
    ) +
    # scale_x_discrete(labels = c("PX", "PMX", "SMX")) +
    scale_fill_manual(values = pal_flame_disc) +
    expand_limits(y = 0) +
    scale_y_continuous(expand = expand_scale(mult = c(0.1, 0.05))) +
    theme_leo() +
    theme(
      text = element_text(family = "Helvetica"),
      legend.position = "none",
      # plot.margin = unit(c(5.5, 5.5, 5.5, 4.5), "pt")
    ) +
    facet_wrap(~cell.type, nrow = 2)
}

modulus <- rmn_violin("modulus") +
  labs(
    x = "Days after induction",
    y = "DMT modulus  (SCW/PCW)"
  )

deformation <- rmn_violin("deformation") +
  labs(
    x = "",
    y = "Deformation (SCW/PCW)"
  )

adhesion <- rmn_violin("adhesion") +
  labs(
    x = "",
    y = "Adhesion  (SCW/PCW)"
  )

pdf("afm_eds.pdf", width = 6.9 / 2.54, height = 13.15 / 2.54)
eds / (deformation | modulus | adhesion) +
  plot_layout(heights = c(0.5, 0.5, 1))
dev.off()
```

## Wild type vessel overview

### Load Raman data

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}
# #### import and tidy Raman data ####
# rmn_files <-
#   list.files(
#     path = "/data/PhD/IRX/RAMAN/nuoendagula/",
#     pattern = "*.txt",
#     recursive = TRUE,
#     full.names = TRUE
#   )
#
# read_plus <- function(flnm) {
#   read_tsv(flnm,
#     comment = "#",
#     col_names = FALSE,
#     skip = 25,
#     cols_only(
#       X1 = col_number(),
#       X2 = col_number()
#     )
#   ) %>%
#     mutate(
#       filename = flnm,
#       exp_time = readLines(flnm) %>%
#         str_split("\t") %>%
#         grep("Exposure Time: .* s", ., value = T) %>%
#         str_remove_all("#Exposure Time: | s") %>%
#         as.numeric(),
#       pwr = readLines(flnm) %>%
#         str_split("\t") %>%
#         grep("Excitation Power: .* mW", ., value = T) %>%
#         str_remove_all("#Excitation Power: | mW") %>%
#         as.numeric(),
#       nd_fltr = 1 - (readLines(flnm) %>%
#         str_split("\t") %>%
#         grep("% \\([0-9]{3}/255\\)", ., value = T) %>%
#         str_remove_all("#ND Filter : | % \\([0-9]{3}/255\\)") %>%
#         as.numeric() / 100)
#     )
# }
#
# rmn_data <- lapply(rmn_files, read_plus) %>%
#   bind_rows()
#
# # rmn_data <- rmn_data %>%
# #   select(c(1:3))
#
# rmn_data$filename <- basename(rmn_data$filename)
#
# rmn_data$filename <-
#   str_replace(rmn_data$filename, fixed("fah 1"), "fah1")
#
# rmn_data <- rmn_data %>%
#   distinct(X1, filename, .keep_all = TRUE) %>%
#   separate(
#     filename,
#     into = c("genotype", "replicate", "cell.type", "technical"),
#     sep = "\\s*#|-|\\s",
#     extra = "merge"
#   ) %>%
#   rename("wavenumber" = X1, "intensity" = X2) %>%
#   mutate(
#     genotype = recode(
#       genotype,
#       "4cl1＆2" = "4cl1x4cl2",
#       "cad4＆5" = "cad4xcad5",
#       "ccoaomt" = "ccoaomt1",
#       "ccr1" = "ccr1-3",
#       "col.o" = "Col-0",
#       "Col.0" = "Col-0",
#       "ccr1＆fah1" = "ccr1xfah1"
#     ),
#     cell.type = recode(cell.type,
#       "px" = "PX",
#       "SX" = "SMX",
#       "MX" = "PMX"
#     ),
#     replicate = str_extract(
#       replicate,
#       "\\d"
#     ),
#     technical = str_extract(
#       technical,
#       "(\\d)+"
#     ),
#     wavenumber = round(wavenumber, digits = 0)
#   ) %>%
#   filter(cell.type != "？？") %>%
#   group_by(genotype, cell.type, replicate, technical) %>%
#   mutate(intensity = intensity / (exp_time * pwr * nd_fltr)) %>%
#   select(-exp_time, -pwr, -nd_fltr)
#
# #### baseline correct Raman data ####
# rmn_data_wide <- rmn_data %>%
#   group_by(genotype, cell.type, replicate, technical) %>%
#   mutate(group_id = row_number()) %>%
#   spread(wavenumber, intensity) %>%
#   select(-group_id) %>%
#   summarise_all(funs(sum(., na.rm = TRUE)))
#
# rmn_data_mat <- as.matrix(rmn_data_wide[, -c(1:4)])
# rownames(rmn_data_mat) <- paste(rmn_data_wide$genotype,
#   rmn_data_wide$cell.type,
#   rmn_data_wide$replicate,
#   rmn_data_wide$technical,
#   sep = "_"
# )
#
# rmn_data_correction <- baseline.als(rmn_data_mat,
#   lambda = 5,
#   p = 0.01,
#   maxit = 100
# )
# rmn_data_corrected <- as_tibble(rmn_data_correction$corrected, rownames = NA)
#
# rmn_data_corrected <- rmn_data_corrected %>%
#   rownames_to_column() %>%
#   separate(rowname, sep = "_", into = c("genotype", "cell.type", "replicate", "technical"), remove = TRUE) %>%
#   gather(key = "wavenumber", value = "corrected.intensity", -c(genotype, cell.type, replicate, technical)) %>%
#   mutate(wavenumber = str_replace(wavenumber, pattern = fixed("X."), replacement = "-")) %>%
#   mutate(wavenumber = str_replace(wavenumber, pattern = fixed("X"), replacement = "")) %>%
#   mutate(wavenumber = as.numeric(wavenumber)) %>%
#   group_by(genotype, cell.type, replicate, technical) %>%
#   inner_join(., rmn_data, by = c("genotype", "cell.type", "replicate", "technical", "wavenumber")) %>%
#   ungroup() %>%
#   filter(genotype != "ccr1xfah1" &
#     wavenumber %in% c(300:1700)) %>%
#   mutate(
#     genotype = ordered(genotype,
#       levels = c(
#         "Col-0",
#         "4cl1",
#         "4cl2",
#         "4cl1x4cl2",
#         "ccoaomt1",
#         "fah1",
#         "omt1",
#         "ccr1-3",
#         "cad4",
#         "cad5",
#         "cad4xcad5"
#       )
#     )
#   )
#
# write_csv(rmn_data_corrected, "rmn_data_corrected.csv")
rmn_data_corrected <- read_csv("rmn_data_corrected.csv", col_types = "ccccnnn") %>%
  mutate(
    genotype = ordered(genotype,
      levels = c(
        "Col-0",
        "4cl1",
        "4cl2",
        "4cl1x4cl2",
        "ccoaomt1",
        "fah1",
        "omt1",
        "ccr1-3",
        "cad4",
        "cad5",
        "cad4xcad5"
      )
    )
  )

rmn_data_corrected <- rmn_data_corrected %>%
  # left_join(lig_peak) %>%
  group_by(genotype, cell.type, replicate, technical) %>%
  mutate(
    AUC = MESS::auc(wavenumber, corrected.intensity),
    scaled = corrected.intensity / AUC,
    scaled_lig = corrected.intensity / (
      (corrected.intensity[wavenumber == 1603] / 2) + corrected.intensity[wavenumber == 1334]
    )
  )

rmn_data_pre <- rmn_data_corrected %>%
  group_by(genotype, cell.type, replicate, wavenumber) %>%
  summarise(
    samples.pre = length(corrected.intensity),
    mean.scaled = mean(scaled, na.rm = TRUE),
    mean.scaled_lig = mean(scaled_lig, na.rm = TRUE),
    # mean.lig_peak = mean(lig_peak, na.rm = TRUE)
  )

rmn_data_avg <- rmn_data_pre %>%
  group_by(genotype, cell.type, wavenumber) %>%
  summarise(
    mean.scaled = mean(mean.scaled, na.rm = TRUE),
    mean.scaled_lig = mean(mean.scaled_lig, na.rm = TRUE),
    # mean.lig_peak = mean(mean.lig_peak, na.rm = TRUE)
  )

#### average Raman data, calculate integrals and detect peaks ####
rmn_sum <- rmn_data_corrected %>%
  group_by(genotype, cell.type, replicate, technical) %>%
  summarise(
    "total_lignin" = (scaled[wavenumber == 1603] / 2) + scaled[wavenumber == 1334],
    "G_rel" = scaled_lig[wavenumber == 1603],
    "GOH_rel" = scaled_lig[wavenumber == 1664],
    "GCHO_rel" = scaled_lig[wavenumber == 1625],
    "S_rel" = scaled_lig[wavenumber == 1334],
    "P_rel" = scaled_lig[wavenumber == 1001],
    "X1174" = scaled_lig[wavenumber == 1174],
    "X1141" = scaled_lig[wavenumber == 1141],
    "cellulose" = scaled[wavenumber == 1099] + scaled[wavenumber == 381],
    "S.G" = scaled_lig[wavenumber == 1334] / scaled_lig[wavenumber == 1276],
    "GOH.G" = scaled_lig[wavenumber == 1664] / scaled_lig[wavenumber == 1603],
    # "GCHO.GOH" = scaled_lig[wavenumber == 1625] /scaled_lig[wavenumber == 1664],
    "AUC" = mean(AUC)
  )
```


### Export scaled spectra

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}

spectra_export.switch <- left_join(read_tsv("/data/PhD/IRX/PMX_which_SMX.tsv",
  col_types = "cccc"
), rmn_data_corrected) %>%
  mutate(
    technical = str_c(technical, "switched", sep = "_"),
    cell.type = "SMX"
  )

spectra_export <- rmn_data_corrected %>%
  anti_join(read_tsv("/data/PhD/IRX/PMX_which_SMX.tsv",
    col_types = "cccc"
  )) %>%
  bind_rows(rmn_irx.switch)

spectra_export_raw <- spectra_export %>%
  select(genotype:corrected.intensity) %>%
  pivot_wider(id_cols = c(genotype:technical), names_from = wavenumber, values_from = corrected.intensity)

spectra_export_scaled <- spectra_export %>%
  select(genotype:wavenumber, scaled) %>%
  pivot_wider(id_cols = c(genotype:technical), names_from = wavenumber, values_from = scaled)

write_excel_csv(spectra_export_raw, "Raman_spectra_raw.csv")
write_excel_csv(spectra_export_scaled, "Raman_spectra_scaled.csv")
```

### Load Wiesner data

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, warning = FALSE, eval = TRUE, echo = FALSE}
#### import and tidy Wiesner test data ####
wiesner.data <-
  read.csv("/home/leonard/Documents/Uni/Phloroglucinol/measurements_revisited.csv",
    skip = 2
  )

wiesner.data$genotype <- recode(wiesner.data$genotype, cad4x5 = "cad4xcad5")

# set cell types according to measurement order
wiesner.data[1:50 + rep(seq(0, (nrow(wiesner.data) - 50), by = 300), each = 50), 4] <-
  "IF"
wiesner.data[51:100 + rep(seq(0, (nrow(wiesner.data) - 50), by = 300), each = 50), 4] <-
  "MX"
wiesner.data[101:150 + rep(seq(0, (nrow(wiesner.data) - 50), by = 300), each = 50), 4] <-
  "XF"
wiesner.data[151:200 + rep(seq(0, (nrow(wiesner.data) - 50), by = 300), each = 50), 4] <-
  "PX"
wiesner.data[201:250 + rep(seq(0, (nrow(wiesner.data) - 50), by = 300), each = 50), 4] <-
  "LP"
wiesner.data[251:300 + rep(seq(0, (nrow(wiesner.data) - 50), by = 300), each = 50), 4] <-
  "PH"

# import SMX measurements
wiesner.data.SMX <- read.csv("/home/leonard/Documents/Uni/Phloroglucinol/measurements_SMX.csv",
  skip = 2
)
wiesner.data.SMX$replicate <- as.factor(wiesner.data.SMX$replicate)

wiesner.data <- rbind(select(wiesner.data, -technical), select(wiesner.data.SMX, -technical))

wiesner.data$genotype <-
  ordered(
    wiesner.data$genotype,
    levels = c(
      "col-0",
      "4cl1",
      "4cl2",
      "4cl1x2",
      "ccoaomt1",
      "fah1",
      "omt1",
      "ccr1-3",
      "ccr1xfah1",
      "cad4",
      "cad5",
      "cad4xcad5"
    )
  )

# calculate the correct hue on the 360 point circular scale
wiesner.data$hue <- ((wiesner.data$h.stained + 128) / 255 * 360)

wiesner.data$replicate <-
  as.factor(as.character(wiesner.data$replicate))

# calculate stained - unstained diff. and adjust for bleaching by subtracting the diff. for the unlignified phloem
wiesner.data$diff <-
  wiesner.data$OD.stained - wiesner.data$OD.unstained
# summary(wiesner.data)

wiesner.data.bg <- wiesner.data %>%
  filter(cell.type == "PH") %>%
  select(1:2, 9) %>%
  ungroup() %>%
  group_by(genotype, replicate) %>%
  summarise(OD.bg = mean(diff, na.rm = TRUE))

wiesner.data <-
  full_join(
    wiesner.data,
    wiesner.data.bg,
    all = TRUE,
    by = c("genotype", "replicate")
  )
wiesner.data$diff.adj <- wiesner.data$diff - wiesner.data$OD.bg
wiesner.data <- subset(wiesner.data, cell.type != "PH")

# average per replicate (for boxplots)
wiesner.data.pre <- wiesner.data %>%
  mutate(
    cell.type = recode(cell.type, "MX" = "PMX"),
    genotype = recode(genotype,
      "col-0" = "Col-0",
      "4cl1x2" = "4cl1x4cl2"
    )
  ) %>%
  group_by(genotype, cell.type, replicate) %>%
  summarise(
    mean.hue1 = mean(hue, na.rm = TRUE),
    SD.hue1 = sd(hue, na.rm = TRUE),
    mean.OD1 = mean(diff.adj, na.rm = TRUE),
    SD.OD1 = sd(diff.adj, na.rm = TRUE)
  ) %>%
  select(genotype, cell.type, replicate, mean.OD1)

#### import and tidy shape data of the vessels from the Raman measurements ####
rmn_irx <- read.csv("file:///data/PhD/IRX/raman_IRX_revisited.csv") %>%
  select(genotype:Perim., Circ., Round, Height, Width) %>%
  filter(object == 2) %>%
  select(-object)

raman.nb <- read.csv("file:///data/PhD/IRX/raman_IRX.csv") %>%
  select(genotype:n_f)
rmn_irx <- left_join(rmn_irx, raman.nb)
rmn_irx$replicate <- as.character(rmn_irx$replicate)
rmn_irx$technical <- as.character(rmn_irx$technical)

raman_convex <- read.csv("/data/PhD/IRX/raman_irx_convex.csv")

raman_convex$File <-
  str_replace(raman_convex$File, fixed("fah 1"), "fah1")

raman_convex <- raman_convex %>%
  select(-X) %>%
  separate(
    File,
    into = c("genotype", "replicate", "cell.type", "technical"),
    sep = "\\s*#|-|\\s",
    extra = "merge"
  ) %>%
  mutate(
    genotype = recode(
      genotype,
      "4cl1＆2" = "4cl1x4cl2",
      "cad4＆5" = "cad4xcad5",
      "ccoaomt" = "ccoaomt1",
      "ccr1" = "ccr1-3",
      "col.o" = "Col-0",
      "Col.0" = "Col-0",
      "ccr1＆fah1" = "ccr1xfah1"
    ),
    cell.type = recode(cell.type,
      "px" = "PX",
      "SX" = "SMX",
      "MX" = "PMX"
    ),
    replicate = str_extract(
      replicate,
      "\\d"
    ),
    technical = str_extract(
      technical,
      "(\\d)+"
    )
  ) %>%
  filter(cell.type != "？？")

rmn_irx <- left_join(rmn_irx, raman_convex)

# write_csv(rmn_irx, "raman_irx_shape.csv")

#### import and tidy plant height data ####
raman.heights <- read.csv("/home/leonard/Documents/Uni/Master/Summer project 16/phenotyping/phenotyping.csv") %>%
  select(genotype, replicate, height) %>%
  mutate(genotype = recode(genotype, "col-0" = "Col-0")) %>%
  rename("Plant.height" = "height") %>%
  rbind(read.csv("file:///data/PhD/IRX/heights_haris.csv"))
raman.heights$replicate <- as.character(raman.heights$replicate)

#### merge data frames ####
rmn_irx <- left_join(rmn_irx, rmn_sum)
rmn_irx.switch <- left_join(read_tsv("/data/PhD/IRX/PMX_which_SMX.tsv",
  col_types = "cccc"
), rmn_irx) %>%
  mutate(
    technical = str_c(technical, "switched", sep = "_"),
    cell.type = "SMX"
  )
rmn_irx <- rmn_irx %>%
  anti_join(read_tsv("/data/PhD/IRX/PMX_which_SMX.tsv",
    col_types = "cccc"
  )) %>%
  bind_rows(rmn_irx.switch)
rmn_irx <- left_join(rmn_irx, raman.heights)
rmn_irx <- left_join(rmn_irx, wiesner.data.pre) %>%
  mutate(
    convexity = Area / ConvexArea,
    "GCHO_term.GCHO_tot" = GCHO_rel / mean.OD1,
    "GCHO.GOH" = mean.OD1 / GOH_rel
  )
# write.csv(rmn_irx, file = "SEM_data.csv")
```

### WT vessel characteristics

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%', fig.width=5, fig.height=4}

rmn_sum_long <- rmn_irx %>%
  filter(cell.type != "IF" & genotype == "Col-0") %>%
  group_by(cell.type, replicate, technical) %>%
  select(-genotype) %>%
  pivot_longer(
    cols = -c(replicate:cell.type),
    names_to = "variable",
    values_to = "value"
  ) %>%
  ungroup() %>%
  mutate(
    cell.type = ordered(cell.type, levels = c("PX", "PMX", "SMX"))
  )

rmn_adj <- rmn_sum_long %>%
  filter(variable %in% c("n_v", "n_f", "n_p")) %>%
  mutate(variable = ordered(as.factor(variable), levels = c("n_p", "n_f", "n_v")))

rmn_adj_letters <- letter_groups(rmn_adj,
  value,
  cell.type,
  "kruskal",
  variable,
  print_position = "below",
  print_adjust = 0.5
) %>%
  mutate(cell.type = ordered(cell.type, levels = c("PX", "PMX", "SMX")))

rmn_adj_box <- ggplot(
  data = rmn_adj,
  aes(x = variable, y = value * 100)
) +
  geom_quasirandom(
    aes(
      group = interaction(cell.type, variable),
      fill = cell.type
    ),
    dodge.width = 0.6,
    shape = 21,
    # width = 0.1,
    # colour = NA,
    alpha = 0.75,
    size = 1,
    stroke = 0.25
  ) +
  geom_violin(aes(
    group = interaction(cell.type, variable),
    # fill = cell.type
  ),
  draw_quantiles = 0.5,
  fill = "white",
  colour = "black",
  alpha = 0.85,
  width = 0.2,
  position = position_dodge(width = 0.6),
  size = 0.2,
  scale = "width"
  ) +
  geom_text(
    data = rmn_adj_letters,
    aes(
      group = interaction(cell.type, variable),
      label = Letters
    ),
    size = 6 / (14 / 5),
    family = "Helvetica",
    position = position_dodge(width = 0.6)
  ) +
  annotate("text",
    x = c(0.8, 1, 1.2),
    y = c(100, 90, 25),
    label = c("PX", "PMX", "SMX"),
    family = "Helvetica",
    size = ggtext_size
  ) +
  annotate("segment",
    x = c(0.8, 1, 1.2),
    xend = c(0.8, 1, 1.2),
    y = c(95, 85, 20),
    yend = c(85, 75, 5),
    colour = "black",
    size = 0.2
  ) +
  scale_x_discrete(labels = c("Parenchyma", "Fibres", "Vessels")) +
  scale_fill_manual(values = pal_flame_disc) +
  labs(
    x = "Neighbouring cell type",
    y = "CW adjacent [%]"
  ) +
  theme_leo() +
  theme(
    text = element_text(family = "Helvetica"),
    legend.position = "none"
  )

pdf("rmn_adj_box.pdf", width = 9 / 2.54, height = 1.5)
rmn_adj_box
dev.off()


rmn_violin <- function(x) {
  letters <- letter_groups(rmn_sum_long %>% filter(variable == x),
    value,
    cell.type,
    "tukey",
    print_position = "below",
    print_adjust = 0.5
  ) %>%
    mutate(cell.type = ordered(cell.type, levels = c("PX", "PMX", "SMX")))

  ggplot(
    data = rmn_sum_long %>% filter(variable == x),
    aes(x = cell.type, y = value)
  ) +
    geom_quasirandom(
      aes(fill = cell.type),
      dodge.width = 0.6,
      shape = 21,
      # width = 0.1,
      # colour = NA,
      alpha = 0.75,
      size = 1,
      stroke = 0.2
    ) +
    geom_violin(aes(
      group = cell.type,
      # fill = cell.type
    ),
    draw_quantiles = 0.5,
    fill = "white",
    colour = "black",
    alpha = 0.85,
    width = 0.2,
    position = position_dodge(width = 0.6),
    size = 0.2,
    scale = "width"
    ) +
    geom_text(
      data = letters,
      aes(label = Letters),
      size = ggtext_size,
      family = "Helvetica",
      position = position_dodge(width = 0.6)
    ) +
    scale_x_discrete(labels = c("PX", "PMX", "SMX")) +
    scale_fill_manual(values = pal_flame_disc) +
    expand_limits(y = 0) +
    theme_leo() +
    theme(
      text = element_text(family = "Helvetica"),
      legend.position = "none",
    )
}

rmn_area <- rmn_violin("Area") +
  labs(
    x = "",
    y = expression(paste("Area [µm"^2, "]"))
  )

rmn_lig <- rmn_violin("total_lignin") +
  labs(
    x = "",
    y = expression(paste("Relative lignin"))
  )

rmn_s <- rmn_violin("S_rel") +
  labs(
    x = "",
    y = expression(paste("Relative " * bold(S)))
  )

rmn_g <- rmn_violin("G_rel") +
  labs(
    x = "",
    y = expression(paste("Relative " * bold(G)))
  )

rmn_goh <- rmn_violin("GOH.G") +
  labs(
    x = "",
    y = expression(paste("Relative " * bold(G)[CHOH]))
  )

rmn_cellu <- rmn_violin("cellulose") +
  labs(
    x = "",
    y = expression(paste("Relative cellulose"))
  )

rmn_gcho.goh <- rmn_violin("GCHO.GOH") +
  labs(
    x = "",
    y = expression(paste(bold(G)[CHO] ~ "/" ~ bold(G)[CHOH]))
  )

rmn_s.g <- rmn_violin("S.G") +
  labs(
    x = "",
    y = expression(paste(Relative ~ bold(S) ~ "/" ~ bold(G)))
  )

rmn_gcho_term.gcho_tot <- rmn_violin("GCHO_term.GCHO_tot") +
  labs(
    x = "",
    y = expression(paste(Terminal ~ bold(G)[CHO] ~ "/" ~ bold(G)[CHO]))
  )

celltypes_WT <-
  rasterGrob(image_read_pdf(
    "/home/leonard/Dropbox/IRX manuscript/Submission/Edits/IRX_WT.pdf",
    density = 900
  ))

```

### Assemble figure

```{r figure 2, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%', fig.width=5, fig.height=4}


top <- plot_grid(celltypes_WT,
  rmn_adj_box,
  labels = c("A", "B"),
  nrow = 1,
  rel_widths = c(1, 2.2),
  vjust = 1,
  label_size = 10,
  label_fontfamily = "Helvetica"
)
bottom <- plot_grid(
  rmn_area,
  rmn_cellu,
  rmn_lig,
  rmn_s.g,
  rmn_gcho.goh,
  rmn_gcho_term.gcho_tot,
  labels = c("C", "D", "E", "F", "G", "H"),
  vjust = 1,
  nrow = 2,
  label_size = 10,
  label_fontfamily = "Helvetica",
  align = "hv",
  axis = "lr"
)

pdf("IRX_figure3.pdf", width = onecol, height = onecol)
plot_grid(top,
  bottom,
  nrow = 2,
  rel_heights = c(1.3, 2)
)
dev.off()
```
## Model figure

### Visual vessel overview

```{r figure 3, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}
irx_data <-
  read.csv("file:///data/PhD/IRX/irx_measurements.csv", sep = "\t")
# irx_data <- subset(irx_data, replicate == 1 & technical == 3)
irx_ref <- subset(irx_data, object == "ref", select = c(1, 2, 3, 6, 7, 27))
irx_ref$ref.y1 <- irx_ref$Y
irx_ref$ref.y2 <- irx_ref$Y
irx_ref$ref.x1 <- irx_ref$X - (irx_ref$Length / 2)
irx_ref$ref.x2 <- irx_ref$X + (irx_ref$Length / 2)


irx_data <- irx_data %>%
  filter(object != "ref") %>%
  left_join(irx_ref[, c(1, 2, 3, 7, 8, 9, 10)],
    by = c("genotype", "replicate", "technical")
  )

irx_data$Distance <-
  apply(
    irx_data[, c("X", "Y", "ref.x1", "ref.x2", "ref.y1", "ref.y2")],
    1,
    function(x) {
      a <- c(x[1], x[2])
      b <- c(x[3], x[5])
      c <- c(x[4], x[6])
      v1 <- b - c
      v2 <- a - b
      m <- cbind(v1, v2)
      d <- abs(det(m)) / sqrt(sum(v1 * v1))
      d
    }
  )

irx_data <- irx_data %>%
  rownames_to_column(var = "number")

irx_x <- irx_data %>%
  select(genotype, replicate, technical, object, number, contains("ROIx")) %>%
  group_by(genotype, replicate, technical, object, number) %>%
  pivot_longer(contains("ROIx"), names_to = "variable", values_to = "ROIx") %>%
  separate(variable, into = c("variable", "point"), sep = "x") %>%
  select(-variable)

irx_y <- irx_data %>%
  select(genotype, replicate, technical, object, number, contains("ROIy")) %>%
  group_by(genotype, replicate, technical, object, number) %>%
  pivot_longer(contains("ROIy"), names_to = "variable", values_to = "ROIy") %>%
  separate(variable, into = c("variable", "point"), sep = "y") %>%
  select(-variable)

irx_roi <- left_join(irx_x, irx_y) %>%
  drop_na() %>%
  filter_at(vars("ROIx", "ROIy"), any_vars(. != 0))

irx_str <- irx_roi %>%
  unite("ID", genotype:number, remove = FALSE, sep = ",") %>%
  group_by(ID) %>%
  mutate(
    number = as.numeric(number),
    ROIx = ROIx - mean(ROIx, na.rm = TRUE),
    ROIy = ROIy - mean(ROIy, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  group_by(ID, point) %>%
  summarise(xy = list(c(ROIx, ROIy))) %>%
  ungroup() %>%
  group_by(ID) %>%
  mutate(
    point = as.numeric(point),
    max_point = max(point)
  ) %>%
  bind_rows(filter(., point == 0) %>%
    group_by(ID) %>%
    mutate(point = max_point + 1)) %>%
  arrange(ID, point) %>%
  summarise(xy = list(c(xy))) %>%
  mutate(geometry = map(
    xy,
    ~ do.call(rbind, .) %>% # make each list a matrix
      list() %>% # st_polygon() requires a list
      st_polygon()
  )) %>%
  st_as_sf()

irx_str <- irx_str %>%
  group_by(ID) %>%
  mutate(
    area = st_area(geometry),
    CH = st_convex_hull(geometry),
    convex_area = st_area(CH),
    convexity = area / convex_area,
    circularity = 4 * pi * (area / (lwgeom::st_perimeter(geometry)^2))
  ) %>%
  separate(ID, into = c("genotype", "replicate", "technical", "cell.type", "number"), sep = ",") %>%
  filter(genotype != "ccr1xfah1") %>%
  mutate(
    genotype = recode(genotype, "col-0" = "Col-0"),
    genotype = ordered(genotype, levels = c(
      "Col-0",
      "4cl1",
      "4cl2",
      "4cl1x4cl2",
      "ccoaomt1",
      "fah1",
      "omt1",
      "ccr1-3",
      "cad4",
      "cad5",
      "cad4xcad5"
    )),
    cell.type = ordered(cell.type, levels = c("PX", "PMX", "SMX"))
  )

sf_ROIs <- ggplot() +
  geom_sf(
    data = irx_str,
    aes(colour = circularity),
    # colour = "black",
    fill = NA,
    lwd = 0.1,
    key_glyph = "point"
  ) +
  theme_leo() +
  scale_x_continuous(expand = expand_scale(mult = 0)) +
  scale_y_continuous(expand = expand_scale(mult = 0)) +
  theme(
    text = element_text(family = "Helvetica"),
    legend.position = "right",
    panel.border = element_blank(),
    # panel.background = element_rect(fill = "black"),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    strip.text = element_text(hjust = 0.5, face = "italic")
  ) +
  scale_colour_gradientn(colours = pal_ostwald_cont, trans = "reverse", name = "Circularity") +
  guides(colour = guide_colourbar(barwidth = unit(2, "mm"))) +
  facet_grid(cell.type ~ genotype, switch = "y")

pdf("sf_ROIs.pdf", width = twocol * 0.7, height = onecol)
sf_ROIs
dev.off()


xdens <- ggplot(irx_str) +
  geom_density(aes(x = circularity, fill = cell.type),
    size = 0.1,
    alpha = 0.5
  ) +
  scale_fill_manual(values = pal_flame_disc) +
  theme_void()

ydens <- ggplot(irx_str) +
  geom_density(aes(x = convexity, fill = cell.type),
    size = 0.1,
    alpha = 0.5
  ) +
  scale_fill_manual(values = pal_flame_disc) +
  coord_flip() +
  theme_void()


pdf("shape_distribution.pdf", height = onecol / 2, width = onecol)
scatter <- ggplot(irx_str) +
  geom_point(aes(x = circularity, y = convexity, shape = cell.type, fill = cell.type),
    size = 1,
    alpha = 0.75,
    stroke = 0.2
  ) +
  scale_fill_manual(values = pal_flame_disc) +
  scale_shape_manual(values = c(21, 22, 23)) +
  labs(
    x = "Circularity",
    y = "Convexity"
  ) +
  theme_leo() +
  theme(
    legend.position = c(0.9, 0.2),
    legend.title = element_blank()
  )
scatter_x <- insert_xaxis_grob(scatter, xdens,
  height = grid::unit(0.1, "null"),
  position = c("top")
)
scatter_xy <- insert_yaxis_grob(scatter_x, ydens,
  width = grid::unit(0.1, "null"),
  position = c("right")
)
ggdraw(scatter_xy)
dev.off()
```

### Model otimisation

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}
step_data <- rmn_irx %>%
  filter(cell.type == "PX") %>%
  select(convexity,
         total_lignin,
         P_rel,
         # G_rel,
         # GOH.G,
         S_rel,
         mean.OD1,
         Perim.,
         n_v,
         GCHO_rel,
         GOH_rel,
         # X1141,
         # X1174
         ) %>%
  drop_na()

# Fit the full model 
full_model <- lm(convexity ~., data = step_data)
# Stepwise regression model
step_model <- MASS::stepAIC(full_model, direction = "both", 
                      trace = FALSE)
summary(step_model)


```

#### Multigroup model (discarded)

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}
TM.multigroup <- function(x) {
  psem(
    lm(Circ. ~
    n_v +
      convexity,
    data = filter(drop_na(rmn_irx), cell.type == x)
    ),
    lm(convexity ~
    # n_v +
    # n_f +
    # mean.OD1 +
    Perim. +
      # S.G +
      # G_tot +
      # GOH_rel +
      GOH.G +
      # G_rel +
      # GOH_tot +
      S_rel +
      # S_tot +
      # GCHO_rel +
      total_lignin,
    data = filter(drop_na(rmn_irx), cell.type == x)
    ),
    # lme(Perim. ~
    #       # n_v,
    #       # n_f +
    #       mean.OD1 +
    #       # cellu.peak +
    #       # lig.peak.pos +
    #       lig.peak,
    #     random = ~1|replicate,
    #     data = filter(drop_na(rmn_irx), cell.type == x)
    # ),
    # lme(HeightByWidth ~
    #       # mean.OD1 +
    #       # lig.peak +
    #       # n_v +
    #       Circ.,
    #     random = ~1|replicate,
    #     data = filter(drop_na(rmn_irx), cell.type == x)
    # ),
    # lme(X1147.1603 ~
    #       Perim.,
    #     random = ~1|replicate,
    #     data = filter(drop_na(rmn_irx), cell.type == x)
    # ),
    # lme(X1603.1099 ~
    #       Perim.,
    #     random = ~1|replicate,
    #     data = filter(drop_na(rmn_irx), cell.type == x)
    # ),
    # lme(mean.OD1 ~
    #       X1664.1603 +
    #       X1603.1099 +
    #       X1621.1603,
    #     random = ~1|replicate,
    #     data = filter(drop_na(rmn_irx), cell.type == x)
    # ),
    data = filter(drop_na(rmn_irx), cell.type == x)
  )
}
```

### SEM for PX

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}

psem_px <- psem(
  lm(Circ. ~
  # n_v +
  convexity,
  data = filter(drop_na(rmn_irx), cell.type == "PX")
  ),
  lm(convexity ~
  # n_v +
  # n_f +
  mean.OD1 +
    Perim. +
    P_rel +
    # cellulose +
    # S.G +
    # G_tot +
    # GOH_rel +
    GOH.G +
    # G_rel +
    # GOH_tot +
    S_rel +
    # S_tot +
    # GCHO_rel +
    # total_lignin:mean.OD1 +
    total_lignin,
  data = filter(drop_na(rmn_irx), cell.type == "PX")
  ),
  data = filter(drop_na(rmn_irx), cell.type == "PX")
)
ind_claims <- summary(psem_px)[["dTable"]][, c(1, 4, 5)] %>%
  mutate(
    model = "PX",
    species = "Arabidopsis",
    fishC = summary(psem_px)[["Cstat"]][["Fisher.C"]],
    fishP = summary(psem_px)[["Cstat"]][["P.Value"]]
  )

coefs <- summary(psem_px)[["coefficients"]][, c(1:8)] %>%
  mutate(
    model = "PX",
    species = "Arabidopsis",
  )

plot(psem_px)
```

### SEM for PMX

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}

psem_pmx <- psem(
  lm(Circ. ~
  n_v +
    convexity,
  data = filter(drop_na(rmn_irx), cell.type == "PMX")
  ),
  lm(convexity ~
  n_v +
    # n_f +
    mean.OD1 +
    P_rel +
    # Perim. +
    # S.G +
    # G_tot +
    # GOH_rel +
    # GOH.G +
    # G_rel +
    # GOH_tot +
    S_rel +
    # S_tot +
    GCHO_rel +
    # total_lignin:mean.OD1 +
    total_lignin,
  data = filter(drop_na(rmn_irx), cell.type == "PMX")
  ),
  data = filter(drop_na(rmn_irx), cell.type == "PMX")
)
ind_claims <- ind_claims %>%
  bind_rows(summary(psem_pmx)[["dTable"]][, c(1, 4, 5)] %>%
    mutate(
      model = "PMX",
      species = "Arabidopsis",
      fishC = summary(psem_pmx)[["Cstat"]][["Fisher.C"]],
      fishP = summary(psem_pmx)[["Cstat"]][["P.Value"]]
    ))

coefs <- coefs %>%
  bind_rows(summary(psem_pmx)[["coefficients"]][, c(1:8)] %>%
    mutate(
      model = "PMX",
      species = "Arabidopsis"
    ))
plot(psem_pmx)
summary(psem_pmx)
```

### SEM for SMX

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}

psem_smx <- psem(
  lm(Circ. ~
  # n_v +
  convexity,
  data = filter(drop_na(rmn_irx), cell.type == "SMX")
  ),
  lm(convexity ~
  # n_v +
  # n_f +
  # mean.OD1 +
  # Perim. +
  # S.G +
  # G_tot +
  # GOH_rel +
  GOH.G +
    P_rel +
    # G_rel +
    # GOH_tot +
    # S_rel +
    # S_tot +
    # GCHO_rel +
    # X1174 +
    # total_lignin:mean.OD1 +
    total_lignin,
  data = filter(drop_na(rmn_irx), cell.type == "SMX")
  ),
  data = filter(drop_na(rmn_irx), cell.type == "SMX")
)
ind_claims <- ind_claims %>%
  bind_rows(summary(psem_smx)[["dTable"]][, c(1, 4, 5)] %>%
    mutate(
      model = "SMX",
      species = "Arabidopsis",
      fishC = summary(psem_smx)[["Cstat"]][["Fisher.C"]],
      fishP = summary(psem_smx)[["Cstat"]][["P.Value"]]
    ))

coefs <- coefs %>%
  bind_rows(summary(psem_smx)[["coefficients"]][, c(1:8)] %>%
    mutate(
      model = "SMX",
      species = "Arabidopsis",
    ))
plot(psem_smx)
summary(psem_smx)
```

## Plant bending

### Hydration influence

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}
medium_data <- read_csv("/data/PhD/IRX/Bending/WT_medium.csv") %>%
  pivot_longer(cols = c(stiffness, strength), values_to = "value", names_to = "variable") %>%
  mutate(medium = ordered(medium, levels = c("Water", "Air", "Sorbitol")))

bending_files <-
  list.files(
    path = "/data/PhD/IRX/Bending/media_raw/",
    pattern = "*.txt",
    recursive = TRUE,
    full.names = TRUE
  )

read_plus <- function(flnm) {
  read_delim(flnm,
    delim = ";",
    skip = grep("Raw Data", readLines(flnm)),
    col_types = "cccccc",
    col_names = FALSE
  ) %>%
    fill(X1) %>%
    filter(X2 != "Time" & X2 != "(s)") %>%
    mutate(
      filename = basename(flnm)
    ) %>%
    rename(
      "section" = X1,
      "time" = X2,
      "displacement" = X3,
      "force" = X4,
      "stress" = X5,
      "strain" = X6
    )
}
medium_curves <- lapply(bending_files, read_plus) %>%
  bind_rows() %>%
  mutate(
    filename = str_remove(filename, fixed("Stem ")),
    filename = str_remove(filename, fixed(".txt"))
  ) %>%
  separate(filename, into = c("replicate", "medium"), sep = " ") %>%
  mutate(
    replicate = as.factor(replicate),
    medium = as.factor(medium)
  ) %>%
  mutate_if(is.character, ~ as.numeric(str_replace(., fixed(","), fixed(".")))) %>%
  mutate(medium = ordered(medium, levels = c("Water", "Air", "Sorbitol")))

medium_breaking <- medium_curves %>%
  group_by(medium, replicate, section) %>%
  filter(strain < 2.5) %>%
  summarise(
    break_point = max(strain[stress == max(stress)]),
    strength = max(stress)
  ) %>%
  pivot_longer(cols = c(break_point, strength), names_to = "variable", values_to = "value")

flex_violin <- function(x, dat = medium_data) {
  letters <- letter_groups(dat %>% filter(variable == x),
    value,
    medium,
    "tukey",
    print_position = "below",
    print_adjust = 0.5
  )

  ggplot(
    data = dat %>% filter(variable == x),
    aes(x = medium, y = value)
  ) +
    geom_quasirandom(
      aes(fill = medium),
      # dodge.width = 0.6,
      shape = 21,
      # width = 0.1,
      # colour = NA,
      alpha = 0.75,
      size = 1,
      stroke = 0.2
    ) +
    geom_violin(
      draw_quantiles = 0.5,
      fill = "white",
      colour = "black",
      alpha = 0.85,
      width = 0.2,
      position = position_dodge(width = 0.6),
      size = 0.2,
      scale = "width"
    ) +
    geom_text(
      data = letters,
      aes(label = Letters),
      size = ggtext_size,
      family = "Helvetica",
      position = position_dodge(width = 0.6)
    ) +
    scale_fill_grey(start = 0.9, end = 0.1) +
    expand_limits(y = 0) +
    scale_y_continuous(expand = expand_scale(mult = c(0.1, 0.05))) +
    theme_leo() +
    theme(
      text = element_text(family = "Helvetica"),
      legend.position = "none"
    )
}

medium_strength <- flex_violin("strength") +
  labs(y = "Strength [MPa]") +
  theme(axis.title.x = element_blank())

medium_stiffness <- flex_violin("stiffness") +
  labs(y = "Stiffness [MPa]") +
  theme(axis.title.y = element_blank()) +
  scale_x_discrete(
    limits = rev(levels(medium_data$medium))
  ) +
  coord_flip()

medium_breakpoint <- flex_violin("break_point", medium_breaking) +
  labs(y = "Strain [%] at breaking") +
  coord_flip() +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  scale_x_discrete(
    limits = rev(levels(medium_data$medium))
  )

pdf("bending_medium.pdf", width = onecol / 1.25, height = onecol / 3)
hydration <- plot_grid(
  medium_strength,
  medium_stiffness,
  labels = c("B", "C"),
  # vjust = 1,
  nrow = 1,
  rel_widths = c(1, 1.1),
  label_size = 10,
  label_fontfamily = "Helvetica"
)
hydration
dev.off()
```

### Mutant bending

#### Load data

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}
#### fah1 ####
bending_files <-
  list.files(
    path = "/data/PhD/IRX/Bending/fah1_raw/",
    pattern = "*.txt",
    recursive = TRUE,
    full.names = TRUE
  )

read_plus <- function(flnm) {
  read_delim(flnm,
    delim = ";",
    skip = grep("Raw Data", readLines(flnm)),
    col_types = "cccccc",
    col_names = FALSE
  ) %>%
    fill(X1) %>%
    filter(X2 != "Time" & X2 != "(s)") %>%
    mutate(
      filename = basename(flnm)
    ) %>%
    rename(
      "section" = X1,
      "time" = X2,
      "displacement" = X3,
      "force" = X4,
      "strain" = X5,
      "stress" = X6
    )
}

fahA <- read_delim(bending_files[1],
  delim = ";",
  skip = grep("Raw Data", readLines(bending_files[1])),
  col_types = "cccccc",
  col_names = FALSE
) %>%
  mutate(filename = basename(bending_files[1])) %>%
  fill(X1) %>%
  filter(X2 != "Displacement" & X2 != "(mm)") %>%
  rename(
    "section" = X1,
    "displacement" = X2,
    "force" = X3,
    "strain" = X4,
    "stress" = X5
  )

bending_data <- lapply(bending_files[-1], read_plus) %>%
  bind_rows()

bending_data <- bending_data %>%
  bind_rows(fahA) %>%
  separate(filename, into = c("genotype", "replicate"), sep = " ", extra = "merge") %>%
  mutate(
    replicate = str_remove(replicate, "Air"),
    replicate = str_remove(replicate, ".txt"),
    replicate = str_remove(replicate, "\\d\\dcm"),
    replicate = str_remove(replicate, " "),
    replicate = as.factor(replicate),
    section = as.factor(section),
    experiment = factor("fah1", levels = c("fah1", "cad")),
    genotype = ordered(recode(genotype, "Fah1" = "fah1"), levels = c("WT", "fah1", "cad4xcad5"))
  ) %>%
  mutate_if(is.character, ~ as.numeric(str_replace(., fixed(","), fixed("."))))

read_stiffness <- function(flnm) {
  read_delim(flnm,
    delim = ";",
    skip = 1,
    col_names = FALSE
  ) %>%
    select(-X1) %>%
    filter(str_detect(X2, "Fah1|WT")) %>%
    rename(
      "name" = X2,
      "modulus" = X3,
      "max.stress" = X4,
      "AUC" = X5,
      "N.max.stress" = X6
    ) %>%
    mutate(
      name = str_remove(name, " Air"),
      experiment = factor("fah1", levels = c("fah1", "cad"))
    ) %>%
    separate(name, into = c("genotype", "replicate", "sample"), extra = "merge") %>%
    mutate(
      genotype = ordered(recode(genotype, "Fah1" = "fah1"), levels = c("WT", "fah1", "cad4xcad5")),
      replicate = as.factor(replicate),
      sample = as.factor(sample)
    ) %>%
    mutate_if(is.character, ~ as.numeric(str_replace(., fixed(","), fixed("."))))
}

stiffness_fah1 <- lapply(bending_files, read_stiffness) %>%
  bind_rows()

#### cad4xcad5 ####
bending_files <-
  list.files(
    path = "/data/PhD/IRX/Bending/cad_raw/",
    pattern = "*.txt",
    recursive = TRUE,
    full.names = TRUE
  )

read_plus <- function(flnm) {
  read_delim(flnm,
    delim = ";",
    skip = grep("Raw Data", readLines(flnm)),
    col_types = "ccccc",
    col_names = FALSE
  ) %>%
    # fill(X1) %>%
    filter(X1 != "Time" & X1 != "(s)") %>%
    mutate(
      filename = basename(flnm),
      filename = str_remove(filename, ".txt"),
      genotype = "cad4xcad5"
    ) %>%
    rename(
      "time" = X1,
      "displacement" = X2,
      "force" = X3,
      "strain" = X4,
      "stress" = X5
    ) %>%
    separate(filename, into = c("section", "replicate"), sep = "_")
}

bending_data_cad <- lapply(bending_files, read_plus) %>%
  bind_rows() %>%
  mutate(
    replicate = as.factor(replicate),
    section = as.factor(section),
    experiment = factor("cad", levels = c("fah1", "cad")),
    genotype = ordered(genotype, levels = c("WT", "fah1", "cad4xcad5"))
  ) %>%
  mutate_if(is.character, ~ as.numeric(str_replace(., fixed(","), fixed("."))))

read_stiffness <- function(flnm) {
  read_delim(flnm,
    delim = ";",
    skip = 1,
    col_names = FALSE
  ) %>%
    select(-X1) %>%
    filter(str_detect(X2, "top|mid|Base")) %>%
    rename(
      "name" = X2,
      "modulus" = X3,
      "max.stress" = X4
    ) %>%
    mutate(
      experiment = factor("cad", levels = c("fah1", "cad")),
      genotype = ordered("cad4xcad5", levels = c("WT", "fah1", "cad4xcad5"))
    ) %>%
    separate(name, into = c("replicate", "sample"), extra = "merge") %>%
    mutate(
      replicate = as.factor(replicate),
      sample = as.factor(sample)
    ) %>%
    mutate_if(is.character, ~ as.numeric(str_replace(., fixed(","), fixed("."))))
}

bending_files_stiff <- bending_files %>%
  str_subset(pattern = fixed("_1."))

stiffness_cad <- lapply(bending_files_stiff, read_stiffness) %>%
  bind_rows()

bending_data_complete <- bind_rows(bending_data, bending_data_cad)

bending_pre <- bending_data_complete %>%
  group_by(genotype, replicate, section) %>%
  summarise(
    break_point = strain[stress == max(stress)],
    strength = max(stress)
  ) %>%
  pivot_longer(cols = c(break_point, strength), names_to = "variable", values_to = "value")

stiffness_complete <- stiffness_fah1 %>%
  bind_rows(stiffness_cad) %>%
  select(-AUC, -N.max.stress) %>%
  pivot_longer(cols = c(modulus, max.stress), names_to = "variable", values_to = "value")
```

#### Figures

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}

flex_violin <- function(x, dat = stiffness_complete) {
  letters <- letter_groups(dat %>% filter(variable == x),
    value,
    genotype,
    "tukey",
    print_position = "below",
    print_adjust = 0.5
  )

  ggplot(
    data = dat %>% filter(variable == x),
    aes(x = genotype, y = value)
  ) +
    # geom_line(aes(group = interaction(cell.type, variable, cell, measurement)),
    #           size = 0.2,
    #           alpha = 0.25) +
    geom_quasirandom(
      aes(fill = genotype),
      dodge.width = 0.6,
      shape = 21,
      # width = 0.1,
      # colour = NA,
      alpha = 0.75,
      size = 1,
      stroke = 0.2
    ) +
    # geom_boxplot(
    #   aes(group = interaction(cell.type, variable), fill = cell.type),
    #   # fill = "white",
    #   alpha = 0.5,
    #   width = 0.6,
    #   position = position_dodge2(width = 1),
    #   outlier.alpha = 0,
    #   lwd = 0.25,
    #   fatten = 1,
    #   width = 0.25
    # ) +
    geom_violin(
      # aes(group = interaction(days, cell.type),
      # fill = cell.type
      # ),
      draw_quantiles = 0.5,
      fill = "white",
      colour = "black",
      alpha = 0.85,
      width = 0.2,
      position = position_dodge(width = 0.6),
      size = 0.2,
      scale = "width"
    ) +
    geom_text(
      data = letters,
      aes(label = Letters),
      size = ggtext_size,
      family = "Helvetica",
      position = position_dodge(width = 0.6)
    ) +
    scale_x_discrete(
      labels = rev(c(
        "WT",
        expression(italic("fah1")),
        expression(paste(italic("cad4"), "x", italic("5")))
      )),
      limits = rev(levels(stiffness_complete$genotype))
    ) +
    scale_fill_manual(values = c(
      pal_ostwald_disc[1],
      pal_ostwald_disc[3],
      pal_ostwald_disc[3]
    )) +
    # scale_fill_gradientn(colours = pal_ostwald_cont, trans = "reverse")
    expand_limits(y = 0) +
    scale_y_continuous(expand = expand_scale(mult = c(0.1, 0.05))) +
    theme_leo() +
    theme(
      text = element_text(family = "Helvetica"),
      legend.position = "none",
      # panel.border = element_blank(),
      # axis.line.x = element_line(size = 0.2),
      # axis.line.y = element_line(size = 0.2)
    )
}


violin_strength <- flex_violin("max.stress") +
  labs(y = "Strength [MPa]") +
  theme(
    # axis.text.y = element_blank(),
    # axis.ticks.y = element_blank(),
    # axis.line.x = element_line(size = 0.2)
  ) + theme(axis.title.x = element_blank()) +
  scale_x_discrete(
    labels = c(
      "WT",
      expression(italic("fah1")),
      expression(paste(italic("cad4"), "x", italic("5")))
    ),
    limits = levels(stiffness_complete$genotype)
  )

violin_stiffness <- flex_violin("modulus") +
  labs(y = "Stiffness [MPa]") +
  scale_y_continuous(
    expand = expand_scale(mult = c(0.1, 0.05)),
    breaks = c(0, 3000, 6000)
  ) +
  theme(axis.title.y = element_blank()) +
  coord_flip()

violin_breakpoint <- flex_violin("break_point", bending_pre) +
  labs(y = "Strain [%] at breaking") +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    # axis.line.x = element_line(size = 0.2)
  ) +
  theme(axis.title.y = element_blank()) +
  coord_flip()

pdf("breaking_point.pdf", width = 2, height = 2)
violin_breakpoint
dev.off()

```

### Mutant lignin composition

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}

heights <- read_csv("/data/Master/Summer project 16/phenotyping/phenotyping.csv") %>%
  select(genotype, height) %>%
  filter(genotype %in% c("col-0", "fah1", "cad4xcad5")) %>%
  mutate(genotype = recode(genotype, "col-0" = "WT")) %>%
  group_by(genotype) %>%
  summarise(
    mean.height = mean(height),
    sd.height = sd(height)
  )

biochem_data <- read_csv("/data/PhD/IRX/Bending/bending_biochem.csv") %>%
  group_by(genotype) %>%
  summarise(
    mean.lig.py = mean(lignin.py),
    sd.lig.py = sd(lignin.py),
    mean.lig.tga = mean(lignin.tga),
    sd.lig.tga = sd(lignin.tga),
    mean.sg = mean(s.g),
    sd.sg = sd(s.g),
    mean.cho.choh = mean(gcho.gchoh),
    sd.cho.choh = sd(gcho.gchoh)
  ) %>%
  ungroup() %>%
  mutate(
    scaled.lig.py = scale_range(mean.lig.py),
    scaled.lig.tga = scale_range(mean.lig.tga),
    scaled.sg = scale_range(mean.sg),
    scaled.cho.choh = scale_range(mean.cho.choh)
  ) %>%
  left_join(heights)

bending_avg <- stiffness_complete %>%
  group_by(genotype, variable) %>%
  summarise(mean = mean(value)) %>%
  pivot_wider(
    id_cols = genotype,
    names_from = variable,
    values_from = mean
  ) %>%
  left_join(biochem_data)

biochem <- ggplot(
  biochem_data,
  aes(x = genotype)
) +
  annotate("rect",
    ymin = c(0.5, 2.5),
    ymax = c(1.5, 3.5),
    xmin = -Inf,
    xmax = Inf,
    fill = "grey95",
    colour = NA
  ) +
  geom_point(aes(
    size = scaled.lig.tga,
    y = 1,
    colour = genotype,
    # alpha = mean.lig.tga / 10
  ),
  alpha = 0.75,
  # size = 8,
  shape = 16
  ) +
  # geom_segment(aes(y = 0.75,
  #                  yend = 0.75 + (scaled.lig.tga * 0.5),
  #                  x = c(0.75, 1.75, 2.75),
  #                  xend = c(0.75, 1.75, 2.75),
  #              colour = genotype)) +
  geom_point(aes(
    size = scaled.sg,
    y = 2,
    colour = genotype,
    # alpha = mean.sg
  ),
  alpha = 0.75,
  # size = 8,
  shape = 16
  ) +
  geom_point(aes(
    size = scaled.cho.choh,
    y = 3,
    colour = genotype,
    # alpha = mean.cho.choh / 10
  ),
  alpha = 0.75,
  # size = 8,
  shape = 16
  ) +
  geom_text(aes(
    label = paste0(
      round(mean.lig.tga, digits = 2), "\n±",
      round(sd.lig.tga, digits = 2)
    ),
    y = 1
  ),
  family = "Helvetica",
  size = ggtext_size,
  lineheight = 0.75,
  hjust = 0.5
  ) +
  geom_text(aes(
    label = paste0(
      round(mean.sg, digits = 2), "\n±",
      round(sd.sg, digits = 2)
    ),
    y = 2
  ),
  family = "Helvetica",
  size = ggtext_size,
  lineheight = 0.75,
  hjust = 0.5
  ) +
  geom_text(aes(
    label = paste0(
      round(mean.cho.choh, digits = 2), "\n±",
      round(sd.cho.choh, digits = 2)
    ),
    y = 3
  ),
  family = "Helvetica",
  size = ggtext_size,
  lineheight = 0.75,
  hjust = 0.5
  ) +
  theme_leo() +
  scale_size_area(
    max_size = 8
  ) +
  theme(
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    axis.ticks.y = element_blank(),
    # panel.border = element_blank(),
    # axis.line.x = element_line(size = 0.2)
  ) +
  scale_y_continuous(
    limits = c(0.5, 3.5),
    breaks = c(1, 2, 3),
    labels = c(
      "Total\nlignin",
      expression(
        bold("S") * "/" *
          bold("G")
      ),
      expression(
        bold("G")[CHO] * "/" *
          bold("G")[CHOH]
      )
    )
  ) +
  scale_colour_manual(values = rev(c(
    pal_ostwald_disc[1],
    pal_ostwald_disc[3],
    pal_ostwald_disc[3]
  ))) +
  scale_x_discrete(limits = rev(levels(stiffness_complete$genotype))) +
  coord_flip()
```

### Assemble bending figure

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}
#### Load images ####
WT <-
  rasterGrob(
    readPNG(
      "/data/PhD/IRX/Bending/Images/bending_adjusted.png"
    )
  )

#### plot grid ####

top <- plot_grid(WT,
  medium_stiffness,
  medium_breakpoint,
  ncol = 3,
  labels = c("", "B", "C"),
  label_fontfamily = "Helvetica",
  label_size = 10,
  rel_widths = c(0.785, 1, 0.785)
  # align = "hv"
)
bottom <- plot_grid(
  violin_stiffness,
  violin_breakpoint,
  biochem,
  ncol = 3,
  labels = c("D", "E", "F"),
  label_fontfamily = "Helvetica",
  label_size = 10,
  # vjust = 1,
  # hjust = 0,
  rel_widths = c(1, 0.785, 0.785),
  align = "h"
)

patch <-
  medium_stiffness + medium_breakpoint + plot_spacer() +
  violin_stiffness + violin_breakpoint + biochem

pdf("IRX_figure5.pdf", width = onecol, height = onecol * 0.85)
patch +
  plot_layout(ncol = 3) &
  plot_annotation(
    tag_levels = "A",
    theme = theme(plot.margin = margin(-5, -5, -5, -5))
  ) &
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))

dev.off()
```

## Drought experiment

### Rosette area

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}
rosettes <- read_csv("/data/PhD/IRX/Drought/drought_area.csv") %>%
  mutate(
    image = str_remove(image, fixed(".JPG")),
    mean_hue = (mean_hue / 255) * 360
  ) %>%
  separate(image, c("genotype", "medium", "replicate"), sep = "-") %>%
  mutate(
    medium = ordered(recode(medium,
      "0%" = "Water",
      "10%" = "10% PEG",
      "20%" = "20% PEG"
    ),
    levels = c("Water", "10% PEG", "20% PEG")
    ),
    genotype = ordered(genotype, levels = c("WT", "cad4x5"))
  )

letters <- letter_groups(rosettes,
  area,
  medium,
  "tukey",
  genotype,
  print_position = "below"
)

area_medium <- ggplot(
  rosettes,
  aes(
    x = medium,
    y = area
  )
) +
  geom_quasirandom(
    aes(fill = genotype),
    shape = 21,
    # width = 0.1,
    # colour = NA,
    alpha = 0.75,
    size = 1,
    stroke = 0.2
  ) +
  geom_violin(aes(
    group = medium,
  ),
  draw_quantiles = 0.5,
  fill = "white",
  colour = "black",
  alpha = 0.85,
  width = 0.15,
  position = position_dodge(width = 0.6),
  size = 0.2,
  scale = "width"
  ) +
  geom_text(
    data = letters,
    aes(
      label = Letters,
      y = area,
    ),
    size = ggtext_size,
    family = "Helvetica",
    lineheight = 0.75,
  ) +
  scale_fill_manual(values = pal_ostwald_disc[c(1, 3)]) +
  scale_colour_manual(values = pal_ostwald_disc[c(1, 3)]) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 6)) +
  expand_limits(y = 0) +
  labs(y = "Rosette area [cm<sup>2</sup>]") +
  theme_leo() +
  theme(
    text = element_text(family = "Helvetica"),
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = ggtext::element_textbox_simple(
      orientation = "left-rotated",
      halign = 0.5
    ),
    axis.text.x = element_text(vjust = 1),
  ) +
  facet_wrap(~genotype)

pdf("area_facet.pdf", width = onecol, height = onecol * 0.25)
area_medium
dev.off()
  
```

### Evapotranspiration

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}
transpiration <- read_csv(
  "/data/PhD/IRX/Drought/evapotranspiration.csv"
) %>%
  pivot_longer(-c(medium, time), names_to = "plant", values_to = "transp_rate") %>%
  mutate(
    genotype = unlist(str_extract_all(plant, "[:alpha:]+")),
    genotype = ordered(recode(genotype, "CAD" = "cad4x5"),
      levels = c("WT", "cad4x5")
    ),
    replicate = unlist(str_extract_all(plant, "[:digit:]+")),
    medium = ordered(medium,
      levels = c("Water", "10% PEG", "20% PEG")
    ),
    transp_rate = transp_rate * 1000
  ) %>%
  select(-plant) %>%
  left_join(rosettes)

transpiration_avg <- transpiration %>%
  group_by(genotype, replicate, medium) %>%
  summarise(mean_transp = mean(transp_rate, na.rm = TRUE))

letters <- letter_groups(transpiration_avg,
  mean_transp,
  medium,
  "tukey",
  genotype,
  print_position = "below",
  print_adjust = 3
) %>%
  mutate(mean_transp = case_when(
    genotype == "WT" ~ mean_transp + 1,
    TRUE ~ mean_transp
  ))

transpiration_medium <- ggplot(
  transpiration,
  aes(
    x = medium,
    y = transp_rate
  )
) +
  geom_quasirandom(
    # aes(colour = genotype),
    # dodge.width = 0.6,
    shape = 16,
    # width = 0.1,
    # colour = NA,
    alpha = 0.15,
    size = 0.25,
    # stroke = 0.2
  ) +
  geom_quasirandom(
    data = transpiration_avg,
    aes(
      y = mean_transp,
      fill = genotype
    ),
    dodge.width = 0.6,
    shape = 21,
    width = 0.2,
    # colour = NA,
    alpha = 0.75,
    size = 1,
    stroke = 0.2
  ) +
  geom_violin(aes(
    group = medium,
    # fill = cell.type
  ),
  draw_quantiles = 0.5,
  fill = "white",
  colour = "black",
  alpha = 0.85,
  width = 0.15,
  position = position_dodge(width = 0.6),
  size = 0.2,
  scale = "width"
  ) +
  geom_text(
    data = letters,
    aes(
      label = Letters,
      y = mean_transp,
    ),
    size = ggtext_size,
    family = "Helvetica",
    lineheight = 0.75
  ) +
  scale_fill_manual(values = pal_ostwald_disc[c(1, 3)]) +
  scale_colour_manual(values = pal_ostwald_disc[c(1, 3)]) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 6)) +
  expand_limits(y = 0) +
  labs(y = "Evapotranspiration [mg min<sup>-1</sup>]") +
  theme_leo() +
  theme(
    text = element_text(family = "Helvetica"),
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = ggtext::element_textbox_simple(
      orientation = "left-rotated",
      halign = 0.5
    ),
    axis.text.x = element_text(vjust = 1),
  ) +
  facet_wrap(~genotype)

pdf("transpiration_facet.pdf", width = onecol, height = onecol * 0.25)
transpiration_medium
dev.off()

```

### Drought recovery

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}
recovery_data <- read_csv("/data/PhD/IRX/Drought/recovery.csv") %>%
  mutate(
    state = ordered(state, levels = rev(c("Recovered", "Partly recovered", "Dead"))),
    medium = ordered(medium, levels = c("Water", "10% PEG", "20% PEG")),
    genotype = ordered(genotype, levels = c("WT", "cad4x5"))
  ) %>%
  group_by(genotype, medium) %>%
  mutate(sum = sum(count)) %>%
  group_by(genotype, medium, state) %>%
  mutate(proportion = (count / sum) * 100)

recovery_collapse <- ggplot(recovery_data, aes(
  x = medium,
  y = proportion,
  fill = state
)) +
  geom_col(
    alpha = 0.85,
    width = 0.6
  ) +
  scale_fill_manual(values = rev(pal_ostwald_disc)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 6)) +
  theme_leo() +
  labs(
    y = "After 95 h in water [%]"
  ) +
  facet_wrap(~genotype, nrow = 1) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = ggtext::element_textbox_simple(
      orientation = "left-rotated",
      halign = 0.5
    ),
    axis.text.x = element_text(vjust = 1),
    legend.key.height = unit(2, "mm"),
    legend.key.width = unit(2, "mm"),
    legend.spacing = unit(c(0, 0, 0, 0), "mm"),
    plot.margin = unit(c(0.5, 6, 0.5, 6), "pt")
  ) +
  coord_cartesian(ylim = c(0, 100))

pdf("recovery.pdf", width = onecol, height = onecol * 0.5)
recovery_collapse
dev.off()
```

### Assemble figure
```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}
patch <- area_medium + transpiration_medium + recovery_collapse

pdf("IRX_Figure6.pdf", width = onecol * 0.55, height = onecol * 0.8)
patch + plot_layout(ncol = 1) &
  plot_annotation(
    theme = theme(plot.margin = margin(0, -2, 0, -2))
  ) &
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
dev.off()
```

## Supplemental figures

### Interaction terms

#### Interaction within the best models

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}
find_interactions <- function(x) {
  rmn_irx <- filter(rmn_irx, cell.type == x) %>%
    select(
      convexity,
      Perim.,
      # n_f,
      # n_p,
      n_v,
      cellulose,
      mean.OD1,
      total_lignin,
      # AUC,
      # G_rel,
      # GOH_rel,
      GOH.G,
      S_rel,
      # GCHO_rel,
      # S.G
    ) %>%
    drop_na()

  model.null <- lm(convexity ~ 1,
    data = rmn_irx
  )
  model.full <- lm(convexity ~ (.)^2,
    data = rmn_irx
  )

  best_model <- step(model.null,
    scope = list(upper = model.full),
    direction = "both",
    data = rmn_irx
  )

  print(tidy(best_model) %>%
    select(term) %>%
    filter(str_detect(term, fixed(":"))))

  best_model
}

#### protoxylem ####
best_model <- find_interactions("PX")

mean.OD1_total_lignin <- interact_plot(best_model,
  pred = mean.OD1, modx = total_lignin,
  colors = c("black", "black", "black"),
  x.label = expression(paste("Total "*bold(G)[CHO])),
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = c(0.75, 0.8),
    panel.background = element_rect(fill = pal_flame_disc[1])
  )
mean.OD1_total_lignin$layers[[1]]$aes_params$size <- 0.4


best_model <- find_interactions("PX")

total_lignin_Perim. <- interact_plot(best_model,
  pred = total_lignin, modx = Perim.,
  colors = c("black", "black", "black"),
  x.label = "Total lignin",
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = c(0.75, 0.8),
    panel.background = element_rect(fill = pal_flame_disc[1])
  ) +
  scale_y_continuous(
    labels = scales::number_format(accuracy = 0.01)
  )
total_lignin_Perim.$layers[[1]]$aes_params$size <- 0.4

S_rel_Perim. <- interact_plot(best_model,
  pred = S_rel, modx = Perim.,
  colors = c("black", "black", "black"),
  x.label = expression(paste(bold(S))),
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = c(0.75, 0.8),
    panel.background = element_rect(fill = pal_flame_disc[1])
  ) +
  scale_y_continuous(
    labels = scales::number_format(accuracy = 0.01)
  ) +
  scale_x_continuous(
    labels = scales::number_format(accuracy = 0.01)
  )
S_rel_Perim.$layers[[1]]$aes_params$size <- 0.4

mean.OD1_Perim. <- interact_plot(best_model,
  pred = mean.OD1, modx = Perim.,
  colors = c("black", "black", "black"),
  x.label = expression(paste("Total "*bold(G)[CHO])),
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = c(0.75, 0.8),
    panel.background = element_rect(fill = pal_flame_disc[1])
  )
mean.OD1_Perim.$layers[[1]]$aes_params$size <- 0.4

GOH.G_mean.OD1 <- interact_plot(best_model,
  pred = GOH.G, modx = mean.OD1,
  colors = c("black", "black", "black"),
  x.label = expression(paste(bold(G)[CHOH])),
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = c(0.75, 0.8),
    panel.background = element_rect(fill = pal_flame_disc[1])
  )
GOH.G_mean.OD1$layers[[1]]$aes_params$size <- 0.4

pdf("PX_interactions.pdf", height = twocol, width = twocol)
px <- plot_grid(
  total_lignin_Perim. + theme(legend.position = "none"),
  S_rel_Perim. + theme(legend.position = "none"),
  mean.OD1_Perim. + theme(legend.position = "none"),
  mean.OD1_total_lignin + theme(legend.position = "none"),
  "",
  GOH.G_mean.OD1 + theme(legend.position = "none"),
  "",
  ncol = 1
)
px
dev.off()

#### primary metaxylem ####
find_interactions <- function(x) {
  rmn_irx <- filter(rmn_irx, cell.type == x) %>%
    select(
      convexity,
      Perim.,
      # n_f,
      # n_p,
      n_v,
      cellulose,
      mean.OD1,
      total_lignin,
      # AUC,
      # G_rel,
      # GOH_rel,
      GOH.G,
      S_rel,
      GCHO_rel,
      # S.G
    ) %>%
    drop_na()

  model.null <- lm(convexity ~ 1,
    data = rmn_irx
  )
  model.full <- lm(convexity ~ (.)^2,
    data = rmn_irx
  )

  best_model <- step(model.null,
    scope = list(upper = model.full),
    direction = "both",
    data = rmn_irx
  )

  print(tidy(best_model) %>%
    select(term) %>%
    filter(str_detect(term, fixed(":"))))

  best_model
}

best_model <- find_interactions("PMX")

n_v_mean.OD1 <- interact_plot(best_model,
  pred = n_v, modx = mean.OD1,
  colors = c("black", "black", "black"),
  x.label = "Vessel adjacency",
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = c(0.75, 0.8),
    panel.background = element_rect(fill = alpha(pal_flame_disc[2], 0.5))
  )
n_v_mean.OD1$layers[[1]]$aes_params$size <- 0.4

GCHO_rel_total_lignin <- interact_plot(best_model,
  pred = GCHO_rel, modx = total_lignin,
  colors = c("black", "black", "black"),
  x.label = expression(paste("Terminal "*bold(G)[CHO])),
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = c(0.75, 0.8),
    panel.background = element_rect(fill = alpha(pal_flame_disc[2], 0.5))
  )
GCHO_rel_total_lignin$layers[[1]]$aes_params$size <- 0.4

mean.OD1_total_lignin <- interact_plot(best_model,
  pred = mean.OD1, modx = total_lignin,
  colors = c("black", "black", "black"),
  x.label = expression(paste("Total "*bold(G)[CHO])),
  y.label = "Convexity"
) +
  scale_y_continuous(
    labels = scales::number_format(accuracy = 0.01)
  ) +
  theme_leo() +
  theme(
    legend.position = c(0.75, 0.8),
    panel.background = element_rect(fill = alpha(pal_flame_disc[2], 0.5))
  )
mean.OD1_total_lignin$layers[[1]]$aes_params$size <- 0.4

pdf("PMX_interactions.pdf", height = onecol, width = onecol / 2)
pmx <- plot_grid(
  "",
  "",
  "",
  mean.OD1_total_lignin + theme(legend.position = "none"),
  GCHO_rel_total_lignin + theme(legend.position = "none"),
  "",
  n_v_mean.OD1 + theme(legend.position = "none"),
  ncol = 1
)
pmx
dev.off()

#### secondary metaxylem ####
find_interactions <- function(x) {
  rmn_irx <- filter(rmn_irx, cell.type == x) %>%
    select(
      convexity,
      Perim.,
      # n_f,
      # n_p,
      n_v,
      cellulose,
      mean.OD1,
      total_lignin,
      # AUC,
      # G_rel,
      # GOH_rel,
      GOH.G,
      S_rel,
      # GCHO_rel,
      # S.G
    ) %>%
    drop_na()

  model.null <- lm(convexity ~ 1,
    data = rmn_irx
  )
  model.full <- lm(convexity ~ (.)^2,
    data = rmn_irx
  )

  best_model <- step(model.null,
    scope = list(upper = model.full),
    direction = "both",
    data = rmn_irx
  )

  print(tidy(best_model) %>%
    select(term) %>%
    filter(str_detect(term, fixed(":"))))

  best_model
}

best_model <- find_interactions("SMX")

GOH.G_mean.OD1 <- interact_plot(best_model,
  pred = GOH.G, modx = mean.OD1,
  colors = c("white", "white", "white"),
  x.label = expression(paste(bold(G)[CHOH])),
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = c(0.75, 0.8),
    panel.background = element_rect(fill = alpha(pal_flame_disc[3], 0.5))
  )
GOH.G_mean.OD1$layers[[1]]$aes_params$size <- 0.4

mean.OD1_total_lignin <- interact_plot(best_model,
  pred = mean.OD1, modx = total_lignin,
  colors = c("white", "white", "white"),
  x.label = expression(paste("Total "*bold(G)[CHO])),
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = c(0.75, 0.8),
    panel.background = element_rect(fill = alpha(pal_flame_disc[3], 0.5))
  )
mean.OD1_total_lignin$layers[[1]]$aes_params$size <- 0.4

pdf("SMX_interactions.pdf", height = onecol, width = onecol / 2)
smx <- plot_grid(
  "",
  "",
  "",
  mean.OD1_total_lignin + theme(legend.position = "none"),
  "",
  GOH.G_mean.OD1 + theme(legend.position = "none"),
  "",
  ncol = 1
)
smx
dev.off()


legend <- get_legend(n_v_total_lignin + guides(
  linetype = guide_legend(title = "Interactor"),
  colour = "none",
  fill = "none"
) +
  theme(legend.justification = "top"))

pdf("interactions.pdf", height = twocol * 1.1, width = onecol)
plot_grid(px,
  pmx,
  smx,
  ncol = 3,
  labels = c("A", "B", "C"),
  vjust = 1,
  label_size = 10,
  label_fontfamily = "Helvetica"
)
dev.off()
```

#### Independent interactions 

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}
find_interactions <- function(x) {
  rmn_irx <- filter(rmn_irx, cell.type == x) %>%
    select(
      convexity,
      Perim.,
      # n_f,
      # n_p,
      n_v,
      cellulose,
      mean.OD1,
      total_lignin,
      # AUC,
      # G_rel,
      # GOH_rel,
      GOH.G,
      S_rel,
      # GCHO_rel,
      # S.G
    ) %>%
    drop_na()

  model.null <- lm(convexity ~ 1,
    data = rmn_irx
  )
  model.full <- lm(convexity ~ (.)^2,
    data = rmn_irx
  )

  best_model <- step(model.null,
    scope = list(upper = model.full),
    direction = "both",
    data = rmn_irx
  )

  print(tidy(best_model) %>%
    select(term) %>%
    filter(str_detect(term, fixed(":"))))

  best_model
}

best_model <- find_interactions("PX")

intrct <- paste0("Perim.", ":", "S_rel")
as.data.frame(effect(c(intrct), best_model))


plot_interactions <- function(ct, t1, t2) {
  ct <- enquo(ct)
  t1 <- enquo(t1)
  t2 <- enquo(t2)

  rmn_irx <- filter(rmn_irx, cell.type == quo_name(ct)) %>%
    select(
      convexity,
      Perim.,
      # n_f,
      # n_p,
      n_v,
      cellulose,
      mean.OD1,
      total_lignin,
      # AUC,
      # G_rel,
      # GOH_rel,
      GOH.G,
      S_rel,
      # GCHO_rel,
      # S.G
    ) %>%
    drop_na()

  model.null <- lm(convexity ~ 1,
    data = rmn_irx
  )
  model.full <- lm(convexity ~ (.)^2,
    data = rmn_irx
  )

  best_model <- step(model.null,
    scope = list(upper = model.full),
    direction = "both",
    data = rmn_irx
  )

  # intrct <- paste0(quo_name(t1), ":", quo_name(t2))

  # interact_fit <- as.data.frame(effect(c(intrct), best_model, xlevels = 3)) %>%
  #   mutate(!!t2 := as.character(!!t2))

  rmn_irx_bins <- rmn_irx %>%
    select(!!t1, !!t2, convexity) %>%
    drop_na() %>%
    mutate(interactor = ordered(case_when(
      !!t2 > mean(!!t2) + sd(!!t2) ~ "High",
      !!t2 < mean(!!t2) + sd(!!t2) & !!t2 > mean(!!t2) - sd(!!t2) ~ "Average",
      !!t2 < mean(!!t2) - sd(!!t2) ~ "Low"
    ), levels = c("High", "Average", "Low")))

  ggplot(rmn_irx_bins, aes(x = !!t1, y = convexity, colour = interactor)) +
    # scale_colour_manual(values = pal_ostwald_disc, name = term2_group) +
    geom_point() +
    labs(title = quo_name(t2)) +
    # geom_line()
    geom_smooth(method = "lm")
}

# plot_interactions(PX, Perim., S_rel)
# plot_interactions(PX, S_rel, Perim.)

# # plot_interactions(PX, Perim., total_lignin)
# plot_interactions(PX, total_lignin, Perim.)
#
# plot_interactions(PX, total_lignin, mean.OD1)
plot_interactions(PX, mean.OD1, total_lignin)
#   coord_cartesian(ylim = c(0.90,1))
#
# # plot_interactions(PX, Perim., mean.OD1)
# plot_interactions(PX, mean.OD1, Perim.)
#
# # plot_interactions(PX, GOH.G, mean.OD1)
# plot_interactions(PX, mean.OD1, GOH.G)
```

### Overview

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}
rmn_irx_long <- rmn_irx %>%
  select(genotype:cell.type, n_v, Area, Perim., Circ., convexity, total_lignin, G_rel, GOH.G, S_rel, cellulose, mean.OD1, GCHO_rel, S.G, GCHO.GOH, GCHO_term.GCHO_tot) %>%
  pivot_longer(cols = -c(genotype:cell.type), names_to = "variable", values_to = "value") %>%
  mutate(
    variable = factor(variable,
                      levels = c(
      "n_v",
      "Area",
      "cellulose",
      "Circ.",
      "convexity",
      "G_rel",
      "GOH.G",
      "mean.OD1",
      "GCHO_rel",
      "Perim.",
      "total_lignin",
      "S_rel",
      "S.G",
      "GCHO.GOH",
      "GCHO_term.GCHO_tot"
    ),
    labels = c("Adj. vessels",
               "Area [µm²]",
               "Cellulose",
               "Circularity",
               "Convexity",
               "bold(G)",
               "bold(G)[CHOH]",
               "Total~bold(G)[CHO]",
               "Terminal~bold(G)[CHO]",
               "Perimeter [µm]",
               "Lignin",
               "bold(S)",
               "bold(S)~'/'~bold(G)",
               "bold(G)[CHO]~'/'~bold(G)[CHOH]",
               "Term~'/'~tot~bold(G)[CHO]"
               )),
    genotype = ordered(genotype,
      levels = c(
        "Col-0",
        "4cl1",
        "4cl2",
        "4cl1x4cl2",
        "ccoaomt1",
        "fah1",
        "omt1",
        "ccr1-3",
        "cad4",
        "cad5",
        "cad4xcad5"
      )
    ),
    cell.type = ordered(cell.type, levels = c("PX", "PMX", "SMX"))
  )

rmn_irx_letters <- letter_groups(
  rmn_irx_long,
  value,
  genotype,
  "tukey",
  cell.type,
  variable,
  print_position = "above",
  print_adjust = 1,
  alpha = 0.05
) %>%
  mutate(value = case_when(variable == "Area [µm²]" & cell.type %in% c("PX", "SMX") ~
           value + 20,
           variable == "bold(G)[CHO]~'/'~bold(G)[CHOH]" & cell.type %in% c("PX", "SMX") ~
             value + 0.5,
           variable == "bold(G)[CHO]~'/'~bold(G)[CHOH]" & cell.type == "PMX" ~
             value + 0.2,
           variable == "Term~'/'~tot~bold(G)[CHO]" & cell.type %in% c("PMX", "SMX") ~
             value + 2,
         TRUE ~ value))

rmn_irx_lig <- ggplot(
  filter(rmn_irx_long, !variable %in%
           c("Adj. vessels",
             "Area [µm²]",
             "Circularity",
             "Convexity",
             "Perimeter [µm]")),
  aes(
    x = genotype,
    y = value
  )
) +
  geom_quasirandom(
    aes(fill = cell.type),
    shape = 21,
    # width = 0.1,
    # colour = NA,
    alpha = 0.75,
    size = 1,
    stroke = 0.25
  ) +
  geom_boxplot(
    fill = "white",
    alpha = 0.5,
    width = 0.6,
    outlier.alpha = 0,
    lwd = 0.25,
    fatten = 1,
    width = 0.25
  ) +
  geom_text(
    data = filter(rmn_irx_letters, 
                  !variable %in%
           c("Adj. vessels",
             "Area [µm²]",
             "Circularity",
             "Convexity",
             "Perimeter [µm]")),
    aes(label = Letters),
    size = 6 / (14 / 5),
    family = "Helvetica"
  ) +
  scale_fill_manual(values = pal_flame_disc) +
  scale_x_discrete(
    labels = c(
      "Col-0",
      expression(italic("4cl1")),
      expression(italic("4cl2")),
      expression(paste(italic("4cl1"), "x", italic("4cl2"))),
      expression(italic("ccoaomt1")),
      expression(italic("fah1")),
      expression(italic("omt1")),
      expression(italic("ccr1")),
      expression(italic("cad4")),
      expression(italic("cad5")),
      expression(paste(italic("cad4"), "x", italic("cad5")))
    )
  ) +
  theme_leo() +
  theme(
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    ),
    axis.title = element_blank()
  ) +
  facet_grid(variable ~ cell.type,
    scales = "free_y",
    labeller = label_parsed
  )

rmn_irx_morph <- ggplot(
  filter(rmn_irx_long, variable %in%
           c("Adj. vessels",
             "Area [µm²]",
             "Circularity",
             "Convexity",
             "Perimeter [µm]")),
  aes(
    x = genotype,
    y = value
  )
) +
  geom_quasirandom(
    aes(fill = cell.type),
    shape = 21,
    # width = 0.1,
    # colour = NA,
    alpha = 0.75,
    size = 1,
    stroke = 0.25
  ) +
  geom_boxplot(
    fill = "white",
    alpha = 0.5,
    width = 0.6,
    outlier.alpha = 0,
    lwd = 0.25,
    fatten = 1,
    width = 0.25
  ) +
  geom_text(
    data = filter(rmn_irx_letters, 
                  variable %in%
           c("Adj. vessels",
             "Area [µm²]",
             "Circularity",
             "Convexity",
             "Perimeter [µm]")),
    aes(label = Letters),
    size = 6 / (14 / 5),
    family = "Helvetica"
  ) +
  scale_fill_manual(values = pal_flame_disc) +
  scale_x_discrete(
    labels = c(
      "Col-0",
      expression(italic("4cl1")),
      expression(italic("4cl2")),
      expression(paste(italic("4cl1"), "x", italic("4cl2"))),
      expression(italic("ccoaomt1")),
      expression(italic("fah1")),
      expression(italic("omt1")),
      expression(italic("ccr1")),
      expression(italic("cad4")),
      expression(italic("cad5")),
      expression(paste(italic("cad4"), "x", italic("cad5")))
    )
  ) +
  theme_leo() +
  theme(
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    ),
    axis.title = element_blank()
  ) +
  facet_grid(variable ~ cell.type,
    scales = "free_y",
    # labeller = label_parsed
  )

pdf("IRX_FigureS1a.pdf", width = onecol * 1.5, height = onecol * 1.5)
rmn_irx_morph
dev.off()

pdf("IRX_FigureS1b.pdf", width = onecol * 1.5, height = twocol)
rmn_irx_lig
dev.off()
```

### In vitro Poplar

#### Load data

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE}

pop_irx <- read_csv("/data/PhD/IRX/Poplar/poplar_irx.csv")

#### create reference at the height of the cambium for distance calculation ####
pop_irx_ref <- pop_irx %>%
  filter(object == "ref") %>%
  select(c(
    "genotype",
    "replicate",
    "technical",
    "X",
    "Y",
    "Length"
  ))
pop_irx_ref$ref.y1 <- pop_irx_ref$Y
pop_irx_ref$ref.y2 <- pop_irx_ref$Y
pop_irx_ref$ref.x1 <- pop_irx_ref$X - (pop_irx_ref$Length / 2)
pop_irx_ref$ref.x2 <- pop_irx_ref$X + (pop_irx_ref$Length / 2)

#### merge reference into data frame ####
pop_irx <- pop_irx %>%
  filter(object != "ref") %>%
  full_join(select(pop_irx_ref, -X, -Y, -Length),
    by = c("genotype", "replicate", "technical")
  ) %>%
  mutate(genotype = ordered(genotype, levels = c("WT", "c4h", "ccr1")))

#### calculate difference of the centre of each vessel to the cambium ####
pop_irx$Distance <-
  apply(
    pop_irx[, c("X", "Y", "ref.x1", "ref.x2", "ref.y1", "ref.y2")],
    1,
    function(x) {
      a <- c(x[1], x[2])
      b <- c(x[3], x[5])
      c <- c(x[4], x[6])
      v1 <- b - c
      v2 <- a - b
      m <- cbind(v1, v2)
      d <- abs(det(m)) / sqrt(sum(v1 * v1))
      d
    }
  )

pop_irx <- pop_irx %>%
  mutate(bin = cut(
    Distance,
    breaks = c(-Inf, 50, 100, Inf),
    labels = c("I", "II", "III")
  ))

pop_irx <- pop_irx %>%
  mutate(
    convexity = Area / ConvexArea,
    Mean = Mean / 255,
    type = recode(type,
      "P" = "PX",
      "S" = "SX"
    )
  )
```

#### Wiesner by distance

```{r}

pop_irx_dist <- pop_irx %>%
  mutate(genotype = recode(genotype, 
                           "c4h" = "C4H-RNAi",
                           "ccr1" = "CCR1-RNAi"),
         type = recode(type, 
                       "PX" = "PV",
                       "SX" = "SV")) %>%
  group_by(genotype, replicate) %>%
  mutate(rel_dist = Distance / max(Distance))

pop_dist <- ggplot(pop_irx_dist,
       aes(x = rel_dist,
           y = Mean)) +
  geom_point(aes(fill = type),
    shape = 21,
    # width = 0.1,
    # colour = NA,
    alpha = 0.75,
    size = 1,
    stroke = 0.25) +
  geom_smooth(aes(colour = type),
              method = "lm",
              size = 0.4) +
  labs(x = "Relative distance",
       y = expression(paste("Relative "*bold(G)[CHO]))) +
  scale_x_continuous(breaks = c(0.05,0.95), labels = c("Cambium", "Pith")) +
  scale_colour_manual(values = pal_flame_disc[c(1,3)]) +
  scale_fill_manual(values = pal_flame_disc[c(1,3)]) +
  theme_leo() +
  theme(legend.position = "right",
        legend.title = element_blank()) +
  facet_wrap(~ genotype)

pdf("pop_dist.pdf", width = twocol * 0.975, height = onecol * 0.33)
pop_dist

```

#### Mutant overview

```{r}

pop_irx_overview <- pop_irx %>%
  ungroup() %>%
  select(
    genotype = genotype,
    type = type,
    `Adj. vessels` = n_v,
    Area = Area,
    Circularity = Circ.,
    Distance = Distance,
    Convexity = convexity
  ) %>%
  pivot_longer(`Adj. vessels`:Convexity, names_to = "variable", values_to = "value") %>%
  drop_na()

write_csv(
  pop_irx_overview %>%
    filter(genotype == "WT") %>%
    group_by(genotype, type, variable) %>%
    summarise(value = mean(value)),
  "WT_values.csv"
)


pop_letters <- letter_groups(pop_irx_overview,
  value,
  genotype,
  "kruskal",
  type,
  variable,
  print_position = "below",
  print_adjust = 0.75
) %>%
  mutate(value = case_when(
    variable == "Convexity" ~ value - 0.1,
    variable == "Circularity" ~ value - 0.1,
    TRUE ~ value
  ))

pop_overview <- ggplot(
  pop_irx_overview,
  aes(
    x = genotype,
    y = value
  )
) +
  geom_quasirandom(
    aes(fill = type),
    shape = 21,
    # width = 0.1,
    # colour = NA,
    alpha = 0.75,
    size = 1,
    stroke = 0.25
  ) +
  geom_violin(aes(
    group = genotype,
    # fill = cell.type
  ),
  draw_quantiles = 0.5,
  fill = "white",
  colour = "black",
  alpha = 0.85,
  width = 0.2,
  position = position_dodge(width = 0.6),
  size = 0.2,
  scale = "width"
  ) +
  # geom_boxplot(
  #   fill = "white",
  #   colour = "black",
  #   alpha = 0.5,
  #   width = 0.6,
  #   outlier.alpha = 0,
  #   lwd = 0.25,
  #   fatten = 1,
  #   width = 0.25
  # ) +
  geom_text(
    data = pop_letters,
    aes(label = Letters),
    size = 6 / (14 / 5),
    family = "Helvetica"
  ) +
  scale_fill_manual(values = pal_flame_disc[c(1, 3)]) +
  scale_x_discrete(
    labels = c(
      "WT",
      expression(paste(italic("C4H"), "-RNAi")),
      expression(paste(italic("CCR"), "-RNAi"))
    )
  ) +
  scale_y_continuous(expand = expand_scale(mult = c(0.1, 0.05))) +
  theme_leo() +
  theme(
    # axis.text.x = element_text(angle = 90,
    #                                hjust = 1),
    axis.title = element_blank()
  ) +
  expand_limits(y = 0) +
  facet_grid(variable ~ type,
    scales = "free_y"
  )


# pop_neighbourhood <- ggplot(pop_irx, aes(x = n_v, y = convexity)) +
#   geom_point(aes(fill = genotype),
#     shape = 21,
#     size = 2,
#     alpha = 0.75,
#     stroke = 0.2
#   ) +
#   scale_fill_few() +
#   scale_y_continuous(limits = c(0.7,1)) +
#   labs(x = "Fraction of perimeter adjacent to other vessels") +
#   theme_leo() +
#   geom_smooth(
#     method = "loess",
#     span = 1,
#     colour = "black",
#     linetype = 2
#   ) +
#   facet_wrap(~ bin, ncol = 1)
#
# pop_distance <- ggplot(pop_irx, aes(x = Distance, y = convexity)) +
#   geom_point(aes(fill = genotype),
#     shape = 21,
#     size = 2,
#     alpha = 0.75,
#     stroke = 0.2
#   ) +
#   scale_fill_few() +
#   scale_y_continuous(limits = c(0.7,1)) +
#   labs(x = "Distance from the cambium [µm]") +
#   theme_leo() +
#   geom_smooth(
#     method = "loess",
#     span = 1,
#     colour = "black",
#     linetype = 2
#   ) +
#   facet_wrap(~ bin, ncol = 1)

pdf("pop_overview.pdf", width = onecol * 0.9, height = onecol * 0.9)
pop_overview
dev.off()
```

#### WT overview

```{r}

pop_adj <- pop_irx %>%
  select(genotype:n_v) %>%
  filter(genotype == "WT") %>%
  pivot_longer(n_r:n_v, names_to = "variable", values_to = "value")

p_values <- pop_adj %>%
  group_by(variable) %>%
  summarise(p_value = broom::tidy(
    wilcox.test(value[type == "PX"], value[type == "SX"])
  )[["p.value"]]) %>%
  mutate(
    p_value = formatC(p_value, format = "e", digits = 1)
    #        case_when(
    # p_value > 0.001 ~ round(p_value, digits = 3),
    # p_value > 0.0001 ~ round(p_value, digits = 4),
    # p_value > 0.00001 ~ round(p_value, digits = 5),
    # p_value > 0.000001 ~ round(p_value, digits = 6),
    # p_value > 0.0000001 ~ round(p_value, digits = 7),
    # TRUE ~ round(p_value, digits = 8)
  )

pop_adj_box <- ggplot(
  data = pop_adj,
  aes(x = variable, y = value * 100)
) +
  geom_quasirandom(
    aes(
      group = interaction(type, variable),
      fill = type
    ),
    dodge.width = 0.6,
    shape = 21,
    # width = 0.1,
    # colour = NA,
    alpha = 0.75,
    size = 1,
    stroke = 0.25
  ) +
  # geom_boxplot(
  #   aes(group = interaction(cell.type, variable), fill = cell.type),
  #   # fill = "white",
  #   alpha = 0.5,
  #   width = 0.6,
  #   position = position_dodge2(width = 1),
  #   outlier.alpha = 0,
  #   lwd = 0.25,
  #   fatten = 1,
  #   width = 0.25
  # ) +
  geom_violin(aes(
    group = interaction(type, variable),
    # fill = cell.type
  ),
  draw_quantiles = 0.5,
  fill = "white",
  colour = "black",
  alpha = 0.85,
  width = 0.2,
  position = position_dodge(width = 0.6),
  size = 0.2,
  scale = "width"
  ) +
  geom_text(
    data = p_values,
    aes(label = paste(("P ="), p_value)),
    # x = 0.5,
    y = -5,
    hjust = 0.5,
    vjust = 1,
    family = "Helvetica",
    size = ggtext_size
  ) +
  annotate("text",
    x = c(0.6, 1.4),
    y = c(100, 100),
    label = c("PV", "SV"),
    family = "Helvetica",
    size = ggtext_size
  ) +
  annotate("segment",
    x = c(0.7, 1.3),
    xend = c(0.8, 1.2),
    y = c(100, 100),
    yend = c(100, 100),
    colour = "black",
    size = 0.2
  ) +
  scale_x_discrete(labels = c("Fibres", "Parenchyma", "Rays", "Vessels")) +
  scale_y_continuous(expand = expand_scale(mult = c(0.15, 0.05))) +
  # scale_fill_grey(start = 1, end = 0) +
  # scale_fill_brewer(palette = "Greys") +
  scale_fill_manual(values = pal_flame_disc[c(1, 3)]) +
  # scale_fill_manual(values = pal_ostwald[c(3, 4, 6)]) +
  # scale_colour_manual(values = pal_ostwald[c(3, 4, 6)]) +
  labs(
    x = "Neighbouring cell type",
    y = "CW adjacent [%]"
  ) +
  theme_leo() +
  theme(
    text = element_text(family = "Helvetica"),
    legend.position = "none"
  )

pdf("pop_adj_box.pdf", width = 9 / 2.54, height = 1.5)
pop_adj_box
dev.off()


rmn_violin <- function(x) {
  pop_irx_overview <- pop_irx %>%
    filter(genotype == "WT") %>%
    select(genotype:type, Mean, Distance, Area) %>%
    pivot_longer(c(Mean, Distance, Area), names_to = "variable", values_to = "value")

  p_values <- pop_irx_overview %>%
    filter(variable == x) %>%
    summarise(p_value = broom::tidy(
      t.test(value[type == "PX"], value[type == "SX"])
    )[["p.value"]]) %>%
    mutate(p_value = formatC(p_value, format = "e", digits = 1))

  ggplot(
    data = pop_irx_overview %>% filter(variable == x),
    aes(x = type, y = value)
  ) +
    geom_quasirandom(
      aes(fill = type),
      dodge.width = 0.6,
      shape = 21,
      # width = 0.1,
      # colour = NA,
      alpha = 0.75,
      size = 1,
      stroke = 0.2
    ) +
    # geom_boxplot(
    #   aes(group = interaction(cell.type, variable), fill = cell.type),
    #   # fill = "white",
    #   alpha = 0.5,
    #   width = 0.6,
    #   position = position_dodge2(width = 1),
    #   outlier.alpha = 0,
    #   lwd = 0.25,
    #   fatten = 1,
    #   width = 0.25
    # ) +
    geom_violin(aes(
      group = type,
      # fill = cell.type
    ),
    draw_quantiles = 0.5,
    fill = "white",
    colour = "black",
    alpha = 0.85,
    width = 0.2,
    position = position_dodge(width = 0.6),
    size = 0.2,
    scale = "width"
    ) +
    geom_text(
      data = p_values,
      aes(label = paste(("P ="), p_value)),
      x = 1.5,
      y = -0,
      hjust = 0.5,
      vjust = 1,
      family = "Helvetica",
      size = ggtext_size
    ) +
    scale_x_discrete(labels = c("PV", "SV")) +
    scale_fill_manual(values = pal_flame_disc[c(1, 3)]) +
    scale_y_continuous(expand = expand_scale(mult = c(0.15, 0.05))) +
    expand_limits(y = 0) +
    theme_leo() +
    theme(
      text = element_text(family = "Helvetica"),
      legend.position = "none",
    )
}

pop_area <- rmn_violin("Area") +
  labs(
    x = "",
    y = expression(paste("Area [µm"^2, "]"))
  )

pop_wiesner <- rmn_violin("Mean") +
  labs(
    x = "",
    y = expression(paste(bold(G)[CHO]))
  )

pop_distance <- rmn_violin("Distance") +
  labs(
    x = "",
    y = expression(paste("Distance [µm]"))
  )


celltypes_WT <-
  rasterGrob(readPNG(
    "/home/leonard/Dropbox/IRX manuscript/Submission/Edits/poplarWT.png"
  ))


right_low <- plot_grid(
  pop_area,
  pop_distance,
  pop_wiesner,
  labels = c("C", "D", "E"),
  nrow = 1,
  # rel_widths = c(1, 2.2),
  vjust = 1,
  label_size = 10,
  label_fontfamily = "Helvetica"
)

right <- plot_grid(
  pop_adj_box,
  right_low,
  labels = c("B", ""),
  nrow = 2,
  # rel_widths = c(1, 2.2),
  vjust = 1,
  label_size = 10,
  label_fontfamily = "Helvetica"
)

pdf("IRX_figure2_pop.pdf", width = onecol * 1.15, height = onecol * 0.75)
plot_grid(
  celltypes_WT,
  right,
  labels = c("A", ""),
  nrow = 1,
  # rel_widths = c(1, 2.2),
  vjust = 1,
  label_size = 10,
  label_fontfamily = "Helvetica",
  rel_widths = c(1, 2)
)
dev.off()
```

#### SEMs

```{r}

pop_irx <- as.data.frame(pop_irx)
poplar_psem <- psem(
  lm(Circ. ~
  n_v +
    Distance +
    # n_p +
    convexity,
  data = filter(pop_irx, type == "PX")
  ),
  lm(convexity ~
  n_v +
    # n_p +
    Mean +
    Distance +
    Perim.,
  data = filter(pop_irx, type == "PX")
  ),
  data = filter(pop_irx, type == "PX")
)

ind_claims <- ind_claims %>%
  bind_rows(summary(poplar_psem)[["dTable"]][, c(1, 4, 5)] %>%
    mutate(
      model = "PX",
      species = "Populus",
      fishC = summary(poplar_psem)[["Cstat"]][["Fisher.C"]],
      fishP = summary(poplar_psem)[["Cstat"]][["P.Value"]]
    ))

coefs <- coefs %>%
  bind_rows(summary(poplar_psem)[["coefficients"]][, c(1:8)] %>%
    mutate(
      model = "PX",
      species = "Populus"
    ))
plot(poplar_psem)

poplar_psem <- psem(
  lm(Circ. ~
  n_v +
    Distance +
    # n_p +
    convexity,
  data = filter(pop_irx, type == "SX")
  ),
  lm(convexity ~
  n_v +
    # n_p +
    Mean +
    Distance +
    Perim.,
  data = filter(pop_irx, type == "SX")
  ),
  data = filter(pop_irx, type == "SX")
)

ind_claims <- ind_claims %>%
  bind_rows(summary(poplar_psem)[["dTable"]][, c(1, 4, 5)] %>%
    mutate(
      model = "SX",
      species = "Populus",
      fishC = summary(poplar_psem)[["Cstat"]][["Fisher.C"]],
      fishP = summary(poplar_psem)[["Cstat"]][["P.Value"]]
    ))

coefs <- coefs %>%
  bind_rows(summary(poplar_psem)[["coefficients"]][, c(1:8)] %>%
    mutate(
      model = "SX",
      species = "Populus"
    ))
plot(poplar_psem)
```

#### Interactions 
```{r}
find_interactions <- function(x) {
  pop_irx <- filter(pop_irx, type == x) %>%
    select(
      convexity,
      Perim.,
      Mean,
      # n_f,
      # n_p,
      n_v,
      Distance
    ) %>%
    drop_na()

  model.null <- lm(convexity ~ 1,
    data = pop_irx
  )
  model.full <- lm(convexity ~ (.)^2,
    data = pop_irx
  )

  best_model <- step(model.null,
    scope = list(upper = model.full),
    direction = "both",
    data = pop_irx
  )

  print(tidy(best_model) %>%
    select(term) %>%
    filter(str_detect(term, fixed(":"))))

  best_model
}

#### protoxylem ####
best_model <- find_interactions("S")

Distance_n_v <- interact_plot(best_model,
  pred = Distance, modx = n_v,
  colors = c("white", "white", "white"),
  x.label = "Distance [µm]",
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = alpha(pal_flame_disc[3], 0.5))
  )
Distance_n_v$layers[[1]]$aes_params$size <- 0.2

Mean_n_v <- interact_plot(best_model,
  pred = n_v, modx = Mean,
  colors = c("white", "white", "white"),
  x.label = "Adjacent vessels",
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = alpha(pal_flame_disc[3], 0.5))
  )
Mean_n_v$layers[[1]]$aes_params$size <- 0.2

Mean_Distance <- interact_plot(best_model,
  pred = Distance, modx = Mean,
  colors = c("white", "white", "white"),
  x.label = "Distance [µm]",
  y.label = "Convexity"
) +
  theme_leo() +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = alpha(pal_flame_disc[3], 0.5))
  )
Mean_Distance$layers[[1]]$aes_params$size <- 0.2

pdf("pop_interactions.pdf", width = onecol, height = onecol / 3)
plot_grid(Distance_n_v,
  Mean_Distance,
  Mean_n_v,
  nrow = 1
)
dev.off()
```

### Woody Poplar

#### Load Raman data

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}
#### import and tidy poplar Raman data ####
rmn_files <-
  list.files(
    path = "/data/PhD/IRX/Poplar/2020-08_soil_poplar",
    pattern = "(TE|FI|RP)(.*).txt",
    recursive = TRUE,
    full.names = TRUE
  )

read_plus <- function(flnm) {
  read_tsv(flnm,
    comment = "#",
    col_names = c("wavenumber", "intensity"),
    col_types = "cc",
    # locale = locale(decimal_mark = ",")
  ) %>%
    mutate(
      filename = str_remove(
        flnm,
        fixed("/data/PhD/IRX/Poplar/2020-08_soil_poplar/")
      )
    ) %>%
    separate(filename, into = c("genotype", "replicate", "TE"), sep = "([/-])") %>%
    mutate(
      wavenumber = as.numeric(str_replace(wavenumber, fixed(","), fixed("."))),
      intensity = as.numeric(str_replace(intensity, fixed(","), fixed("."))),
      TE = str_remove(TE, fixed(".txt")),
      technical = unlist(str_extract_all(TE, "[:alpha:]+")),
      TE = unlist(str_extract_all(TE, "[:digit:]+")),
      cell_type = case_when(
        str_detect(technical, "TE") ~ "TE",
        str_detect(technical, "FI") ~ "fibre",
        TRUE ~ "ray"
      )
    ) %>%
    pivot_wider(
      id_cols = c(genotype, replicate, technical, cell_type, TE),
      names_from = wavenumber,
      values_from = intensity
    ) %>%
    group_by(genotype, replicate, technical, cell_type, TE) %>%
    nest()
}

rmn_pop <- lapply(rmn_files, read_plus) %>%
  bind_rows() %>%
  filter(cell_type == "TE") %>%
  ungroup() %>%
  select(-cell_type) %>%
  anti_join(read_csv("/data/PhD/IRX/Poplar/2020-08_soil_poplar/pop_rmn_exclude.csv",
    col_types = "cccc"
  ))

#### baseline correction ####
rmn_als <- function(x) {
  corrected_spectra <- baseline.als(as.matrix(x),
    lambda = 5,
    p = 0.01,
    maxit = 100
  )
  as_tibble(corrected_spectra$corrected, rownames = NA)
}

rmn_pop_corrected <- rmn_pop %>%
  mutate(corrected_data = map(data, rmn_als)) %>%
  select(-data) %>%
  unnest() %>%
  pivot_longer(-c(genotype, replicate, technical, TE),
    names_to = "wavenumber",
    values_to = "corrected.intensity"
  ) %>%
  drop_na() %>%
  mutate(wavenumber = as.numeric(wavenumber))

#### aligment, filtering and scaling ####

# alignment
lig_peak <- rmn_pop_corrected %>%
  unite("ID",
    genotype, replicate, technical, TE,
    remove = F,
    sep = "_"
  ) %>%
  filter(round(wavenumber, digits = 0) %in% c(1560:1640)) %>%
  group_by(genotype, replicate, technical, TE) %>%
  mutate(
    peak_pos = wavenumber[which.max(corrected.intensity)]
  ) %>%
  select(-wavenumber, -corrected.intensity) %>%
  unique() %>%
  ungroup() %>%
  mutate(peak_pos_ref = case_when(
    genotype == "T89" ~
    peak_pos[ID == "T89_4_alignedTE_1"],
    genotype == "CCR1" ~
    peak_pos[ID == "CCR1_1_TE_1"],
    genotype == "C4H" ~
    peak_pos[ID == "C4H_4_cTE_1"],
    TRUE ~ 0
  ))

rmn_pop_aligned <- rmn_pop_corrected %>%
  left_join(lig_peak) %>%
  group_by(genotype, replicate, technical, TE, wavenumber) %>%
  mutate(wavenumber = wavenumber + (peak_pos_ref - peak_pos))

# scaling and filtering
rmn_pop_scaled <- rmn_pop_aligned %>%
  group_by(genotype, replicate, technical, TE) %>%
  mutate(
    AUC = MESS::auc(wavenumber, corrected.intensity),
    scaled = corrected.intensity / AUC,
    scaled_lig = corrected.intensity / (
      (corrected.intensity[abs(wavenumber - 1600) == min(abs(wavenumber - 1600))][1] / 2) + corrected.intensity[abs(wavenumber - 1330) == min(abs(wavenumber - 1330))][1]
    )
  ) %>%
  filter(wavenumber > 200 & wavenumber < 1700)

rmn_pop_avg <- rmn_pop_scaled %>%
  mutate(wavenumber = round(wavenumber, digits = 1)) %>%
  group_by(genotype) %>%
  arrange(genotype, wavenumber) %>%
  summarise(
    # roll_mean = RcppRoll::roll_mean(scaled, 50, fill = NA, na.rm = TRUE),
    roll_mean = slider::slide_index_dbl(scaled, wavenumber, mean, .before = 1, .after = 1),
    roll_sd = slider::slide_index_dbl(scaled, wavenumber, sd, .before = 2, .after = 2),
    wavenumber = wavenumber
    )

rmn_pop_sum <- rmn_pop_scaled %>%
  group_by(genotype, replicate, technical, TE) %>%
  summarise(
    "unscaled_lignin" = 
      (corrected.intensity[abs(wavenumber - 1600) == min(abs(wavenumber - 1600))][1] / 2) +
      corrected.intensity[abs(wavenumber - 1330) == min(abs(wavenumber - 1330))][1],
    "broad_lignin" = 
      corrected.intensity[abs(wavenumber - 1600) == min(abs(wavenumber - 1600))][1] +
      corrected.intensity[abs(wavenumber - 1330) == min(abs(wavenumber - 1330))][1] +
      corrected.intensity[abs(wavenumber - 1660) == min(abs(wavenumber - 1660))][1] +
      corrected.intensity[abs(wavenumber - 1625) == min(abs(wavenumber - 1625))][1],
    "total_lignin" = (scaled[abs(wavenumber - 1600) == min(abs(wavenumber - 1600))][1] / 2) +
      scaled[abs(wavenumber - 1330) == min(abs(wavenumber - 1330))][1],
    "P_rel" = scaled_lig[abs(wavenumber - 1000) == min(abs(wavenumber - 1000))][1],
    "G_rel" = scaled_lig[abs(wavenumber - 1600) == min(abs(wavenumber - 1600))][1],
    "G_tot" = scaled[abs(wavenumber - 1600) == min(abs(wavenumber - 1600))][1],
    "GOH_rel" = scaled_lig[abs(wavenumber - 1660) == min(abs(wavenumber - 1660))][1],
    "GCHO_rel" = scaled_lig[abs(wavenumber - 1625) == min(abs(wavenumber - 1625))][1],
    "GCHO_tot" = scaled[abs(wavenumber - 1625) == min(abs(wavenumber - 1625))][1],
    "S_rel" = scaled_lig[abs(wavenumber - 1330) == min(abs(wavenumber - 1330))][1],
    "S_tot" = scaled[abs(wavenumber - 1330) == min(abs(wavenumber - 1330))][1],
    "cellulose" =
      scaled[abs(wavenumber - 1095) == min(abs(wavenumber - 1095))][1] +
        scaled[abs(wavenumber - 375) == min(abs(wavenumber - 375))][1] +
        scaled[abs(wavenumber - 1120) == min(abs(wavenumber - 1120))][1],
    "lig.cellu" = total_lignin / cellulose,
    "G.cellu" = G_tot / cellulose,
    "S.G" = scaled_lig[abs(wavenumber - 1330) == min(abs(wavenumber - 1330))][1] /
      scaled_lig[abs(wavenumber - 1276) == min(abs(wavenumber - 1276))][1],
    # "GOH.G" = scaled_lig[wavenumber == 1660] / scaled_lig[wavenumber == 1600],
    "GCHO.GOH" = scaled_lig[abs(wavenumber - 1625) == min(abs(wavenumber - 1625))][1] /
      scaled_lig[abs(wavenumber - 1660) == min(abs(wavenumber - 1660))][1],
    "AUC" = mean(AUC)
  )


rmn_pop_test <- ggplot(
  rmn_pop_scaled,
  aes(
    x = wavenumber,
    y = scaled
  )
) +
  scale_x_continuous(trans = "reverse",
                     # limits = c(1020, 980)
                     ) +
  # scale_y_continuous(limits = c(-0.001, 0.003)) +
  geom_vline(xintercept = c(1660, 1600, 1330, 1120, 1095, 1001, 375),
             size = 0.2) +
  annotate("text",
           label = c("1660", "1600", "1330", "1120", "1095", "1001", "375"),
           y = 0.007,
           x = c(1661, 1601, 1331, 1170, 1096, 1002, 376),
           hjust = 0,
           family = "Helvetica",
           size = ggtext_size) +
  # geom_line(aes(
  #   group = interaction(genotype, replicate, technical, TE),
  #   colour = genotype
  # ),
  # alpha = 0.25,
  # size = 0.1
  # ) +
  geom_ribbon(data = rmn_pop_avg,
            aes(fill = genotype,
                y = roll_mean,
                ymin = roll_mean - roll_sd,
                ymax = roll_mean + roll_sd),
            colour = NA,
            alpha = 0.25) +
  geom_line(data = rmn_pop_avg,
            aes(colour = genotype,
                y = roll_mean),
            size = 0.2) +
  scale_colour_manual(values = pal_ostwald_disc[c(3, 2, 1)]) +
  scale_fill_manual(values = pal_ostwald_disc[c(3, 2, 1)]) +
  theme_leo() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  ) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 1)))

# plotly::ggplotly(rmn_pop_test)
pdf("rmn_pop.pdf", height = 3)
rmn_pop_test
dev.off()
```

#### Export scaled spectra

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}
export_raw <- rmn_pop_scaled %>%
  select(-c(AUC, scaled, scaled_lig)) %>%
  mutate(wavenumber = round(wavenumber, digits = 0)) %>%
  pivot_wider(id_cols = c(genotype, replicate, technical, TE), 
              names_from = wavenumber,
              values_from = corrected.intensity) %>%
  select(genotype, replicate, technical, TE, order(as.numeric(colnames(.))))

export_scaled <- rmn_pop_scaled %>%
  select(-c(AUC, corrected.intensity, scaled_lig)) %>%
  mutate(wavenumber = round(wavenumber, digits = 0)) %>%
  pivot_wider(id_cols = c(genotype, replicate, technical, TE), 
              names_from = wavenumber,
              values_from = scaled) %>%
  select(genotype, replicate, technical, TE, order(as.numeric(colnames(.))))

write_excel_csv(export_raw, "soil_poplar_raw.csv")
write_excel_csv(export_scaled, "soil_poplar_scaled.csv")
```

#### Load shape data

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}

pop_irx <- read_csv(
  "/data/PhD/IRX/Poplar/2020-08_soil_poplar/pop_rmn_coords.csv"
  )

#### create reference at the height of the cambium for distance calculation ####
pop_irx_ref <- pop_irx %>%
  filter(TE == 0) %>%
  group_by(genotype,
           replicate,
           technical,
           ref_line) %>%
  summarise(ref.y1 = unique(y),
            ref.y2 = unique(y),
            ref.x1 = min(x),
            ref.x2 = max(x))
  
  # select(c(
  #   "genotype",
  #   "replicate",
  #   "technical",
  #   "ref_line",
  #   "X",
  #   "Y",
  #   "Length"
  # )) %>%
  # mutate(
  #   "ref.y1" = Y,
  #   "ref.y2" = Y,
  #   "ref.x1" = X - (Length / 2),
  #   "ref.x2" = X + (Length / 2)
  # )

#### merge reference into data frame ####
pop_irx <- pop_irx %>%
  filter(TE != 0) %>%
  left_join(pop_irx_ref,
    by = c("genotype", "replicate", "technical", "ref_line")
  ) %>%
  mutate(
    TE = as.character(TE),
    replicate = as.character(replicate)
  )

#### calculate difference of the centre of each vessel to the cambium ####
pop_irx$Distance <-
  apply(
    pop_irx[, c("X", "Y", "ref.x1", "ref.x2", "ref.y1", "ref.y2")],
    1,
    function(x) {
      a <- c(x[1], x[2])
      b <- c(x[3], x[5])
      c <- c(x[4], x[6])
      v1 <- b - c
      v2 <- a - b
      m <- cbind(v1, v2)
      d <- abs(det(m)) / sqrt(sum(v1 * v1))
      d
    }
  )

celltypes <- read_csv(
  "/data/PhD/IRX/Poplar/2020-08_soil_poplar/poplar_cell_types.csv",
  col_types = "cccccc")

rmn_pop_data <- rmn_pop_sum %>%
  left_join(pop_irx) %>%
  left_join(celltypes) %>%
  nest(point, x, y) %>%
  mutate(genotype = recode(genotype, "T89" = "WT"),
         genotype = ordered(genotype, levels = c("WT", "CCR1", "C4H")))

rmn_pop_n <- rmn_pop_data %>%
  group_by(genotype, cell_type) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(overall = sum(n))

```

#### Scatter plots

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}


ggplot(rmn_pop_data,
       aes(x = total_lignin,
           y = Solidity)
       )+
  geom_point(aes(colour = genotype)) +
  geom_smooth(method = "lm")

rmn_lignin_dist <- rmn_pop_data %>%
  unite("ID", genotype, replicate,
        sep = "_",
        remove = FALSE) %>%
  filter(Distance < 2500) %>%
  # filter(ID != "WT_1") %>%
  select(genotype, replicate, cell_type, technical, Distance, TE,
         # "Cellulose (381 + 1099)" = cellulose,
         # "AUC scaled lignin" = total_lignin,
         # "Total G-units (1603)" = G_tot,
         # "Total S-units (1334)" = S_tot,
         # "Phenyl" = P_rel,
         "Lignin/Cellulose" = lig.cellu,
         # "Unscaled lignin" = unscaled_lignin,
         # "AUC" = AUC,
         # "Complete lignin" = broad_lignin,
         # "Total G/Cellulose" = G.cellu,
         # "Relative G-units" = G_rel, 
         # "Relative S-units" = S_rel, 
         # "Relative aldehyde end-units" = GCHO_rel,
         "S/G" = S.G,
         "G-CHO/G-CHOH" = GCHO.GOH,
         # "Convexity" = Solidity
         ) %>%
  drop_na() %>%
  pivot_longer(-c(genotype, replicate, cell_type, technical, Distance, TE), 
               names_to = "variable", 
               values_to = "value")

# bulk model
regressions <- rmn_lignin_dist %>%
  ungroup() %>%
  filter(cell_type == "SV") %>%
  select(-replicate, -technical) %>%
  nest(data = c(TE, Distance, value)) %>%
  group_by(variable) %>%
  mutate(max = max(map_dbl(data, ~ max(.x$value))),
         min = max(map_dbl(data, ~ min(.x$value)))) %>%
  group_by(genotype, variable) %>%
  mutate(
    fit = map(data, ~ lm(value ~ Distance, data = .x)),
    tidied = map(fit, glance),
    Distance = map_dbl(data, ~ max(.x$Distance))
  ) %>%
  unnest(tidied) %>%
  select(-data, -fit) %>%
  mutate(face = case_when(p.value > 0.05 ~ "plain",
                          TRUE ~ "bold")) %>%
  group_by(variable) %>%
  mutate(
    Distance = max(Distance),
    spacer = max - min,
    value = case_when(
      genotype == "WT" ~ max + (0.2 * spacer),
      genotype == "CCR1" ~ max + (0.1 * spacer),
      genotype == "C4H" ~ max
    )
  )

# mixed model
regressions <- rmn_lignin_dist %>%
  ungroup() %>%
  filter(cell_type == "SV") %>%
  select(-technical) %>%
  nest(data = c(TE, replicate, Distance, value)) %>%
  group_by(variable) %>%
  mutate(max = max(map_dbl(data, ~ max(.x$value))),
         min = max(map_dbl(data, ~ min(.x$value)))) %>%
  group_by(genotype, variable) %>%
  mutate(
    fit = map(data, ~ lme4::lmer(value ~ Distance + 
                                   # (1 + Distance | replicate),
                                   (1 | replicate),
                                 data = .x)),
    pred = map(fit, predict),
    tidied = map(fit, broom.mixed::glance),
    r2 = map(fit, MuMIn::r.squaredGLMM),
    r2 = map(r2, 2),
    r2 = as.numeric(r2),
    Distance = map_dbl(data, ~ max(.x$Distance))
  ) %>%
  group_by(variable) %>%
  mutate(
    Distance = max(Distance),
    spacer = case_when(variable == "Lignin/Cellulose" ~ (max - min) * 1.3,
                       TRUE ~ max - min),
    value = case_when(
      genotype == "WT" ~ max + (0.2 * spacer),
      genotype == "CCR1" ~ max + (0.1 * spacer),
      genotype == "C4H" ~ max
    )
  )

rmn_lignin_dist_pred <- regressions %>%
  unnest(data, pred)

dist_lig <- ggplot(
  rmn_lignin_dist_pred %>% filter(cell_type == "SV"),
  aes(
    x = Distance,
    y = value,
    colour = genotype,
    # fill = replicate
  )
) +
  geom_point(
    aes(shape = replicate),
    alpha = 0.5
    ) +
  geom_smooth(
    size = 0.5,
    method = "lm",
    se = F
  ) +
  ggtext::geom_richtext(
    data = regressions,
    size = ggtext_size,
    vjust = 0,
    hjust = 1,
    # colour = "black",
    label.colour = NA,
    family = "Helvetica",
    fill = NA,
    aes(label =
          paste0(genotype,
                 ": R<sup>2</sup> = ",
                 round(r2, 2)
                 # ", <i>P</i> = ",
                 # format(p.value, scientific = TRUE, digits = 2)
                 ),
        colour = genotype,
        # fontface = face
        )
  ) +
  # ggrepel::geom_text_repel(aes(label = TE)) +
  theme_leo() +
  theme(axis.title.y = element_blank()) +
  labs(x = "Distance [px]") +
  # coord_cartesian(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
  scale_colour_manual(values = pal_ostwald_disc) +
  facet_wrap(vars(variable),
             scales = "free_y",
             nrow = 1)

pdf("post_mortem_lignification.pdf", width = twocol, height = onecol * 0.5)
dist_lig
dev.off()


dist_lig1 <- ggplot(
  rmn_lignin_dist_pred %>% filter(cell_type == "SV"),
  aes(
    x = Distance,
    y = value,
    colour = genotype,
    # fill = replicate
  )
) +
  geom_point(
    aes(shape = replicate),
    size = 1,
    alpha = 0.25
    ) +
  geom_smooth(
    aes(group = interaction(genotype, replicate),
        y = pred),
    size = 0.5,
    method = "lm",
    se = F
  ) +
  ggtext::geom_richtext(
    data = regressions,
    size = ggtext_size,
    vjust = 0,
    hjust = 1,
    # colour = "black",
    label.colour = NA,
    family = "Helvetica",
    fill = NA,
    aes(label =
          paste0(genotype,
                 ": R<sup>2</sup> = ",
                 round(r2, 2)
                 # ", <i>P</i> = ",
                 # format(p.value, scientific = TRUE, digits = 2)
                 ),
        colour = genotype,
        # fontface = face
        )
  ) +
  # ggrepel::geom_text_repel(aes(label = TE)) +
  theme_leo() +
  theme(axis.title.y = element_blank()) +
  labs(x = "Distance [px]") +
  # coord_cartesian(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
  scale_colour_manual(values = pal_ostwald_disc) +
  facet_wrap(vars(variable),
             scales = "free_y",
             nrow = 1)

pdf("individual_intercept.pdf", width = twocol, height = onecol * 0.5)
dist_lig1
dev.off()

dist_lig2 <- ggplot(
  rmn_lignin_dist_pred %>% filter(cell_type == "SV"),
  aes(
    x = Distance,
    y = value,
    colour = genotype,
    # fill = replicate
  )
) +
  geom_point(
    aes(shape = replicate),
    alpha = 0.5
    ) +
  geom_smooth(
    aes(group = interaction(genotype, replicate),
        y = value),
    size = 0.5,
    method = "loess",
    span = 5,
    se = F
  ) +
  # ggtext::geom_richtext(
  #   data = regressions,
  #   size = ggtext_size,
  #   vjust = 0,
  #   hjust = 1,
  #   # colour = "black",
  #   label.colour = NA,
  #   family = "Helvetica",
  #   fill = NA,
  #   aes(label =
  #         paste0(genotype,
  #                ": R<sup>2</sup> = ",
  #                round(r2, 2)
  #                # ", <i>P</i> = ",
  #                # format(p.value, scientific = TRUE, digits = 2)
  #                ),
  #       colour = genotype,
  #       # fontface = face
  #       )
  # ) +
  # ggrepel::geom_text_repel(aes(label = TE)) +
  theme_leo() +
  theme(axis.title.y = element_blank()) +
  labs(x = "Distance [px]") +
  coord_cartesian(clip = "off") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
  scale_colour_manual(values = pal_ostwald_disc) +
  facet_wrap(vars(variable),
             scales = "free_y",
             nrow = 1)

pdf("individual_slope+intercept.pdf", width = twocol, height = onecol * 0.5)
dist_lig2
dev.off()

```

#### Violin plots

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}

letters <- letter_groups(rmn_pop_data,
                         Solidity,
                         genotype,
                         "tukey",
                         cell_type,
                         print_position = "below"
                         )

test <- rmn_pop_data %>%
  group_by(genotype, cell_type) %>%
  drop_na(Solidity, cell_type, genotype) %>%
  summarise(n = n())

convexity_poplar <- ggplot(drop_na(rmn_pop_data, genotype, Solidity, cell_type),
       aes(x = genotype,
           y = Solidity)) +
  geom_quasirandom(
      aes(fill = genotype),
      # dodge.width = 0.6,
      shape = 21,
      # width = 0.1,
      # colour = NA,
      alpha = 0.75,
      size = 1,
      stroke = 0.2
    ) +
    geom_violin(aes(
      group = genotype,
      # fill = cell.type
    ),
    draw_quantiles = 0.5,
    fill = "white",
    colour = "black",
    alpha = 0.85,
    width = 0.2,
    # position = position_dodge(width = 0.6),
    size = 0.2,
    scale = "width"
    ) +
    geom_text(
      data = letters,
      aes(label = Letters),
      size = ggtext_size,
      family = "Helvetica"
    ) +
    # scale_x_discrete(labels = c("PX", "PMX", "SMX")) +
    scale_fill_manual(values = pal_ostwald_disc) +
    # scale_fill_grey(start = 0.9, end = 0.3) +
    # expand_limits(y = 0) +
    # scale_y_continuous(expand = expand_scale(mult = c(0.1, 0.05))) +
    # labs(y = "Convexity") +
    theme_leo() +
    theme(
      text = element_text(family = "Helvetica"),
      legend.position = "none",
      axis.title.x = element_blank(),
      # plot.margin = unit(c(0.2, 0.2, 0.1, 0.1), "mm")
    ) +
  facet_wrap(~ cell_type)

pdf("poplar_convexity.pdf", width = onecol * 0.5, height = onecol * 0.5)
convexity_poplar
dev.off()
```

#### Model otimisation

```{r, fig.showtext = TRUE, out.width = '100%', message = FALSE, eval = TRUE, echo = FALSE}

psem_data_pop <- rmn_pop_data %>%
  ungroup() %>%
  filter(cell_type == "SV") %>%
  select(
    Solidity,
    # n_v,
    # Perim.,
    cellulose,
    total_lignin,
    G_rel,
    S_rel,
    P_rel,
    GOH_rel,
    GCHO_rel,
    lig.cellu,
    Distance,
  ) %>%
  drop_na()

# Fit the full model 
full_model <- lm(Solidity ~., data = psem_data_pop)
# Stepwise regression model
step_model <- MASS::stepAIC(full_model, direction = "both", 
                      trace = FALSE)
summary(step_model)


```

#### SEM 

```{r}

psem_data_pop <- rmn_pop_data %>%
  filter(cell_type == "SV") %>%
  select(-data, -Length, -n_v) %>%
  drop_na()

psem_pop <- psem(
  lm(Circ. ~
  # n_v +
    Solidity,
  data = psem_data_pop
  ),
  lm(Solidity ~
  # n_v +
    Distance +
    # cellulose +
    P_rel +
    # n_f +
    # mean.OD1 +
    # Perim. +
    # S.G +
    # G_tot +
    GOH_rel +
    # GOH.G +
    # G_rel +
    # GOH_tot +
    # S_rel +
    # S_tot +
    GCHO_rel +
    # total_lignin:mean.OD1 +
    total_lignin,
  data = data.frame(psem_data_pop)
  ),
  data = data.frame(psem_data_pop)
)
# ind_claims <- ind_claims %>%
#   bind_rows(summary(psem_pmx)[["dTable"]][, c(1, 4, 5)] %>%
#     mutate(
#       model = "PMX",
#       species = "Arabidopsis",
#       fishC = summary(psem_pmx)[["Cstat"]][["Fisher.C"]],
#       fishP = summary(psem_pmx)[["Cstat"]][["P.Value"]]
#     ))
# 
# coefs <- coefs %>%
#   bind_rows(summary(psem_pmx)[["coefficients"]][, c(1:8)] %>%
#     mutate(
#       model = "PMX",
#       species = "Arabidopsis"
#     ))
plot(psem_pop)
summary(psem_pop)

```

### Bending

#### Curves

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}

curves <- ggplot(
  data = bending_data_complete,
  aes(x = strain, y = stress)
) +
  stat_smooth(aes(group = interaction(genotype, replicate, section)),
    geom = "line",
    size = 0.2,
    alpha = 0.15,
    colour = "black",
    se = F
  ) +
  geom_smooth(aes(colour = genotype),
    size = 0.4
  ) +
  # geom_smooth(
  #   data = filter(
  #     bending_data_complete,
  #     genotype == "WT" &
  #       strain < 0.25
  #   ),
  #   aes(
  #     x = strain,
  #     y = stress
  #   ),
  #   method = "lm",
  #   se = F,
  #   fullrange = T,
  #   colour = "black",
  #   linetype = 2,
  #   size = 0.2
  # ) +
  # geom_line(
  #   data = filter(bending_data_complete, genotype == "WT"),
  #   aes(x = strain, y = 14),
  #   linetype = 1,
  #   colour = "black",
  #   size = 0.2
  # ) +
  # geom_segment(
  #   data = filter(bending_data_complete, genotype == "WT")[1, ],
  #   aes(x = 0.75, xend = 0.75, y = 0, yend = 50),
  #   linetype = 3,
  #   colour = "black",
  #   size = 0.2
  # ) +
  # geom_text(
  #   data = filter(bending_data_complete, genotype == "WT")[1, ],
  #   aes(x = 0.71, y = 45, label = br),
  #   hjust = 1,
  #   vjust = 0,
  #   angle = 90,
  #   family = "Helvetica",
  #   size = ggtext_size
  # ) +
  # geom_text(
  #   data = filter(bending_data_complete, genotype == "WT")[1, ],
  #   aes(x = 3, y = 15.5, label = str),
  #   hjust = 1,
  #   vjust = 0,
  #   family = "Helvetica",
  #   size = ggtext_size
  # ) +
  # geom_text(
  #   data = filter(bending_data_complete, genotype == "WT")[1, ],
  #   aes(x = 1.53, y = 45, label = stiff),
  #   hjust = 1,
  #   vjust = 0,
  #   angle = 52,
  #   family = "Helvetica",
  #   size = ggtext_size
  # ) +
  scale_colour_manual(values = c(
    pal_ostwald_disc[1],
    pal_ostwald_disc[3],
    pal_ostwald_disc[3]
  )) +
  coord_cartesian(ylim = c(0, 45)) +
  theme_leo() +
  xlim(0, 3) +
  labs(
    x = "Strain [%]",
    y = "Stress [MPa]"
  ) +
  facet_wrap(~genotype, nrow = 3)

medium_curves <- medium_curves %>%
  mutate(
    stiff = "Stiffness",
    str = "Strength",
    br = "Breaking"
  )

curves_media <- ggplot(
  data = medium_curves,
  aes(x = strain, y = stress)
) +
  stat_smooth(aes(group = interaction(medium, replicate, section)),
    geom = "line",
    size = 0.2,
    alpha = 0.15,
    colour = "black",
    se = F
  ) +
  geom_smooth(
    colour = pal_ostwald_disc[1],
    size = 0.4
  ) +
  coord_cartesian(ylim = c(0, 20)) +
  theme_leo() +
  xlim(0, 3) +
  labs(
    x = "Strain [%]",
    y = "Stress [MPa]"
  ) +
  geom_smooth(
    data = filter(
      medium_curves,
      medium == "Water" &
        strain < 0.25
    ),
    aes(
      x = strain,
      y = stress
    ),
    method = "lm",
    se = F,
    fullrange = T,
    colour = "black",
    linetype = 2,
    size = 0.2
  ) +
  geom_line(
    data = filter(medium_curves, medium == "Water"),
    aes(x = strain, y = 9),
    linetype = 1,
    colour = "black",
    size = 0.2
  ) +
  geom_segment(
    data = filter(medium_curves, medium == "Water")[1, ],
    aes(x = 0.675, xend = 0.675, y = 0, yend = 50),
    linetype = 3,
    colour = "black",
    size = 0.2
  ) +
  geom_text(
    data = filter(medium_curves, medium == "Water")[1, ],
    aes(x = 0.6, y = 20.5, label = br),
    hjust = 1,
    vjust = 0,
    angle = 90,
    family = "Helvetica",
    size = ggtext_size
  ) +
  geom_text(
    data = filter(medium_curves, medium == "Water")[1, ],
    aes(x = 3, y = 10, label = str),
    hjust = 1,
    vjust = 0,
    family = "Helvetica",
    size = ggtext_size
  ) +
  geom_text(
    data = filter(medium_curves, medium == "Water")[1, ],
    aes(x = 1.3, y = 20, label = stiff),
    hjust = 1,
    vjust = 0,
    angle = 61,
    family = "Helvetica",
    size = ggtext_size
  ) +
  facet_wrap(~medium, nrow = 3)

# pdf("media_curves.pdf", width = 2, height = 4)
# curves_media
# dev.off()
```

#### Hydration influence

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}
medium_data <- read_csv("/data/PhD/IRX/Bending/WT_medium.csv") %>%
  pivot_longer(cols = c(stiffness, strength), values_to = "value", names_to = "variable") %>%
  mutate(variable = recode(variable,
    "stiffness" = "Stiffness",
    "strength" = "Strength"
  )) %>%
  mutate(medium = ordered(medium, levels = c("Water", "Air", "Sorbitol")))

medium_diameter_line <- ggplot(medium_data) +
  geom_point(aes(x = d.microscope, y = value, fill = medium),
    shape = 21,
    alpha = 0.75,
    size = 1.5,
    stroke = 0.2
  ) +
  geom_smooth(aes(x = d.microscope, y = value, fill = medium, linetype = medium),
    method = "lm",
    colour = "black",
    size = 0.2
  ) +
  labs(
    y = "MPa",
    x = "Stem diameter [mm]"
  ) +
  scale_fill_grey(start = 0.9, end = 0.1) +
  theme_leo() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  ) +
  facet_wrap(~variable, scales = "free_y")

letters <- letter_groups(medium_data %>% filter(variable == "Stiffness"),
  d.microscope,
  medium,
  "tukey",
  print_position = "below",
  print_adjust = 1
)

medium_diameter_box <- ggplot(
  data = medium_data %>% filter(variable == "Stiffness"),
  aes(x = medium, y = d.microscope)
) +
  # geom_line(aes(group = interaction(cell.type, variable, cell, measurement)),
  #           size = 0.2,
  #           alpha = 0.25) +
  geom_quasirandom(
    aes(fill = medium),
    dodge.width = 0.6,
    shape = 21,
    # width = 0.1,
    # colour = NA,
    alpha = 0.75,
    size = 1.5,
    stroke = 0.2
  ) +
  # geom_boxplot(
  #   aes(group = interaction(cell.type, variable), fill = cell.type),
  #   # fill = "white",
  #   alpha = 0.5,
  #   width = 0.6,
  #   position = position_dodge2(width = 1),
  #   outlier.alpha = 0,
  #   lwd = 0.25,
  #   fatten = 1,
  #   width = 0.25
  # ) +
  geom_violin(
    # aes(group = interaction(days, cell.type),
    # fill = cell.type
    # ),
    draw_quantiles = 0.5,
    fill = "white",
    colour = "black",
    alpha = 0.85,
    width = 0.2,
    position = position_dodge(width = 0.6),
    size = 0.2,
    scale = "width"
  ) +
  geom_text(
    data = letters,
    aes(label = Letters),
    size = ggtext_size,
    family = "Helvetica",
    position = position_dodge(width = 0.6)
  ) +
  # scale_x_discrete(labels = c("PX", "PMX", "SMX")) +
  # scale_fill_manual(values = pal_flame_disc) +
  scale_fill_grey(start = 0.9, end = 0.1) +
  expand_limits(y = 0) +
  scale_y_continuous(expand = expand_scale(mult = c(0.1, 0.05))) +
  labs(y = "Stem diameter [mm]") +
  theme_leo() +
  theme(
    text = element_text(family = "Helvetica"),
    legend.position = "none",
    axis.title.x = element_blank()
  )
```

#### Composition influence

```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}
s_data <- read_csv("/data/PhD/IRX/Bending/stiffness_s.csv") %>%
  bind_rows(read_csv("/data/PhD/IRX/Bending/cad_diameters.csv")) %>%
  tidyr::fill(everything()) %>%
  mutate(
    genotype = ordered(genotype, levels = c("WT", "fah1", "cad4xcad5"))
  )

heights <- read_csv("/data/Master/Summer project 16/phenotyping/phenotyping.csv") %>%
  select(genotype, height) %>%
  filter(genotype %in% c("col-0", "fah1", "cad4xcad5")) %>%
  mutate(
    genotype = recode(genotype, "col-0" = "WT"),
    genotype = ordered(genotype, levels = c("WT", "fah1", "cad4xcad5"))
  )

letters <- letter_groups(s_data,
  d.microscope,
  genotype,
  "tukey",
  print_position = "below",
  print_adjust = 1
)


mutant_diameter_box <- ggplot(
  data = s_data,
  aes(x = genotype, y = d.microscope)
) +
  # geom_line(aes(group = interaction(cell.type, variable, cell, measurement)),
  #           size = 0.2,
  #           alpha = 0.25) +
  geom_quasirandom(
    aes(fill = genotype),
    dodge.width = 0.6,
    shape = 21,
    # width = 0.1,
    # colour = NA,
    alpha = 0.75,
    size = 1.5,
    stroke = 0.2
  ) +
  # geom_boxplot(
  #   aes(group = interaction(cell.type, variable), fill = cell.type),
  #   # fill = "white",
  #   alpha = 0.5,
  #   width = 0.6,
  #   position = position_dodge2(width = 1),
  #   outlier.alpha = 0,
  #   lwd = 0.25,
  #   fatten = 1,
  #   width = 0.25
  # ) +
  geom_violin(
    # aes(group = interaction(days, cell.type),
    # fill = cell.type
    # ),
    draw_quantiles = 0.5,
    fill = "white",
    colour = "black",
    alpha = 0.85,
    width = 0.2,
    position = position_dodge(width = 0.6),
    size = 0.2,
    scale = "width"
  ) +
  geom_text(
    data = letters,
    aes(label = Letters),
    size = ggtext_size,
    family = "Helvetica",
    position = position_dodge(width = 0.6)
  ) +
  scale_x_discrete(labels = c(
    "WT",
    expression(italic("fah1")),
    expression(paste(italic("cad4"), "x", italic("cad5")))
  )) +
  # scale_fill_manual(values = pal_flame_disc) +
  scale_fill_manual(values = c(
    pal_ostwald_disc[1],
    pal_ostwald_disc[3],
    pal_ostwald_disc[3]
  )) +
  expand_limits(y = 0) +
  scale_y_continuous(expand = expand_scale(mult = c(0.1, 0.05))) +
  labs(y = "Stem diameter [mm]") +
  theme_leo() +
  theme(
    text = element_text(family = "Helvetica"),
    legend.position = "none",
    axis.title.x = element_blank()
  )

letters <- letter_groups(heights,
  height,
  genotype,
  "tukey",
  print_position = "below",
  print_adjust = 2
)


mutant_height_box <- ggplot(
  data = heights,
  aes(x = genotype, y = height)
) +
  # geom_line(aes(group = interaction(cell.type, variable, cell, measurement)),
  #           size = 0.2,
  #           alpha = 0.25) +
  geom_quasirandom(
    aes(fill = genotype),
    dodge.width = 0.6,
    shape = 21,
    # width = 0.1,
    # colour = NA,
    alpha = 0.75,
    size = 1.5,
    stroke = 0.2
  ) +
  # geom_boxplot(
  #   aes(group = interaction(cell.type, variable), fill = cell.type),
  #   # fill = "white",
  #   alpha = 0.5,
  #   width = 0.6,
  #   position = position_dodge2(width = 1),
  #   outlier.alpha = 0,
  #   lwd = 0.25,
  #   fatten = 1,
  #   width = 0.25
  # ) +
  geom_violin(
    # aes(group = interaction(days, cell.type),
    # fill = cell.type
    # ),
    draw_quantiles = 0.5,
    adjust = 2,
    fill = "white",
    colour = "black",
    alpha = 0.85,
    width = 0.2,
    position = position_dodge(width = 0.6),
    size = 0.2,
    scale = "width"
  ) +
  geom_text(
    data = letters,
    aes(label = Letters),
    size = ggtext_size,
    family = "Helvetica",
    position = position_dodge(width = 0.6)
  ) +
  scale_x_discrete(labels = c(
    "WT",
    expression(italic("fah1")),
    expression(paste(italic("cad4"), "x", italic("cad5")))
  )) +
  # scale_fill_manual(values = pal_flame_disc) +
  scale_fill_manual(values = c(
    pal_ostwald_disc[1],
    pal_ostwald_disc[3],
    pal_ostwald_disc[3]
  )) +
  expand_limits(y = 0) +
  scale_y_continuous(expand = expand_scale(mult = c(0.1, 0.05))) +
  labs(y = "Plant height [cm]") +
  theme_leo() +
  theme(
    text = element_text(family = "Helvetica"),
    legend.position = "none",
    axis.title.x = element_blank()
  )

right <- plot_grid(medium_diameter_box,
  mutant_diameter_box,
  mutant_height_box,
  labels = c("B", "C", "D"),
  align = "v",
  axis = "l",
  label_fontfamily = "Helvetica",
  label_size = 10,
  nrow = 3
)

```

#### Assemble figure
```{r, echo=FALSE, fig.showtext=TRUE, message=FALSE, warning=FALSE, out.width='100%'}

# pdf("bending_supplemental.pdf", width = onecol, height = onecol)
# bending_supp <- plot_grid(
#   curves,
#   right,
#   labels = c("A", ""),
#   # vjust = 1,
#   nrow = 1,
#   align = "v",
#   axis = "lb",
#   label_size = 10,
#   label_fontfamily = "Helvetica"
# )
# bending_supp
# dev.off()

# patch <- curves | (medium_diameter_box / mutant_diameter_box / mutant_height_box)

patch <- 
  curves_media |
  (medium_strength / medium_diameter_box) |
  curves | 
  (violin_strength / mutant_diameter_box / mutant_height_box)

pdf("bending_supplemental.pdf", width = twocol, height = onecol)
patch + plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 10, face = "bold", family = "Helvetica"))
dev.off()
```

### SEM tables

```{r} 
ind_claims <- ind_claims %>%
  select(species, model, fishC, fishP, Independ.Claim, Crit.Value, P.Value)
fit_tab <- kable(ind_claims,
  "latex",
  # align = "lllllcccc",
  # caption = "Raman bands validated by Py--GC/MS.",
  # col.names = c(
  #   "Band no.",
  #   "Band intensity at",
  #   "Normalised to",
  #   "Assigned polymer",
  #   "Assigned structure",
  #   "Pearson's r",
  #   "P-value",
  #   "Pearson's r",
  #   "P-value"
  # ),
  booktabs = TRUE,
  linesep = ""
) %>%
  kable_styling(latex_options = c("hold_position", "striped", "scale_down"))
# add_header_above(c(" " = 5, "Total pyrolysate" = 2, "Total lignin" = 2)) %>%
# save_kable("IRX_tableS1.pdf")

readr::write_file(fit_tab, "sem_fit_tab.txt")

coefs <- coefs %>%
  select(species, model, Response, Predictor, Estimate, Std.Error, Crit.Value, P.Value, Std.Estimate)

coef_tab <- kable(coefs,
  "latex",
  # align = "lllllcccc",
  # caption = "Raman bands validated by Py--GC/MS.",
  # col.names = c(
  #   "Band no.",
  #   "Band intensity at",
  #   "Normalised to",
  #   "Assigned polymer",
  #   "Assigned structure",
  #   "Pearson's r",
  #   "P-value",
  #   "Pearson's r",
  #   "P-value"
  # ),
  booktabs = TRUE,
  linesep = ""
) %>%
  kable_styling(latex_options = c("hold_position", "striped", "scale_down"))
# add_header_above(c(" " = 5, "Total pyrolysate" = 2, "Total lignin" = 2)) %>%
# save_kable("IRX_tableS2.pdf")

readr::write_file(coef_tab, "sem_coef_tab.txt")
```

### Drought experiment conditions

```{r} 
transpiration <- read_csv(
  "/data/PhD/IRX/Drought/evapotranspiration.csv"
) %>%
  pivot_longer(-c(medium, time), names_to = "plant", values_to = "transp_rate") %>%
  mutate(
    genotype = unlist(str_extract_all(plant, "[:alpha:]+")),
    genotype = ordered(recode(genotype, "CAD" = "cad4x5"),
      levels = c("WT", "cad4x5")
    ),
    replicate = unlist(str_extract_all(plant, "[:digit:]+")),
    medium = ordered(medium,
      levels = c("Water", "10% PEG", "20% PEG")
    ),
    transp_rate = transp_rate * 1000
  ) %>%
  select(-plant) %>%
  group_by(medium, genotype, time) %>%
  mutate(mean_transp = mean(transp_rate))

conditions <- read_csv(
  "/data/PhD/IRX/Drought/chamber_conditions.csv"
) %>%
  fill(everything()) %>%
  group_by(genotype) %>%
  mutate(
    genotype = ordered(recode(genotype, "CAD" = "cad4x5"),
      levels = c("WT", "cad4x5")
    ),
    time = row_number()
  ) %>%
  pivot_longer(-c(time, genotype))


transp <- ggplot(
  transpiration,
  aes(
    x = time,
    y = transp_rate,
    linetype = medium,
    colour = genotype
  )
) +
  geom_line(aes(group = interaction(medium, replicate)),
    size = 0.1,
    alpha = 0.25
  ) +
  # geom_line(aes(y = mean_transp)) +
  geom_smooth(aes(
    group = medium,
    # fill = medium
  ),
  size = 0.2,
  se = T
  ) +
  labs(
    y = "Evapotranspiration [mg min<sup>-1</sup>]",
    x = "Time [min]"
  ) +
  expand_limits(x = 0) +
  theme_leo() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.title.y = ggtext::element_textbox_simple(
      orientation = "left-rotated",
      halign = 0.5,
      vjust = 0
    ),
    strip.text = element_blank()
  ) +
  scale_x_continuous(breaks = c(0, 5, 10, 15, 20)) +
  scale_colour_brewer(palette = "Set2") +
  # scale_fill_brewer(palette = "Set2") +
  scale_colour_manual(values = pal_ostwald_disc[c(1,3)]) +
  # scale_colour_grey(start = 0.9, end = 0.1) +
  guides(colour = "none",
         linetype = guide_legend(override.aes = list(colour = "black",
                                                     fill = "white",
                                                     size = 0.4))) +
  facet_wrap(~genotype,
    ncol = 2
  )

hum <- ggplot(conditions %>% filter(name == "humidity"),
                aes(x = time,
                    y = value,
                    colour = genotype)) +
  geom_line(size = 0.2) +
  expand_limits(y = 0) +
  theme_leo() +
  labs(y = "Humidity [%]") +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = ggtext::element_textbox_simple(
          orientation = "left-rotated",
          halign = 0.5,
          vjust = 0
          )) +
  scale_colour_manual(values = pal_ostwald_disc[c(1,3)]) +
  facet_wrap(~genotype, ncol = 2)

ill <- ggplot(conditions %>% filter(name == "illumination"),
                aes(x = time,
                    y = value,
                    colour = genotype)) +
  geom_line(size = 0.2) +
  expand_limits(y = 0) +
  theme_leo() +
  labs(y = "Illumination [cm<sup>-2</sup> s<sup>-1</sup>]") +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = ggtext::element_textbox_simple(
          orientation = "left-rotated",
          halign = 0.5,
          vjust = 0
          ),
        strip.text = element_blank()) +
  scale_colour_manual(values = pal_ostwald_disc[c(1,3)]) +
  facet_wrap(~genotype, ncol = 2)

temp <- ggplot(conditions %>% filter(name == "temperature"),
                aes(x = time,
                    y = value,
                    colour = genotype)) +
  geom_line(size = 0.2) +
  expand_limits(y = 0) +
  theme_leo() +
  labs(y = "Temperature [°C]") +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = ggtext::element_textbox_simple(
          orientation = "left-rotated",
          halign = 0.5,
          vjust = 0
          ),
        strip.text = element_blank()) +
  scale_colour_manual(values = pal_ostwald_disc[c(1,3)]) +
  facet_wrap(~genotype, ncol = 2)


pdf("transpiration_supplement.pdf", width = twocol, height = onecol * 0.9)
hum / ill / temp / transp + plot_layout(heights = c(0.5, 0.5, 0.5 , 1))
dev.off()

```

