---
title: "C4H field trials"
author: "Leonard Blaschek"
date: "11/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(showtext)
library(tidyverse)
library(broom)
library(lubridate)
library(patchwork)

#### import Helvetica ####
font_add(
  "Helvetica",
  regular = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-Lt.otf",
  italic = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-LtIt.otf",
  bold = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-Bd.otf",
  bolditalic = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-BdIt.otf"
)
showtext_auto()

#### generating plot theme ####
theme_leo <- function(base_size = 6,
                      base_family = "Helvetica") {
  theme_minimal(
    base_size = base_size,
    base_family = base_family
  ) %+replace%
    theme(
      strip.text = element_text(hjust = 0, face = "italic"),
      axis.ticks = element_line(
        size = 0.125,
        lineend = "square",
        color = "black"
      ),
      axis.text.x = element_text(
        size = 6,
        colour = "black", # flipped coords
        margin = margin(1, 1, 1, 1)
      ),
      axis.text.y = element_text(
        colour = "black",
        size = 6,
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1)
      ),
      axis.title = element_text(
        colour = "black",
        size = 6
      ),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", size = 0.25),
      axis.line = element_line(size = 0.2),
      panel.spacing = unit(1.5, "mm"),
      legend.position = "none",
      legend.text = element_text(size = 6),
      legend.key.height = unit(4, "mm"),
      plot.title = element_text(
        size = 6,
        hjust = 0
      ),
      complete = TRUE
    )
}

pal_ostwald_disc <- c(
  "#275d95",
  "#e8c245",
  "#d25952"
)

pal_flame <- c(
  "#1F265D",
  # "#203F8F",
  "#404678",
  "#897B88",
  "#E7BBA2",
  "#FFEDC3",
  "#FEFEFE"
)

ggtext_size <- 6 / (14 / 5)
cm_size <- function(x) x / 2.54
twocol <- 18 / 2.54
onehalfcol <- 14 / 2.54
onecol <- 9 / 2.54
```

## Growth data

### Load data

```{r}

#growth instance 1
data1 <- read_csv("/home/leonard/Dropbox/2020_P_manuscript/P_data_1.csv") %>%
  pivot_longer(cols = -sample, names_to = "ID", values_to = "value") %>%
  separate(sample, into = c("genotype", "plant")) %>%
  separate(ID, into = c("variable", "date"), sep = " ") %>%
  mutate(instance = "1")

#growth instance 2
data2 <- read_csv("/home/leonard/Dropbox/2020_P_manuscript/P_data_2.csv") %>%
  pivot_longer(cols = -sample, names_to = "ID", values_to = "value") %>%
  separate(sample, into = c("genotype", "plant")) %>%
  separate(ID, into = c("variable", "date"), sep = " ") %>%
  mutate(instance = "2")

data <- data1 %>%
  bind_rows(data2) %>%
  mutate(date = as_date(parse_date_time(date, "ym")))
```

### View growth kinetics

```{r}
ggplot(data %>% filter(variable == "height")) +
  geom_line(aes(x = date, y = value, colour = genotype, group = interaction(genotype, plant)),
    alpha = 0.1
  ) +
  geom_smooth(aes(x = date, y = value, colour = genotype),
    method = "loess",
    span = 0.5
  ) +
  theme_leo() +
  theme(legend.position = "bottom") +
  facet_wrap(~instance, nrow = 2)
```

### Split plant data into growth periods

```{r}
year_height <- data %>%
  mutate(
    year = isoyear(date),
    month = month(date)
  ) %>%
  group_by(instance, year, variable, genotype, plant) %>%
  # identify the last measurement of each year
  mutate(
    "season" = case_when(
    date == max(date) ~ "endofyear",
    TRUE ~ "midyear")
    ) %>%
  # filter out last measurements that were taken before the end of the growth period
  filter(season == "endofyear" & variable == "height" & month > 9) %>%
  group_by(instance, genotype, plant) %>%
  # calculate plant age at time of measurement
  mutate(age = year - min(year)) %>%
  arrange(date, .by_group = TRUE) %>%
  # calculate the yearly size change and convert from mm to cm
  mutate(height_growth = (value - lag(value, default = first(value))) / 10) %>%
  # filter out the starting years
  filter(year != min(year)) %>%
  ungroup() %>%
  select(genotype, plant, instance, year, age, height_growth)

year_width <- data %>%
  mutate(
    year = isoyear(date),
    month = month(date)
  ) %>%
  group_by(instance, year, variable, genotype, plant) %>%
  # identify the last measurement of each year
  mutate("season" = case_when(
    date == max(date) ~ "endofyear",
    TRUE ~ "midyear"
  )) %>%
  # filter out last measurements that were taken before the end of the growth period
  filter(season == "endofyear" & variable == "width" & month > 9) %>%
  group_by(instance, genotype, plant) %>%
  arrange(date, .by_group = TRUE) %>%
  # calculate the yearly size change and convert from mm to cm
  mutate(width_growth = (value - lag(value, default = first(value))) / 10) %>%
  # filter out the starting years
  filter(year != min(year)) %>%
  ungroup() %>%
  select(genotype, plant, instance, year, width_growth)

year_injuries <- data %>%
  mutate(
    year = isoyear(date),
    month = month(date)
  ) %>%
  group_by(instance, year, variable, genotype, plant) %>%
  filter(variable == "injuries") %>%
  group_by(instance, genotype, plant, year) %>%
  # summarise yearly injuries as the worst injuries sustained at any time in the year
  summarise(injuries = max(value)) %>%
  # filter out the starting years
  filter(year != min(year)) %>%
  ungroup() %>%
  select(genotype, plant, instance, year, injuries)

year_survival <- data %>%
  mutate(
    year = isoyear(date),
    month = month(date)
  ) %>%
  group_by(instance, year, variable, genotype, plant) %>%
  filter(variable == "survival") %>%
  group_by(instance, genotype, plant, year) %>%
  summarise(survival = min(value)) %>%
  # identify plants that died in the respective or previous years
  filter(survival == 0) %>%
  ungroup() %>%
  select(genotype, plant, instance, year, survival)

year_data <- year_height %>%
  full_join(year_width, by = c("genotype", "plant", "instance", "year")) %>%
  full_join(year_injuries, by = c("genotype", "plant", "instance", "year")) %>%
  # remove all data after plant death
  anti_join(year_survival, by = c("genotype", "plant", "instance", "year")) %>%
  ungroup() %>%
  mutate(
    line = genotype,
    genotype = case_when(
      genotype == "WT" ~ "WT",
      TRUE ~ "<i>C4H</i>-RNAi"
    )
  )
```

### View yearly growth by genotype

```{r}
year_growth <- ggplot(
  year_data,
  aes(x = genotype, y = height_growth)
) +
  ggbeeswarm::geom_quasirandom(
    aes(fill = age),
    shape = 21,
    stroke = 0.2,
    size = 1,
    alpha = 0.5,
    width = 0.4
  ) +
  geom_violin(
    # aes(group = interaction(instance,genotype)),
    draw_quantiles = 0.5,
    colour = "black",
    alpha = 0.5,
    size = 0.2,
    width = 0.4
  ) +
  scale_fill_gradientn(colours = pal_flame,
                       guide = guide_legend(
                         direction = "horizontal",
                         title.position = "top",
                         title = "Plant age [y]"
                         )) +
  labs(y = "Yearly height change [cm]") +
  theme_leo() +
  theme(legend.position = c(0.76,0.92),
        legend.key.height = unit(1, "mm"),
        legend.key.width = unit(1, "mm"),
        legend.title.align = 0,
        axis.text.x = ggtext::element_markdown(),
        axis.title.x = element_blank()) +
  facet_wrap(~instance)

pdf("year_growth.pdf", height = onecol * 0.5, width = onecol * 0.5)
year_growth
dev.off()
```

## Environmental data

### Load data

```{r}
envir <- read_csv("/home/leonard/Dropbox/2020_P_manuscript/env_data.csv") %>%
  mutate(
    date = as_date(parse_date_time(date, "ym")),
    year = isoyear(date),
    month = month(date)
  )

```

### View temperature and humidty timelines

```{r}

temps <- ggplot(envir) +
  geom_ribbon(aes(x = date, ymin = low_temp, ymax = high_temp)) +
  geom_line(aes(x = date, y = mean_temp)) +
  scale_x_date(limits = c(as_date(parse_date_time("2011-06-01", "ymd")),
                          as_date(parse_date_time("2018-01-01", "ymd"))),
               breaks = "12 month",
               minor_breaks = "3 month") +
  theme_leo() +
  theme(panel.grid.major.x = element_line(colour = "grey50"),
        panel.grid.minor.x = element_line(colour = "grey50"))
  

humid <- ggplot(envir) +
  geom_ribbon(aes(x = date, ymin = low_humidity, ymax = high_humidity)) +
  geom_line(aes(x = date, y = mean_humidity)) +
  scale_x_date(limits = c(as_date(parse_date_time("2011-06-01", "ymd")),
                          as_date(parse_date_time("2018-01-01", "ymd"))),
               breaks = "12 month",
               minor_breaks = "3 month") +
  theme_leo() +
  theme(panel.grid.major.x = element_line(colour = "grey50"),
        panel.grid.minor.x = element_line(colour = "grey50"))
  
temps / humid

```

### Summarise yearly conditions

```{r}

year_envir <- envir %>%
  group_by(year) %>%
  summarise(
    #temperature
    lowest_temp = min(low_temp),
    highest_temp = max(high_temp),
    average_high_temp = mean(high_temp),
    average_low_temp = mean(low_temp),
    average_temp = mean(mean_temp),
    #summer temperature
    lowest_summer_temp = min(low_temp[month %in% c(6:8)]),
    highest_summer_temp = max(high_temp[month %in% c(6:8)]),
    average_summer_high_temp = mean(high_temp[month %in% c(6:8)]),
    average_summer_low_temp = mean(low_temp[month %in% c(6:8)]),
    average_summer_temp = mean(mean_temp[month %in% c(6:8)]),
    #winter temperature
    lowest_winter_temp = min(low_temp[month %in% c(11, 12, 1, 2, 3, 4)]),
    highest_winter_temp = max(high_temp[month %in% c(11, 12, 1, 2, 3, 4)]),
    average_winter_high_temp = mean(high_temp[month %in% c(11, 12, 1, 2, 3, 4)]),
    average_winter_low_temp = mean(low_temp[month %in% c(11, 12, 1, 2, 3, 4)]),
    average_winter_temp = mean(mean_temp[month %in% c(11, 12, 1, 2, 3, 4)]),
    #humidity
    average_high_humidity = mean(high_humidity),
    average_low_humidity = mean(low_humidity),
    average_humidity = mean(mean_humidity),
    lowest_humidity = min(low_humidity),
    highest_humidity = max(high_humidity),
    #precipitation
    average_precipitation = mean(mean_precip),
    lowest_precipitation = min(total_precip),
    #wind
    highest_wind = max(wind),
    average_wind = mean(wind),
    #light
    highest_light = max(irradiation),
    average_light = mean(irradiation)
  )

cumul_temp <- read_csv("/home/leonard/Dropbox/2020_P_manuscript/cumul_temp.csv") %>%
  mutate(
    date = as_date(parse_date_time(date, "ym")),
    year = isoyear(date),
    month = month(date)
  )

year_cumul_temp <- cumul_temp %>%
  group_by(year) %>%
  summarise(
    may_0deg = `cummul h above 0`[month == 5],
    may_5deg = `cummul h above 5`[month == 5],
    may_10deg = `cummul h above 10`[month == 5])
```

### Environmental pSEM

```{r}

pdf("c4h_envir_correlations.pdf", width = 10, height = 10)
PerformanceAnalytics::chart.Correlation(envir[,-c(1,12, 13)])
dev.off()

envir_psem <- piecewiseSEM::psem(
  lm(mean_humidity ~
       high_temp +
       irradiation +
       wind +
       mean_precip,
     data = envir),
  lm(high_temp ~
       irradiation +
       wind +
       mean_precip,
     data = envir)
)

plot(envir_psem)
env_coefs <- summary(envir_psem)$coefficients
env_model <- summary(envir_psem)$IC
```

## Export merged data

```{r}
export_data <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  mutate(genotype = case_when(
      genotype == "WT" ~ "WT",
      TRUE ~ "C4H-RNAi"
    ))

write_csv(export_data, "yearly_data.csv")
```

## Model optimisation

```{r}

#### WT model ####
optim_data <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  filter(genotype == "WT") %>%
  select(-c(genotype:year, line), 
         -highest_light, -average_light
         ) %>%
  #scale and centre data
  mutate(across(where(is.numeric), ~ scale(.x))) %>%
  drop_na(height_growth)

# Fit the full model 
full_model <- lm(height_growth ~., data = optim_data %>% select(-width_growth))
# Stepwise regression model
step_model_WT <- MASS::stepAIC(full_model, direction = "both", 
                      trace = FALSE)
summary(step_model_WT)

#### C4H model ####
optim_data <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  filter(genotype == "<i>C4H</i>-RNAi") %>%
  select(-c(genotype:year, line), 
         -highest_light, -average_light
         ) %>%
  #scale and centre data
  mutate(across(where(is.numeric), ~ scale(.x))) %>%
  drop_na(height_growth)

# Fit the full model 
full_model <- lm(height_growth ~., data = optim_data %>% select(-width_growth))
# Stepwise regression model
step_model_c4h <- MASS::stepAIC(full_model, direction = "both", 
                      trace = FALSE)
summary(step_model_c4h)

#### Models for individual mutant lines ####
optim_data <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  filter(line == "1B") %>%
  select(-c(genotype:year, line), 
         -highest_light, -average_light
         ) %>%
  #scale and centre data
  mutate(across(where(is.numeric), ~ scale(.x))) %>%
  drop_na(height_growth)

# Fit the full model 
full_model <- lm(height_growth ~., data = optim_data %>% select(-width_growth))
# Stepwise regression model
step_model_1b <- MASS::stepAIC(full_model, direction = "both", 
                      trace = FALSE)
summary(step_model_1b)

optim_data <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  filter(line == "3B") %>%
  select(-c(genotype:year, line), 
         -highest_light, -average_light
         ) %>%
  #scale and centre data
  mutate(across(where(is.numeric), ~ scale(.x))) %>%
  drop_na(height_growth)

# Fit the full model 
full_model <- lm(height_growth ~., data = optim_data %>% select(-width_growth))
# Stepwise regression model
step_model_3b <- MASS::stepAIC(full_model, direction = "both", 
                      trace = FALSE)
summary(step_model_3b)

```

### Export data

```{r}
c4h_coefficients <- 
  tidy(step_model_c4h) %>%
  mutate(genotype = "C4H-RNAi",
         line = "1B + 3B") %>%
  add_row(tidy(step_model_1b) %>% mutate(genotype = "C4H-RNAi",
                                         line = "1B")) %>%
  add_row(tidy(step_model_3b) %>% mutate(genotype = "C4H-RNAi",
                                         line = "3B")) %>%
  add_row(tidy(step_model_WT) %>% mutate(genotype = "WT",
                                         line = "T89"))

c4h_models <- 
  glance(step_model_c4h) %>%
  mutate(genotype = "C4H-RNAi",
         line = "1B + 3B") %>%
  add_row(glance(step_model_1b) %>% mutate(genotype = "C4H-RNAi",
                                         line = "1B")) %>%
  add_row(glance(step_model_3b) %>% mutate(genotype = "C4H-RNAi",
                                         line = "3B")) %>%
  add_row(glance(step_model_WT) %>% mutate(genotype = "WT",
                                         line = "T89"))

pdf("model_comparisons.pdf", width = onecol, height = onecol)
ggplot(
  c4h_coefficients %>%
    filter(genotype == "C4H-RNAi" & term != "(Intercept)"),
  aes(
    x = term,
    y = estimate,
    fill = line
  )
) +
  geom_point(
    shape = 21,
    size = 4
  ) +
  scale_fill_brewer(palette = "RdYlBu") +
  theme_leo() +
  theme(legend.position = "bottom") +
  coord_flip()
dev.off()

writexl::write_xlsx(list(
  "growth models" = c4h_models,
  "growth coefficients" = c4h_coefficients,
  "environmental model" = env_model,
  "environmental coefficients" = env_coefs
),
path = "Serk2020_tables.xlsx",
col_names = TRUE,
format_headers = TRUE,
use_zip64 = FALSE
)

```

## Partial regression and partial residual plots

```{r}

#### height_growth ~ age ####
partial_model <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  drop_na(height_growth)

pdf("age_growth.pdf", height = onecol * 0.5, width = onecol * 0.5)
ggplot(partial_model,
       aes(
         x = age,
         y = height_growth
         )
       ) +
  geom_point(shape = 16, alpha = 0.5, size = 0.5) +
  geom_smooth(span = 1.5, size = 0.1) +
  theme_leo() +
  theme(strip.text = ggtext::element_textbox_simple()) +
  facet_wrap(~ genotype)
dev.off()

#### added-variable plot (partial regression plot) for age ####
partial_model <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  drop_na(height_growth) %>%
  nest(data = -c(genotype)) %>%
  mutate(
    growth_model = map(data, ~ lm(height_growth ~
    injuries +
      lowest_temp +
      highest_temp +
      average_high_temp +
      average_low_temp,
    data = .x
    )),
    age_model = map(data, ~ lm(age ~
    injuries +
      lowest_temp +
      highest_temp +
      average_high_temp +
      average_low_temp,
    data = .x
    )),
    growth_model_resid = map(growth_model, residuals),
    age_model_resid = map(age_model, residuals),
  ) %>%
  select(genotype, growth_model_resid, age_model_resid) %>%
  unnest(c(growth_model_resid, age_model_resid),
         names_repair = "unique")

pdf("age_partial_regression.pdf", height = onecol * 0.5, width = onecol * 0.5)
ggplot(partial_model,
       aes(
         x = age_model_resid,
         y = growth_model_resid
         )
       ) +
  labs(x = "Plant age | other predictors",
       y = "Plant Growth | other predictors") +
  geom_point(shape = 16, alpha = 0.5, size = 0.5) +
  geom_smooth(span = 1.5, size = 0.1) +
  theme_leo() +
  theme(strip.text = ggtext::element_textbox_simple()) +
  facet_wrap(~ genotype)
dev.off()

#### partial residual plot for age ####
partial_model <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  drop_na(height_growth) %>%
  nest(data = -c(genotype)) %>%
  mutate(
    model = map(data, ~ lm(height_growth ~
    injuries +
      age +
      lowest_temp +
      highest_temp +
      average_high_temp +
      average_low_temp,
    data = .x
    )),
    resid = map(model, residuals),
    age = map(data, "age"),
    age_coef = map(model, ~{coef(.x)["age"]})
  ) %>%
  select(genotype, resid, age, age_coef) %>%
  unnest(c(resid, age, age_coef),
         names_repair = "unique") %>%
  mutate(partial_resid = resid + age * age_coef)

pdf("age_partial_residuals.pdf", height = onecol * 0.5, width = onecol * 0.5)
ggplot(partial_model,
       aes(
         x = age,
         y = partial_resid
         )
       ) +
  # geom_line(aes(y = age * age_coef)) +
  labs(x = "Plant age [y]",
       y = "Partial residuals of height change [cm]") +
  geom_point(shape = 16, alpha = 0.5, size = 0.5) +
  geom_smooth(span = 1.5, size = 0.1) +
  theme_leo() +
  theme(strip.text = ggtext::element_textbox_simple()) +
  facet_wrap(~ genotype)
dev.off()
```

### Export data

```{r}
export_data <- partial_model %>%
  mutate(genotype = case_when(
      genotype == "WT" ~ "WT",
      TRUE ~ "C4H-RNAi"
    )) %>%
  arrange(genotype, age)

write_csv(export_data, "partial_residuals.csv")
```

## Check C4H-RNAi line 2B

### Load data

```{r}

#growth instance 1
data2B <- read_csv("/home/leonard/Dropbox/2020_P_manuscript/P_data_2B.csv") %>%
  pivot_longer(cols = -sample, names_to = "ID", values_to = "value") %>%
  separate(sample, into = c("genotype", "plant")) %>%
  separate(ID, into = c("variable", "date"), sep = " ") %>%
  mutate(instance = "1")

data <- data2B %>%
  mutate(date = as_date(parse_date_time(date, "ym")))
```

### Split plant data into growth periods

```{r}
year_height <- data %>%
  mutate(
    year = isoyear(date),
    month = month(date)
  ) %>%
  group_by(instance, year, variable, genotype, plant) %>%
  # identify the last measurement of each year
  mutate(
    "season" = case_when(
    date == max(date) ~ "endofyear",
    TRUE ~ "midyear")
    ) %>%
  # filter out last measurements that were taken before the end of the growth period
  filter(season == "endofyear" & variable == "height" & month > 9) %>%
  group_by(instance, genotype, plant) %>%
  # calculate plant age at time of measurement
  mutate(age = year - min(year)) %>%
  arrange(date, .by_group = TRUE) %>%
  # calculate the yearly size change and convert from mm to cm
  mutate(height_growth = (value - lag(value, default = first(value))) / 10) %>%
  # filter out the starting years
  filter(year != min(year)) %>%
  ungroup() %>%
  select(genotype, plant, instance, year, age, height_growth)

year_width <- data %>%
  mutate(
    year = isoyear(date),
    month = month(date)
  ) %>%
  group_by(instance, year, variable, genotype, plant) %>%
  # identify the last measurement of each year
  mutate("season" = case_when(
    date == max(date) ~ "endofyear",
    TRUE ~ "midyear"
  )) %>%
  # filter out last measurements that were taken before the end of the growth period
  filter(season == "endofyear" & variable == "width" & month > 9) %>%
  group_by(instance, genotype, plant) %>%
  arrange(date, .by_group = TRUE) %>%
  # calculate the yearly size change and convert from mm to cm
  mutate(width_growth = (value - lag(value, default = first(value))) / 10) %>%
  # filter out the starting years
  filter(year != min(year)) %>%
  ungroup() %>%
  select(genotype, plant, instance, year, width_growth)

year_injuries <- data %>%
  mutate(
    year = isoyear(date),
    month = month(date)
  ) %>%
  group_by(instance, year, variable, genotype, plant) %>%
  filter(variable == "injuries") %>%
  group_by(instance, genotype, plant, year) %>%
  # summarise yearly injuries as the worst injuries sustained at any time in the year
  summarise(injuries = max(value)) %>%
  # filter out the starting years
  filter(year != min(year)) %>%
  ungroup() %>%
  select(genotype, plant, instance, year, injuries)

year_survival <- data %>%
  mutate(
    year = isoyear(date),
    month = month(date)
  ) %>%
  group_by(instance, year, variable, genotype, plant) %>%
  filter(variable == "survival") %>%
  group_by(instance, genotype, plant, year) %>%
  summarise(survival = min(value)) %>%
  # identify plants that died in the respective or previous years
  filter(survival == 0) %>%
  ungroup() %>%
  select(genotype, plant, instance, year, survival)

year_data <- year_height %>%
  full_join(year_width, by = c("genotype", "plant", "instance", "year")) %>%
  full_join(year_injuries, by = c("genotype", "plant", "instance", "year")) %>%
  # remove all data after plant death
  anti_join(year_survival, by = c("genotype", "plant", "instance", "year")) %>%
  ungroup() %>%
  mutate(
    line = genotype,
    genotype = case_when(
      genotype == "WT" ~ "WT",
      TRUE ~ "<i>C4H</i>-RNAi"
    )
  )
```

### Model optimisation

```{r}

#### WT model ####
optim_data <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  # filter(genotype == "WT") %>%
  select(-c(genotype:year, line), 
         -highest_light, -average_light
         ) %>%
  #scale and centre data
  mutate(across(where(is.numeric), ~ scale(.x))) %>%
  drop_na(height_growth)

# Fit the full model 
full_model <- lm(height_growth ~., data = optim_data %>% select(-width_growth))
# Stepwise regression model
step_model_2B <- MASS::stepAIC(full_model, direction = "both", 
                      trace = FALSE)

summary(step_model_2B)

```

### Partial regression and partial residual plots

```{r}

#### height_growth ~ age ####
partial_model <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  drop_na(height_growth)

pdf("age_growth.pdf", height = onecol * 0.5, width = onecol * 0.5)
ggplot(partial_model,
       aes(
         x = age,
         y = height_growth
         )
       ) +
  geom_point(shape = 16, alpha = 0.5, size = 0.5) +
  geom_smooth(span = 1.5, size = 0.1) +
  theme_leo() +
  theme(strip.text = ggtext::element_textbox_simple()) +
  facet_wrap(~ genotype)
dev.off()

#### added-variable plot (partial regression plot) for age ####
partial_model <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  drop_na(height_growth) %>%
  nest(data = -c(genotype)) %>%
  mutate(
    growth_model = map(data, ~ lm(height_growth ~
    injuries +
      lowest_temp +
      highest_temp +
      average_high_temp +
      average_low_temp,
    data = .x
    )),
    age_model = map(data, ~ lm(age ~
    injuries +
      lowest_temp +
      highest_temp +
      average_high_temp +
      average_low_temp,
    data = .x
    )),
    growth_model_resid = map(growth_model, residuals),
    age_model_resid = map(age_model, residuals),
  ) %>%
  select(genotype, growth_model_resid, age_model_resid) %>%
  unnest(c(growth_model_resid, age_model_resid),
         names_repair = "unique")

pdf("age_partial_regression.pdf", height = onecol * 0.5, width = onecol * 0.5)
ggplot(partial_model,
       aes(
         x = age_model_resid,
         y = growth_model_resid
         )
       ) +
  labs(x = "Plant age | other predictors",
       y = "Plant Growth | other predictors") +
  geom_point(shape = 16, alpha = 0.5, size = 0.5) +
  geom_smooth(span = 1.5, size = 0.1) +
  theme_leo() +
  theme(strip.text = ggtext::element_textbox_simple()) +
  facet_wrap(~ genotype)
dev.off()

#### partial residual plot for age ####
partial_model <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year") %>%
  ungroup() %>%
  drop_na(height_growth) %>%
  nest(data = -c(genotype)) %>%
  mutate(
    model = map(data, ~ lm(height_growth ~
    injuries +
      age +
      lowest_temp +
      highest_temp +
      average_high_temp +
      average_low_temp,
    data = .x
    )),
    resid = map(model, residuals),
    age = map(data, "age"),
    age_coef = map(model, ~{coef(.x)["age"]})
  ) %>%
  select(genotype, resid, age, age_coef) %>%
  unnest(c(resid, age, age_coef),
         names_repair = "unique") %>%
  mutate(partial_resid = resid + age * age_coef)

pdf("age_partial_residuals.pdf", height = onecol * 0.5, width = onecol * 0.5)
ggplot(partial_model,
       aes(
         x = age,
         y = partial_resid
         )
       ) +
  # geom_line(aes(y = age * age_coef)) +
  labs(x = "Plant age [y]",
       y = "Partial residuals of height change [cm]") +
  geom_point(shape = 16, alpha = 0.5, size = 0.5) +
  geom_smooth(span = 1.5, size = 0.1) +
  theme_leo() +
  theme(strip.text = ggtext::element_textbox_simple()) +
  facet_wrap(~ genotype)
dev.off()
```

#### Export data

```{r}

export_data <- year_data %>%
  left_join(year_envir, by = "year") %>%
  left_join(year_cumul_temp, by = "year")

write_csv(export_data, "yearly_data_2B.csv")

c4h_coefficients <- 
  tidy(step_model_2B) %>%
  mutate(genotype = "C4H-RNAi",
         line = "2B")

write_csv(c4h_coefficients, "model_coefficients_2B.csv")

c4h_models <- 
  glance(step_model_2B) %>%
  mutate(genotype = "C4H-RNAi",
         line = "2B")

write_csv(c4h_models, "model_statistics_2B.csv")

export_residuals <- partial_model %>%
  mutate(genotype = case_when(
      genotype == "WT" ~ "WT",
      TRUE ~ "C4H-RNAi"
    )) %>%
  arrange(genotype, age)

write_csv(export_residuals, "partial_residuals_2B.csv")
```
