---
title: "Laccase Phenotyping 2020"
author: "Leonard Blaschek"
date: "16/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(qrencoder)
library(cowplot)
library(grid)
library(gridExtra)
library(tidyverse)

cm_size <- function(x) x / 2.54
```

## Experimental plan

This phenotyping experiment will hopefully create the basis for the coming laccase papers.

### Planting & growth

Planting:

* 3 independent growth instances in Jan-Mar, Mar-May and May-Jul
* 6 plants per genotype per growth instance of the triples, quadruples and quintuple 
  + of this, 1 plant is to be harvested at 30 cm stem height and ground for RNA and possibly bulk activity
* 2/3 of soil to 1/3 of vermiculite
* Pest treatment with Bt-toxin, to avoid the weird phenotypes associated with (excessive) nematode treatment

Growth conditions:

* 23:00-06:00
  + 18&deg;C
  + 50% humidity 
  + 0 µmol m^-2^ s^-1^ irradiance
* 06:00-07:00
  + 20&deg;C
  + 50% humidity 
  + 60 µmol m^-2^ s^-1^ irradiance
* 07:00-22:00
  + 22&deg;C
  + 50% humidity 
  + 100 µmol m^-2^ s^-1^ irradiance
* 22:00-23:00
  + 20&deg;C
  + 50% humidity 
  + 60 µmol m^-2^ s^-1^ irradiance
  

  
### Phenotyping

Singling out should be done ~1 day after germination. After that, the trays will be imaged every day from the top. Colour card central on the right side, tray QR-code on the top right side.

Camera settings (prel.):

* 35 mm objective
* 1/60 s exposure time
* f16 aperture
* 12800 ISO sensitivity

After bolting, plants should be imaged every day from the side. Colour card on the right side, level with the pot. QR-code taped to the pot. Blue plastic stick for support (ordered today). I might do images from 2 sides, with two QR-codes specifying the angle.

Camera settings (prel.):

* 35 mm objective
* 1/60 s exposure time
* f16 aperture
* 12800 ISO sensitivity


## QR-code generation

```{r cars}

genotypes <- c("WT", 
               "lac4x5x10x12x17", 
               "lac5x10x12x17",
               "lac4x10x12x17",
               "lac4x5x12x17",
               "lac4x5x10x17",
               "lac4x5x10x12")
replicates <- c(1:5)

plants <- expand_grid("genotype" = genotypes, "replicate" = replicates) %>%
  unite("plant", genotype, replicate)

qrc <- function(plant) {
png(paste0("qr_codes/", plant, ".png"), height = 900, width = 900, units = "px")
par(mar=c(0,0,0,0))
plot1 <- image(qrencode_raster(plant), 
      asp=1, col=c("white", "black"), axes=FALSE, 
      xlab="", ylab="",
      xlim = c(0, 1), ylim = c(0, 1.1))
text(plant, x = 0.5, y = 1.05, cex = 4)
dev.off()
}

apply(plants, 1, qrc)
  
```

