---
title: "Laccase Phenotyping 2020"
author: "Leonard Blaschek"
date: "16/01/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(showtext)
# library(qrencoder)
library(cowplot)
library(grid)
library(gridExtra)
library(tidyverse)
library(ggrepel)
library(ggridges)
library(cowplot)
library(jsonlite)
library(slider)
library(lubridate)

#### import Helvetica ####
font_add(
  "Helvetica",
  regular = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-Lt.otf",
  italic = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-LtIt.otf",
  bold = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-Bd.otf",
  bolditalic = "/prop_fonts/01. Helvetica   [1957 - Max Miedinger]/HelveticaNeueLTStd-BdIt.otf"
)
showtext_auto()

#### generating plot theme ####
theme_leo <- function(base_size = 10,
                      base_family = "Helvetica") {
  theme_minimal(
    base_size = base_size,
    base_family = base_family
  ) %+replace%
    theme(
      strip.text = element_text(hjust = 0, face = "italic"),
      # axis.ticks = element_line(
      #   size = 0.125,
      #   lineend = "square",
      #   color = "black"
      # ),
      axis.text.x = element_text(
        size = 10,
        colour = "black", # flipped coords
        margin = margin(1, 1, 1, 1)
      ),
      axis.text.y = element_text(
        colour = "black",
        size = 10,
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1)
      ),
      axis.title = element_text(
        colour = "black",
        size = 10
      ),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      # panel.border = element_rect(fill = NA, color = "black", size = 0.25),
      panel.spacing = unit(1.5, "mm"),
      legend.position = "none",
      legend.text = element_text(size = 10),
      legend.key.height = unit(4, "mm"),
      plot.title = element_text(
        size = 10,
        hjust = 0
      ),
      complete = TRUE
    )
}

pal_ostwald_disc <- c(
  "#275d95",
  "#e8c245",
  "#d25952"
)

ggtext_size <- 10 / (14 / 5)
cm_size <- function(x) x / 2.54
```

## Experimental plan

This phenotyping experiment will hopefully create the basis for the coming laccase papers.

### Planting & growth

Planting:

* 3 independent growth instances in Jan-Mar, Mar-May and May-Jul
* 6 plants per genotype per growth instance of the triples, quadruples and quintuple 
  + of this, 1 plant is to be harvested at 30 cm stem height and ground for RNA and possibly bulk activity
* 2/3 of soil to 1/3 of vermiculite
* Pest treatment with Bt-toxin, to avoid the weird phenotypes associated with (excessive) nematode treatment

Growth conditions:

* 23:00-06:00
  + 18&deg;C
  + 50% humidity 
  + 0 µmol m^-2^ s^-1^ irradiance
* 06:00-07:00
  + 20&deg;C
  + 50% humidity 
  + 60 µmol m^-2^ s^-1^ irradiance
* 07:00-22:00
  + 22&deg;C
  + 50% humidity 
  + 100 µmol m^-2^ s^-1^ irradiance
* 22:00-23:00
  + 20&deg;C
  + 50% humidity 
  + 60 µmol m^-2^ s^-1^ irradiance
  

  
### Phenotyping

Singling out should be done ~1 day after germination. After that, the trays will be imaged every day from the top. Colour card central on the right side, tray QR-code on the top right side.

Camera settings (prel.):

* 35 mm objective
* 1/60 s exposure time
* f16 aperture
* 12800 ISO sensitivity

After bolting, plants should be imaged every day from the side. Colour card on the right side, level with the pot. QR-code taped to the pot. Blue plastic stick for support (ordered today). I might do images from 2 sides, with two QR-codes specifying the angle.

Camera settings (prel.):

* 35 mm objective
* 1/60 s exposure time
* f16 aperture
* 12800 ISO sensitivity


## QR-code generation

```{r, message=FALSE, results=FALSE, eval=FALSE}

genotypes <- c(
  "WT",
  "lac4x5x10x12x17",
  "lac5x10x12x17",
  "lac4x10x12x17",
  "lac4x5x12x17",
  "lac4x5x10x17",
  "lac4x5x10x12"
)
replicates <- c(1:5)

plants <- expand_grid("genotype" = genotypes, "replicate" = replicates) %>%
  unite("plant", genotype, replicate)

qrc <- function(plant) {
  png(paste0("qr_codes/", plant, ".png"), height = 900, width = 900, units = "px")
  par(mar = c(0, 0, 0, 0))
  plot1 <- image(qrencode_raster(plant),
    asp = 1, col = c("white", "black"), axes = FALSE,
    xlab = "", ylab = "",
    xlim = c(0, 1), ylim = c(0, 1.1)
  )
  text(plant, x = 0.5, y = 1.05, cex = 4)
  dev.off()
}

apply(plants, 1, qrc)
```

## Adjusted camera settings

To avoid saturation in the colour reference card, the following settings have proven optimal:

* 50 mm objective
* 1/125 s exposure time
* f5.6 aperture
* 2000 ISO sensitivity
* Incandescent white balance

## Harvesting 2020-04-16--2020-04-21

```{r message=FALSE}
siliques <- read_csv(
  "/home/leonard/Dropbox/2020-01_LAC_phenotyping/2020-04-16_siliques.csv"
) %>%
  mutate(rep = as.character(rep))

pheno_data <- read_csv(
  "/home/leonard/Dropbox/2020-01_LAC_phenotyping/2020-04-16_harvesting.csv"
) %>%
  mutate(rep = as.character(rep)) %>%
  left_join(siliques %>%
    pivot_longer(-c(genotype, rep), names_to = "sil", values_to = "length") %>%
    group_by(genotype, rep) %>%
    summarise(mean.sil.length = mean(length))) %>%
  group_by(genotype, rep) %>%
  mutate(
    total.sil = mat.sil + green.sil,
    senescence = green.sil / total.sil
  )

pheno_pca <- pheno_data %>%
  ungroup() %>%
  mutate(genotype = recode(genotype,
      "lac4/5/10/12/17" = "Q",
      "lac5/10/12/17" = "Q-4",
      "lac4/10/12/17" = "Q-5",
      "lac4/5/12/17" = "Q-10",
      "lac4/5/10/17" = "Q-12",
      "lac4/5/10/12" = "Q-17"
    )) %>%
  unite("ID", genotype, rep)

pca_result <- prcomp(pheno_pca[, -1], center = TRUE, scale. = TRUE)
rownames(pca_result$x) <- pheno_pca$ID

gg_pca <- as_tibble(pca_result$x, rownames = NA) %>%
  rownames_to_column(var = "ID") %>%
  separate(ID, sep = "_", into = c("genotype", "replicate"))

pca_plot <- ggplot(gg_pca, aes(x = PC1, y = PC2, fill = genotype)) +
  geom_hline(yintercept = 0, linetype = 1, size = 0.2) +
  geom_vline(xintercept = 0, linetype = 1, size = 0.2) +
  stat_ellipse(geom = "polygon", alpha = 0.5, type = "t", level = 0.4) +
  stat_ellipse(aes(fill = NULL), colour = "black", linetype = 2, size = 0.2) +
  geom_point(
    size = 3,
    stroke = 0.4,
    alpha = 0.75,
    shape = 21
  ) +
  geom_text_repel(
    data = filter(gg_pca, replicate == "4"),
    aes(label = genotype),
    min.segment.length = 0,
    family = "Helvetica",
    size = ggtext_size
  ) +
  labs(
    x = "PC 1 (Size and weight)",
    y = "PC 2 (Degree of senescence)"
  ) +
  scale_fill_viridis_d() +
  theme_leo() +
  theme(
    plot.margin = unit(c(2, 2, 3, 2), "mm")
  )

plot_grid(
  factoextra::fviz_contrib(pca_result, choice = "var", axes = 1, top = 5),
  factoextra::fviz_contrib(pca_result, choice = "var", axes = 2, top = 5)
)

pdf("pheno_PCA.pdf", width = cm_size(20), height = cm_size(10))
pca_plot
dev.off()
```

### Image based phenotyping

#### Hue

```{r message=FALSE}
pcv_data <- fromJSON("/home/leonard/Dropbox/2020-01_LAC_phenotyping/images/front/output/2020-05-11_lab_segmentation/results.json")
hues <- pcv_data$entities$observations$hue_frequencies %>%
  mutate(
    date = pcv_data$entities$metadata$timestamp$value,
    replicate = pcv_data$entities$metadata$frame$value,
    genotype = pcv_data$entities$metadata$id$value
  ) %>%
  select(genotype, replicate, date, "count" = value, "hue" = label, ) %>%
  unnest(cols = c(count, hue))

circular_mean_hue <- pcv_data$entities$observations$hue_circular_mean %>%
  mutate(
    date = pcv_data$entities$metadata$timestamp$value,
    replicate = pcv_data$entities$metadata$frame$value,
    genotype = pcv_data$entities$metadata$id$value
  ) %>%
  select(genotype, replicate, date, "mean_hue" = value)

hues <- hues %>%
  filter(hue != 1) %>%
  group_by(genotype, replicate) %>%
  left_join(circular_mean_hue) %>%
  mutate(
    pixel_count = sum(count),
    count_scaled = count / pixel_count,
    rolling_count = slide_dbl(count_scaled, mean, .before = 3)
  ) 

hues_avg <- hues %>%
  ungroup() %>%
  mutate(genotype = ordered(genotype,
                            levels = c(
                              "Col-0",
                              "Q",
                              "Q-4",
                              "Q-5",
                              "Q-10",
                              "Q-12",
                              "Q-17"
                            ))) %>%
  group_by(genotype, hue) %>%
  summarise(count_mean = mean(count_scaled),
            mean_mean = mean(mean_hue)) %>%
  group_by(genotype) %>%
  mutate(rolling_mean = slide_dbl(count_mean, mean, .before = 3))

H <- round(((1:20) / 20), digits = 2)
S <- 0.7
V <- 0.85

hue_values <- NULL
for (i in H) {
  hue_values[i * 20] <- (hsv(i, S, V))
}


pheno_avg <- pheno_data %>%
  group_by(genotype) %>%
  summarise_if(is.numeric, mean) %>%
  mutate(genotype = recode(genotype,
      "lac4/5/10/12/17" = "Q",
      "lac5/10/12/17" = "Q-4",
      "lac4/10/12/17" = "Q-5",
      "lac4/5/12/17" = "Q-10",
      "lac4/5/10/17" = "Q-12",
      "lac4/5/10/12" = "Q-17"
    ))

hue_dens <- ggplot() +
  geom_density_ridges_gradient(
    data = hues_avg,
    aes(
      x = hue,
      height = rolling_mean,
      y = reorder(genotype, mean_mean),
      # y = genotype,
      fill = hue
    ),
    # scale = 10,
    stat = "identity",
    size = 0.2
  ) +
  # geom_density_ridges(
  #   data = hues,
  #   aes(
  #     x = hue,
  #     height = rolling_count,
  #     y = genotype,
  #     group = interaction(genotype, replicate)
  #   ),
  #   # scale = 10,
  #   fill = NA,
  #   colour = rgb(0, 0, 0, 0.25),
  #   stat = "identity",
  #   size = 0.2
  # ) +
  # geom_text(data = pheno_avg,
  #           aes(label = paste("Senescence index =", round(senescence, digits = 2)),
  #               y = genotype),
  #           x = 100,
  #           family = "Helvetica",
  #           size = ggtext_size,
  #           hjust = 0,
  #           vjust = 1.1) +
  scale_fill_gradientn(colors = hue_values) +
  scale_x_continuous(limits = c(0, 150)) +
  labs(x = "Hue",
       y = "") +
  theme_leo()

pdf("harvest_hue.pdf", width = cm_size(10), height = cm_size(10))
hue_dens
dev.off()
```

#### Height
```{r message=FALSE}
pcv_data <- fromJSON("/home/leonard/Dropbox/2020-01_LAC_phenotyping/images/front/output/2020-05-11_lab_segmentation/results.json")
heights <- as.tibble(pcv_data$entities$observations$height$value) %>%
  mutate(
    longest_path = pcv_data$entities$observations$longest_path$value,
    date = pcv_data$entities$metadata$timestamp$value,
    rep = pcv_data$entities$metadata$frame$value,
    genotype = pcv_data$entities$metadata$id$value
  )

scale_factor <- read_csv("/home/leonard/Dropbox/2020-01_LAC_phenotyping/images/front/output/2020-05-11_lab_segmentation/size_marker.csv",
  col_names = c("id", "space")
) %>%
  mutate(id = str_remove(id, fixed(".jpg"))) %>%
  separate(id, into = c("date", "genotype", "rep"), sep = "_")

heights <- heights %>%
  left_join(scale_factor) %>%
  mutate(
    scaled_height = value / space,
    scaled_path = longest_path / space,
    genotype = recode(genotype,
      Q = "lac4/5/10/12/17",
      `Q-4` = "lac5/10/12/17",
      `Q-5` = "lac4/10/12/17",
      `Q-10` = "lac4/5/12/17",
      `Q-12` = "lac4/5/10/17",
      `Q-17` = "lac4/5/10/12"
    )
  ) %>%
  left_join(pheno_data %>% select(genotype, rep, apparent.height, stretched.height),
    by = c("genotype", "rep")
  )

height_validation <- ggplot(
  data = heights,
  aes(
    x = scaled_height,
    y = apparent.height
  )
) +
  geom_point() +
  geom_smooth(method = "lm")

height_validation

summary(lm(heights$scaled_height ~ heights$apparent.height))

height_validation <- ggplot(
  data = heights,
  aes(
    x = scaled_path,
    y = stretched.height,
    colour = genotype
  )
) +
  geom_point() +
  geom_smooth(
    method = "lm",
    colour = "blue"
  )

height_validation

summary(lm(heights$scaled_path ~ heights$stretched.height))
```

#### Rosette growth

```{r message=FALSE}
#### read-in function ###
read_plus <- function(flnm) {
  json <- fromJSON(flnm)
  json$observations$area$value %>%
    as_tibble() %>%
    mutate(filename = basename(flnm))
}

#### load data ###
rosette_files <-
  list.files(
    path = "/home/leonard/Dropbox/2020-01_LAC_phenotyping/images/top/renamed/nikon_images/",
    pattern = "*.json",
    recursive = FALSE,
    full.names = TRUE
  )

tray_id <- read_csv("/home/leonard/Dropbox/2020-01_LAC_phenotyping/images/top/renamed/tray_id.csv",
  col_types = "cccc"
)

size_marker <- read_csv("/home/leonard/Dropbox/2020-01_LAC_phenotyping/images/top/renamed/nikon_images/output/size_marker_trays.csv",
  col_names = c("filename", "space")
) %>%
  separate(filename, into = c("date", "tray"), sep = "_") %>%
  mutate(
    date = ymd(date),
    tray = str_remove(tray, fixed(".jpg"))
  )

rosette_area <- lapply(rosette_files, read_plus) %>%
  bind_rows() %>%
  separate(filename, into = c("date", "tray", "plant"), sep = "_") %>%
  mutate(
    date = ymd(date),
    tray = str_remove(tray, fixed(".jpg")),
    plant = str_remove(plant, fixed(".json"))) %>%
  left_join(tray_id) %>%
  left_join(size_marker) %>%
  mutate(area = value / space,
         mutation = case_when(genotype == "Col-0" ~ "WT",
                              TRUE ~ "mutant")) %>%
  group_by(genotype, date) %>%
  mutate(area_avg = mean(area)) %>%
  group_by(genotype) %>%
  mutate(max = max(area_avg)) %>%
  filter(date > "2020-01-27")

rosette_growth <- ggplot(
  data = rosette_area,
  aes(
    x = date,
    y = area
  )
) +
  geom_line(aes(
    colour = mutation,
    group = interaction(genotype, replicate)
  ),
  size = 0.2,
  alpha = 0.25
  ) +
  geom_line(aes(
    colour = mutation,
    group = genotype,
    y = area_avg
  )) +
  theme_leo() +
  coord_cartesian(clip = "off") +
  scale_x_date(limits = c(ymd("20200127"), ymd("20200214"))) +
  geom_text_repel(
    data = distinct(rosette_area, max),
    x = ymd("20200212"),
    aes(
      y = max,
      label = genotype
    ),
    family = "Helvetica",
    segment.colour = "grey75",
    hjust = 0,
    direction = "y",
    nudge_x = 0.5,
    min.segment.length = 0
  ) +
  labs(x = "Date",
       y = "Rosette Area [scaled px]") +
  scale_colour_manual(values = pal_ostwald_disc[c(3,1)])

pdf("rosette_growth.pdf", width = cm_size(10), height = cm_size(10))
rosette_growth
dev.off()
```

#### Root growth

```{r}
root_data <- read_csv("/data/PhD/Phenotyping/2020-10_lac_root_growth/2020-10_root_growth.csv")

ggplot(root_data, aes(y = length.cm,
                      x = genotype)) +
  geom_violin(draw_quantiles = 0.5) +
  geom_quasirandom() +
  theme_leo()

```
