---
title: "Raman Multivariate Analysis"
author: "Leonard Blaschek"
date: "16 May 2019"
output: 
    html_document
#   pdf_document:
#     latex_engine: lualatex
#     fig_caption: yes
#     fig_height: 6
#     includes:
#       in_header: rmd_temp.tex
# sansfont: Cooper Hewitt
# monofont: IBM Plex Mono
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ropls)
library(factoextra)
library(cowplot)
library(showtext)
library(ggthemes)
library(plotly)
library(zoo)
library(kableExtra)
library(tidyverse)
library(ggrepel)
library(gplots)
library(viridis)

#### import CooperHewitt Neue ####
font_add(
  "CooperHewitt",
  regular = "/OFL_fonts/CooperHewitt-OTF-public/CooperHewitt-Book.otf",
  italic = "/OFL_fonts/CooperHewitt-OTF-public/CooperHewitt-BookItalic.otf",
  bold = "/OFL_fonts/CooperHewitt-OTF-public/CooperHewitt-Semibold.otf",
  bolditalic = "/OFL_fonts/CooperHewitt-OTF-public/CooperHewitt-SemiboldItalic.otf"
)
showtext_auto()

#### generating plot theme ####
theme_leo <- function(base_size = 14,
                      base_family = "CooperHewitt"){
  theme_minimal(base_size = base_size,
                base_family = base_family) %+replace%
    theme(
      strip.text = element_text(hjust = 0, face = "italic"),
      axis.ticks = element_line(
        size = 0.25,
        lineend = "square",
        color = "black"
      ),
      axis.text.x = element_text(colour = "black", # flipped coords
                                 margin = margin(1, 1, 1, 1)),
      axis.text.y = element_text(
        colour = "black",
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1)
      ),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", size = 0.25),
      panel.spacing = unit(1.5, "mm"),
      legend.position = "none",
      legend.text = element_text(size = rel(0.8)),
      legend.key.height = unit(4, "mm"),
      complete = TRUE
    )
}
```

## Raman peaks from the literature

```{r literature, warning=FALSE, message=FALSE}
rmn_tab <- read_csv("/home/leonard/Documents/Uni/PhD/Raman/raman_peak_assignments.csv")
kable(rmn_tab, caption = "Table 1: Peaks previously identified as diagnostic for lignin and cellulose.") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Read and tidy the Raman data

```{r load data, warning=FALSE, message=FALSE}
rmn_data <- read_csv("Raman_corrected_and_filtered.csv")
rmn_data$replicate <- as.factor(rmn_data$replicate)
rmn_data$technical <- as.factor(rmn_data$technical)
head(rmn_data)
```

## Plot wild type spectra
```{r plot spectra, out.width = '100%', warning=FALSE, message=FALSE}
rmn_data_pre <- rmn_data %>%
  group_by(genotype, cell.type, replicate, wavenumber) %>%
  summarise(
    mean.intensity.pre = mean(corrected.intensity, na.rm = TRUE),
    sd.intensity.pre = sd(corrected.intensity, na.rm = TRUE),
    samples.pre = length(corrected.intensity),
    mean.scaled = mean(scaled, na.rm = TRUE)
  ) 

rmn_data_avg <- rmn_data_pre %>%
  group_by(genotype, cell.type, wavenumber) %>%
  summarise(
    mean.intensity = mean(mean.intensity.pre, na.rm = TRUE),
    sd.intensity = sd(mean.intensity.pre, na.rm = TRUE),
    plants = paste("plants: ", length(mean.intensity.pre)),
    bundles = ifelse(length(mean.intensity.pre) > 1, 
                     paste("cells/plant: ", min(samples.pre), "—", max(samples.pre), sep = ""),
                     paste("cells/plant: ", min(samples.pre))),
    mean.scaled = mean(mean.scaled, na.rm = TRUE)
  )

rmn_data_avg_WT <- filter(rmn_data_avg, genotype == "Col-0" & cell.type %in% c("PMX", "IF"))
rmn_data_pre_WT <- filter(rmn_data_pre, genotype == "Col-0" & cell.type %in% c("PMX", "IF"))
rmn_data_WT <- filter(rmn_data, genotype == "Col-0" & cell.type %in% c("PMX", "IF"))
rmn_data_WT <- rmn_data_WT %>%
  unite("grouptech", replicate, technical, remove = FALSE)

#### plotting unscaled spectra ####
spectra.WT.MX <- ggplot() +
  geom_line(data = rmn_data_WT,
            aes(
              x = wavenumber, 
              y = rollmean(corrected.intensity, 3, na.pad=TRUE),
              colour = replicate,
              group = grouptech
              ),
            size = 0.2,
            alpha = 0.25
            ) +
  geom_line(data = rmn_data_avg_WT,
            aes(x = wavenumber, y = rollmean(mean.intensity, 3, na.pad=TRUE)),
            size = 0.1) +
  scale_x_continuous(limits = c(300, 2000)) +
  scale_y_continuous(limits = c(-1000, 14500)) +
  scale_fill_viridis_d() +
  scale_colour_few() +
  labs(x = "Wavenumber",
       y = "Corrected intensity") +
  theme_leo() +
  geom_text(
    data = subset(rmn_data_avg_WT, wavenumber == 1430),
    x = 1950,
    y = 14500,
    aes(label = plants),
    stat = "identity",
    family = "CooperHewitt",
    size = 2.5,
    hjust = 1
  ) +
  geom_text(
    data = subset(rmn_data_avg_WT, wavenumber == 1430),
    x = 1950,
    y = 13500,
    aes(label = bundles),
    stat = "identity",
    family = "CooperHewitt",
    size = 2.5,
    hjust = 1
  ) +
  facet_grid(cell.type ~ genotype) 
ggplotly(spectra.WT.MX)

#### plotting scaled spectra ####
spectra.WT.scaled <- ggplot() +
  geom_line(data = rmn_data_avg_WT,
            aes(
              x = wavenumber, 
              y = rollmean(mean.scaled, 3, na.pad=TRUE),
              colour = cell.type
            ),
            size = 1,
            alpha = 1
  ) +
  scale_x_continuous(limits = c(300, 2000)) +
  scale_y_continuous(limits = c(-0.1, 1)) +
  scale_fill_viridis_d() +
  scale_colour_few() +
  theme_leo() +
  labs(x = "Wavenumber",
       y = "Normalised intensity")
ggplotly(spectra.WT.scaled)
```

# Clustering

```{r cluster, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_data_spread <- rmn_data %>%
  filter(cell.type %in% c("PMX", "IF")) %>%
  select(-intensity, -corrected.intensity) %>%
  spread(key = wavenumber, value = scaled)

rmn_data_mat <- rmn_data_spread %>%
  select(-replicate, -technical) %>%
  group_by(genotype, cell.type) %>%
  summarise_all(list(mean)) %>%
  ungroup()
rmn_mat <- as.matrix(select(rmn_data_mat, -genotype, -cell.type))
rwnms <- unite(rmn_data_mat,"ID", genotype, cell.type, sep = "_")
rownames(rmn_mat) <- rwnms$ID
heatmap.2(rmn_mat,
          Colv = FALSE,
          col = viridis,
          trace = "none",
          symbreaks = FALSE)
```

Basic clustering of the range normalised spectra groups the lignin deficient IFs together based on their high relative cellulose content (*ccr1*, *4cl1*, *4cl1/2*), but other patterns are harder todistinguish. Let's move on to methods for reducing dimensionality.

# Principle component analysis (PCA)

```{r PCA, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_pca <- prcomp(select(rmn_data_spread, -genotype, -cell.type, -replicate, -technical), center = TRUE, scale. = TRUE)
rmn_pca_gg <- data.frame(rmn_pca$x, genotype = rmn_data_spread$genotype, cell.type = rmn_data_spread$cell.type)

shapes = c("IF" = 21,
           "PMX" = 22)

ggplotly(ggplot(rmn_pca_gg, aes(x = PC1, y = PC2)) +
  geom_hline(yintercept = 0, linetype = 1) +
  geom_vline(xintercept = 0, linetype = 1) +
  stat_ellipse(aes(fill = genotype, 
                   group = interaction(genotype, cell.type)),
               geom = "polygon", alpha = 0.5, type = "t", level = 0.4) +
  stat_ellipse(aes(fill = NULL), colour = "black", linetype = 2) +
  geom_point(aes(shape = cell.type,
                 fill = genotype),
             size = 4,
             stroke = 0.5,
             alpha = 1) +
  scale_shape_manual(values = shapes) +
  scale_colour_viridis_d() +
  theme_leo())

top <- plot_grid(
  fviz_eig(rmn_pca, 
           main = "", 
           font.family = "CooperHewitt",
           ggtheme = theme_leo(),
           barcolor = NA,
           barfill = "#1d91c0",
           font.xtickslab = c(20, "plain", "black"), 
           font.ytickslab = c(20, "plain", "black"),
           font.x = c(20, "plain", "black"),
           font.y = c(20, "plain", "black")
  ),
  labels = c("(a)"),
  label_fontfamily = "CooperHewitt",
  label_size = 20,
  rel_widths = c(2,1),
  nrow = 1,
  align = "h"
)

mid <- plot_grid(
  fviz_contrib(rmn_pca, 
               choice = "var", 
               axes = 1, 
               top = 12, 
               title = "", 
               font.family = "CooperHewitt",
               ggtheme = theme_leo(),
               color = NA,
               fill = "#1d91c0",
           font.xtickslab = c(20, "plain", "black"), 
           font.ytickslab = c(20, "plain", "black"),
           font.y = c(20, "plain", "black")
  ),
  fviz_contrib(rmn_pca, 
               choice = "var", 
               axes = 2, 
               top = 12, 
               title = "", 
               font.family = "CooperHewitt",
               ggtheme = theme_leo(),
               color = NA,
               fill = "#1d91c0",
           font.xtickslab = c(20, "plain", "black"), 
           font.ytickslab = c(20, "plain", "black"),
           font.y = c(20, "plain", "black")
  ),
  labels = c("(b)", "(c)"),
  label_fontfamily = "CooperHewitt",
  label_size = 20,
  nrow = 2
)

plot_grid(
top,
mid,
nrow = 1
)
```
**(b)** represents PC1, **(c)** represents PC2.

# Orthogonal partial least square discriminant analysis (OPLS-DA)

## Raman signals diagnostic for S, G and 5-OH-G units

First, let's look at rough PCA and OPLS-DA differentiating IF (S-unit rich), from PMX (G-unit rich).

```{r opls-da, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_mat <- data.matrix(select(rmn_data_spread, -genotype, -cell.type, -replicate, -technical))
rwnms <- unite(rmn_data_spread,"ID", genotype, cell.type, replicate, technical, sep = "_")
rownames(rmn_mat) <- rwnms$ID
rmn_meta <- as.matrix(select(rmn_data_spread, genotype, cell.type, replicate, technical))

rmn_pca <- opls(rmn_mat)
cell.type <- rmn_meta[, "cell.type"]
plot(rmn_pca,
     typeVc = "x-score",
     parAsColFcVn = cell.type)

rmn_oplsda <- opls(rmn_mat, cell.type, predI = 1, orthoI = NA)
```

Next, let's have a look at the normalised spectra of the IF in the WT (S-units), *omt1* (5-OH-G-units), and *fah1* (only G-units).

```{r S/G/5-OH-G spectra, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_data_avg_SG <- filter(rmn_data_avg, genotype %in% c("Col-0", "fah1", "omt1") & cell.type == "IF")
rmn_data_pre_SG <- filter(rmn_data_pre, genotype %in% c("Col-0", "fah1", "omt1") & cell.type == "IF")
rmn_data_SG <- filter(rmn_data, genotype %in% c("Col-0", "fah1", "omt1") & cell.type == "IF")
rmn_data_SG <- rmn_data_SG %>%
  unite("grouptech", replicate, technical, remove = FALSE)

#### plotting scaled spectra ####
spectra.SG.scaled <- ggplot() +
  geom_line(data = rmn_data_avg_SG,
            aes(
              x = wavenumber, 
              y = rollmean(mean.scaled, 3, na.pad=TRUE),
              colour = genotype
            ),
            size = 1,
            alpha = 1
  ) +
  scale_x_continuous(limits = c(300, 2000)) +
  scale_y_continuous(limits = c(-0.1, 1)) +
  scale_fill_viridis_d() +
  scale_colour_few() +
  theme_leo() +
  labs(x = "Wavenumber",
       y = "Normalised intensity")
ggplotly(spectra.SG.scaled)

```

### OPLS-DA of PMX vs. IF in the wild type
To see how the IF/PMX differentiation is affected by *fah1* and *omt1* mutants, here are OPLS-DA plots for only the WT spectra.

```{r opls-da WT, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_mat <- data.matrix(rmn_data_spread %>%
                         filter(genotype == "Col-0") %>%
                         select(-genotype, -cell.type, -replicate, -technical))
rwnms <- rmn_data_spread %>%
  filter(genotype == "Col-0") %>%
  select(genotype, cell.type, replicate, technical) %>%
  unite("ID", genotype, cell.type, replicate, technical, sep = "_")
rownames(rmn_mat) <- rwnms$ID
rmn_meta <- as.matrix(rmn_data_spread %>%
                         filter(genotype == "Col-0") %>%
                         select(genotype, cell.type, replicate, technical))

cell.type <- rmn_meta[, "cell.type"]
rmn_oplsda <- opls(rmn_mat, cell.type, predI = 1, orthoI = NA)
```

Much clearer! So now let's have a look at the loadings.

```{r opls-da WT loadings, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_oplsda_load <- as_tibble(getLoadingMN(rmn_oplsda), rownames = "wavenumber") %>%
  rename(loading = p1) %>%
  mutate(wavenumber = as.numeric(wavenumber)) %>%
  left_join(select(rmn_tab, wavenumber, Polymer, Substructure) %>% 
              unite(lbl, Polymer, Substructure, sep = ", ") %>%
              mutate(wavenumber = as.numeric(wavenumber),
                     lbl = str_replace_all(lbl, "\\~", ""))) %>%
  ggplot(aes(x = wavenumber, y = rollmean(loading, 5, na.pad=TRUE))) +
  geom_line() +
  geom_hline(yintercept = 0, linetype =2) +
  labs(x = "Wavenumber",
       y = "Loading") +
  theme_leo() +
  geom_label_repel(aes(label = lbl), 
                   nudge_y = 0.02,
                   family = "CooperHewitt",
                   segment.colour = "red")
rmn_oplsda_load
```

Cellulose peaks are quite important in the OPLS-DA model for differentiating between IF and PMX. The most important lignin peaks are the ones assigned to G-OH and G-CHO unit C=C and C=O stretches at 1662, and the aryl-OH/OCH~3~ at 1336. Those two are arguably the ones most related to the S/G ratio, so that's no surprise.

### OPLS-DA of PMX vs. IF in *fah1* and *omt1*
In the converse, that should mean that S-units are crucial for distinction of PMX/IF spectra, and *fah1* and *omt1* should overlap closer.

```{r opls-da noS, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_mat <- data.matrix(rmn_data_spread %>%
                         filter(genotype %in% c("fah1", "omt1")) %>%
                         select(-genotype, -cell.type, -replicate, -technical))
rwnms <- rmn_data_spread %>%
  filter(genotype %in% c("fah1", "omt1")) %>%
  select(genotype, cell.type, replicate, technical) %>%
  unite("ID", genotype, cell.type, replicate, technical, sep = "_")
rownames(rmn_mat) <- rwnms$ID
rmn_meta <- as.matrix(rmn_data_spread %>%
                         filter(genotype %in% c("fah1", "omt1")) %>%
                         select(genotype, cell.type, replicate, technical))

cell.type <- rmn_meta[, "cell.type"]
rmn_oplsda <- opls(rmn_mat, cell.type, predI = 1, orthoI = NA)
```

A little less clear of a divide between PMX and IF, but still there, meaning that we have more that just S/G to differentiate between those to lignins. So let's look at the loadings again.

```{r opls-da noS loadings, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_oplsda_load <- as_tibble(getLoadingMN(rmn_oplsda), rownames = "wavenumber") %>%
  rename(loading = p1) %>%
  mutate(wavenumber = as.numeric(wavenumber)) %>%
  left_join(select(rmn_tab, wavenumber, Polymer, Substructure) %>% 
              unite(lbl, Polymer, Substructure, sep = ", ") %>%
              mutate(wavenumber = as.numeric(wavenumber),
                     lbl = str_replace_all(lbl, "\\~", ""))) %>%
  ggplot(aes(x = wavenumber, y = rollmean(loading, 5, na.pad=TRUE))) +
  geom_line() +
  geom_hline(yintercept = 0, linetype =2) +
  labs(x = "Wavenumber",
       y = "Loading") +
  theme_leo() +
  geom_label_repel(aes(label = lbl), 
                   nudge_y = 0.02,
                   family = "CooperHewitt",
                   segment.colour = "red")
rmn_oplsda_load
```

The loadings look rather similar, with cellulose as the strongest predictor followed by the peaks at 1662 and 1336. Which means that these might be diagnostic for differences between G and 5-OH-G units. Let's see.

### OPLS-DA of *fah1* vs. *omt1* in the IF

```{r opls-da G to 5-OH-G, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_mat <- data.matrix(rmn_data_spread %>%
                         filter(genotype %in% c("fah1", "omt1") & cell.type == "IF") %>%
                         select(-genotype, -cell.type, -replicate, -technical))
rwnms <- rmn_data_spread %>%
  filter(genotype %in% c("fah1", "omt1") & cell.type == "IF") %>%
  select(genotype, cell.type, replicate, technical) %>%
  unite("ID", genotype, cell.type, replicate, technical, sep = "_")
rownames(rmn_mat) <- rwnms$ID
rmn_meta <- as.matrix(rmn_data_spread %>%
                         filter(genotype %in% c("fah1", "omt1") & cell.type == "IF") %>%
                         select(genotype, cell.type, replicate, technical))

genotype <- rmn_meta[, "genotype"]
rmn_oplsda <- opls(rmn_mat, genotype, predI = 1, orthoI = NA)
```

Still a good, predictive (Q2Y = 0.65), separation. Let's look at the loadings again.

```{r opls-da G to 5-OH-G loadings, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_oplsda_load <- as_tibble(getLoadingMN(rmn_oplsda), rownames = "wavenumber") %>%
  rename(loading = p1) %>%
  mutate(wavenumber = as.numeric(wavenumber)) %>%
  left_join(select(rmn_tab, wavenumber, Polymer, Substructure) %>% 
              unite(lbl, Polymer, Substructure, sep = ", ") %>%
              mutate(wavenumber = as.numeric(wavenumber),
                     lbl = str_replace_all(lbl, "\\~", ""))) %>%
  ggplot(aes(x = wavenumber, y = rollmean(loading, 5, na.pad=TRUE))) +
  geom_line() +
  geom_hline(yintercept = 0, linetype =2) +
  labs(x = "Wavenumber",
       y = "Loading") +
  theme_leo() +
  geom_label_repel(aes(label = lbl), 
                   nudge_y = 0.02,
                   family = "CooperHewitt",
                   segment.colour = "red")
rmn_oplsda_load
```

So the peak at 1662 is in fact specific to G units and not S or 5-OH-G units. So next, let's look how we can differentiate S from 5-OH-G units.

### OPLS-DA of Col-0 vs. *omt1* in the IF

```{r opls-da S to 5-OH-G, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_mat <- data.matrix(rmn_data_spread %>%
                         filter(genotype %in% c("Col-0", "omt1") & cell.type == "IF") %>%
                         select(-genotype, -cell.type, -replicate, -technical))
rwnms <- rmn_data_spread %>%
  filter(genotype %in% c("Col-0", "omt1") & cell.type == "IF") %>%
  select(genotype, cell.type, replicate, technical) %>%
  unite("ID", genotype, cell.type, replicate, technical, sep = "_")
rownames(rmn_mat) <- rwnms$ID
rmn_meta <- as.matrix(rmn_data_spread %>%
                         filter(genotype %in% c("Col-0", "omt1") & cell.type == "IF") %>%
                         select(genotype, cell.type, replicate, technical))

genotype <- rmn_meta[, "genotype"]
rmn_oplsda <- opls(rmn_mat, genotype, predI = 1, orthoI = NA)
```

It works, let's see how.

```{r opls-da S to 5-OH-G loadings, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_oplsda_load <- as_tibble(getLoadingMN(rmn_oplsda), rownames = "wavenumber") %>%
  rename(loading = p1) %>%
  mutate(wavenumber = as.numeric(wavenumber)) %>%
  left_join(select(rmn_tab, wavenumber, Polymer, Substructure) %>% 
              unite(lbl, Polymer, Substructure, sep = ", ") %>%
              mutate(wavenumber = as.numeric(wavenumber),
                     lbl = str_replace_all(lbl, "\\~", ""))) %>%
  ggplot(aes(x = wavenumber, y = rollmean(loading, 5, na.pad=TRUE))) +
  geom_line() +
  geom_hline(yintercept = 0, linetype =2) +
  labs(x = "Wavenumber",
       y = "Loading") +
  theme_leo() +
  geom_label_repel(aes(label = lbl), 
                   nudge_y = 0.02,
                   family = "CooperHewitt",
                   segment.colour = "red")
rmn_oplsda_load
```

So it looks like the peak at 1336 is specific to aryl-O-CH~3~, and does not actually respond to aryl-OH bonds. This could also be due to different concentrations of S-units in WT IF vs. 5-OH-G units in *omt1* IF, however.

### Three-way PLS-DA of Col-0 vs. *fah1* vs. *omt1* in the IF

In his FT-IR OPLS-DA paper, Andras' writes that he used OPLS-DA with three classes (fibres, vessels, rays). As far as I can make out however, having more than 2 classes somewhat defeats the purpose of OPLS-DA, which, to my understanding, is to compress the differentiating variation into one axis, while all variation that is not relevant to the discrimination between the two groups is expressed in the orthogonal axis. More than 2 classes would therefore remove the O(rthogonal) aspect, making it a PLS-DA. Accordingly, I did not find a way to perform a OPLS-DA with three classes in R. I could however perform a PLS-DA. 

```{r opls-da S/G/5-OH-G, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_mat <- data.matrix(rmn_data_spread %>%
                         filter(genotype %in% c("Col-0", "fah1", "omt1") & cell.type == "IF") %>%
                         select(-genotype, -cell.type, -replicate, -technical))
rwnms <- rmn_data_spread %>%
  filter(genotype %in% c("Col-0", "fah1", "omt1") & cell.type == "IF") %>%
  select(genotype, cell.type, replicate, technical) %>%
  unite("ID", genotype, cell.type, replicate, technical, sep = "_")
rownames(rmn_mat) <- rwnms$ID
rmn_meta <- as.matrix(rmn_data_spread %>%
                         filter(genotype %in% c("Col-0", "fah1", "omt1") & cell.type == "IF") %>%
                         select(genotype, cell.type, replicate, technical))

genotype <- rmn_meta[, "genotype"]
rmn_plsda <- opls(rmn_mat, genotype)
```


```{r opls-da S/G/5-OH-G loadings, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_plsda_load <- as_tibble(getLoadingMN(rmn_oplsda), rownames = "wavenumber") %>%
  rename(loading = p1) %>%
  mutate(wavenumber = as.numeric(wavenumber)) %>%
  left_join(select(rmn_tab, wavenumber, Polymer, Substructure) %>% 
              unite(lbl, Polymer, Substructure, sep = ", ") %>%
              mutate(wavenumber = as.numeric(wavenumber),
                     lbl = str_replace_all(lbl, "\\~", ""))) %>%
  ggplot(aes(x = wavenumber, y = rollmean(loading, 5, na.pad=TRUE))) +
  geom_line() +
  geom_hline(yintercept = 0, linetype =2) +
  labs(x = "Wavenumber",
       y = "Loading") +
  theme_leo() +
  geom_label_repel(aes(label = lbl), 
                   nudge_y = 0.02,
                   family = "CooperHewitt",
                   segment.colour = "red")
rmn_plsda_load
```

## Raman signals diagnostic for aldehyde rich lignin

### OPLS-DA of Col-0 vs. *cad4*x*cad5* in the IF

```{r CHO spectra, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_data_avg_CHO <- filter(rmn_data_avg, genotype %in% c("Col-0", "cad4xcad5") & cell.type == "IF")
rmn_data_pre_CHO <- filter(rmn_data_pre, genotype %in% c("Col-0", "cad4xcad5") & cell.type == "IF")
rmn_data_CHO <- filter(rmn_data, genotype %in% c("Col-0", "cad4xcad5") & cell.type == "IF")
rmn_data_CHO <- rmn_data_CHO %>%
  unite("grouptech", replicate, technical, remove = FALSE)

#### plotting scaled spectra ####
spectra.CHO.scaled <- ggplot() +
  geom_line(data = rmn_data_avg_CHO,
            aes(
              x = wavenumber, 
              y = rollmean(mean.scaled, 3, na.pad=TRUE),
              colour = genotype
            ),
            size = 1,
            alpha = 1
  ) +
  scale_x_continuous(limits = c(300, 2000)) +
  scale_y_continuous(limits = c(-0.1, 1)) +
  scale_fill_viridis_d() +
  scale_colour_few() +
  theme_leo() +
  labs(x = "Wavenumber",
       y = "Normalised intensity")
ggplotly(spectra.CHO.scaled)

```

```{r opls-da CHO, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_mat <- data.matrix(rmn_data_spread %>%
                         filter(genotype %in% c("Col-0", "cad4xcad5") & cell.type == "IF") %>%
                         select(-genotype, -cell.type, -replicate, -technical))
rwnms <- rmn_data_spread %>%
  filter(genotype %in% c("Col-0", "cad4xcad5") & cell.type == "IF") %>%
  select(genotype, cell.type, replicate, technical) %>%
  unite("ID", genotype, cell.type, replicate, technical, sep = "_")
rownames(rmn_mat) <- rwnms$ID
rmn_meta <- as.matrix(rmn_data_spread %>%
                         filter(genotype %in% c("Col-0", "cad4xcad5") & cell.type == "IF") %>%
                         select(genotype, cell.type, replicate, technical))

genotype <- rmn_meta[, "genotype"]
rmn_oplsda <- opls(rmn_mat, genotype, predI = 1, orthoI = NA)
```


```{r opls-da CHO loadings, fig.showtext=TRUE, out.width = '100%', warning=FALSE, message=FALSE}
rmn_oplsda_load <- as_tibble(getLoadingMN(rmn_oplsda), rownames = "wavenumber") %>%
  rename(loading = p1) %>%
  mutate(wavenumber = as.numeric(wavenumber)) %>%
  left_join(select(rmn_tab, wavenumber, Polymer, Substructure) %>% 
              unite(lbl, Polymer, Substructure, sep = ", ") %>%
              mutate(wavenumber = as.numeric(wavenumber),
                     lbl = str_replace_all(lbl, "\\~", ""))) %>%
  ggplot(aes(x = wavenumber, y = rollmean(loading, 5, na.pad=TRUE))) +
  geom_line() +
  geom_hline(yintercept = 0, linetype =2) +
  labs(x = "Wavenumber",
       y = "Loading") +
  theme_leo() +
  geom_label_repel(aes(label = lbl), 
                   nudge_y = 0.02,
                   family = "CooperHewitt",
                   segment.colour = "red")
rmn_oplsda_load
```