---
title: "SEM course"
always_allow_html: yes
output: 
  # pdf_document:
  #   fig_height: 6
  #   includes:
  #     in_header: rmd_temp.tex
  html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Building a Model 

### Model Saturation

Models can be fitted if
$$t \le \frac{n(n+1)}{2}$$
where

$n$ = knowns (variables/variation)

$t$ = unknowns (correlations/covariation)

This condition is sually only unsatisfied when fitting bidirectional realtionships.

## Global Estimation with Lavaan

```{r testrun, message = FALSE}
library(lavaan)
library(piecewiseSEM)
data(keeley)
head(keeley, n = 1)

#fit linear model
keeley_lm <- lm(cover ~ age, keeley)
summary(keeley_lm)
coefs(keeley_lm)

#fit linear model with lavaan
keeley_formula1 <- "cover ~ age"
keeley_sem1 <- sem(keeley_formula1, data = keeley) #"meanstructure = T" for intercept
summary(keeley_sem1, standardized = TRUE, rsq = TRUE)

#fit partial mediation SEM
keeley_formula2 <- "
firesev ~ age
cover ~ age + firesev
"
keeley_sem2 <- sem(keeley_formula2, data = keeley)
# standardized variances are error correlations, 1-them is explained variance
summary(keeley_sem2, standardized = TRUE, rsq = TRUE)

#indirect effects
keeley_formula2.1 <- "
firesev ~ af * age
cover ~ ac * age + fc * firesev

indirect := af * fc 
total := ac + af * fc
"
summary(sem(keeley_formula2.1, keeley), standardize = TRUE, rsq = TRUE)

#fit distance -> rich (directly + via abiotic and hetero)
#chi-square should be high to show that model is not sig. diff. from data
keeley_formula3 <- "
abiotic ~ da * distance
hetero ~ dh * distance
rich ~ dr * distance + hr * hetero + ar * abiotic 

direct := dr
indirect := ar * da + hr * dh
"
summary(sem(keeley_formula3, keeley, estimator = "MLM"), standardize = TRUE, rsq = TRUE)

#fix mediated path coefficients to zero
keeley_formula4 <- "
firesev ~ 1
cover ~ age
"
summary(sem(keeley_formula4, keeley), standardize = TRUE, rsq = TRUE)

#correlated errors
keeley_formula5 <- "
cover ~ age
firesev ~ age
cover ~~ firesev"
summary(sem(keeley_formula5, keeley), standardize = TRUE, rsq = TRUE)
```

## Local Estimation with piecewiseSEM

```{r piecewise test, message = FALSE}
keeley.sem <- psem(
  lm(abiotic ~ distance, data = keeley),
  lm(hetero ~ distance, data = keeley),
  lm(rich ~ abiotic + hetero, data = keeley),
  data = keeley
)
basisSet(keeley.sem)

#Conduct d-sep tests
keeley.dsep <- dSep(keeley.sem, conditioning = TRUE)
keeley.dsep

#Fisher's C test
fisherC(keeley.sem)

#Update the model
keeley.sem2 <- psem(
  lm(abiotic ~ distance, data = keeley),
  lm(hetero ~ distance, data = keeley),
  lm(rich ~ distance + abiotic + hetero, data = keeley),
  data = keeley
)
basisSet(keeley.sem2)

#Conduct d-sep tests on updated model
keeley.dsep <- dSep(keeley.sem2, conditioning = TRUE)
keeley.dsep

#Fisher's C test of updated model
fisherC(keeley.sem2)

#get coefficients
coefs(keeley.sem2)

#get R-squared
rsquared(keeley.sem2)

#get summary
summary(keeley.sem2)

#compare to lavaan
keeley.model <- "
abiotic ~ distance
hetero ~ distance
rich ~ distance + hetero + abiotic
"
summary(sem(keeley.model, data = keeley), standardize = TRUE, rsq = TRUE)

#correlated errors, in case the causality between rich and abiotic is unclear
keeley.sem3 <- psem(
  lm(abiotic ~ distance, data = keeley),
  lm(hetero ~ distance, data = keeley),
  lm(rich ~ distance + abiotic + hetero, data = keeley),
  rich %~~% abiotic,
  data = keeley
)
summary(keeley.sem3)

#keep a variable (abiotic) in the system, without paths
keeley.sem4 <- psem(
  lm(hetero ~ distance, data = keeley),
  lm(rich ~ distance + hetero, data = keeley),
  abiotic ~ 1,
  data = keeley
)
summary(keeley.sem4)

#poisson distributed response variable (rich)
keeley.glm.sem <- psem(
  lm(abiotic ~ distance, data = keeley),
  lm(hetero ~ distance, data = keeley),
  glm(rich ~ distance + abiotic + hetero, family = "poisson", data = keeley),
  data = keeley
)
summary(keeley.glm.sem)

#shipley pSEM
library(nlme)
library(lme4)

data(shipley)

shipley.sem <- psem(
  lme(DD ~ lat, random = ~1|site/tree, na.action = na.omit, data = shipley),
  lme(Date ~ DD, random = ~1|site/tree, na.action = na.omit, data = shipley),
  lme(Growth ~ Date, random = ~1|site/tree, na.action = na.omit, data = shipley),
  glmer(Live ~ Growth + (1|site) + (1|tree), family = binomial(link = "logit"),
        data = shipley)
)
summary(shipley.sem, .progressBar = FALSE)

#durocher SEM
durocher <- read.csv("file:///home/leonard/Dropbox/2019 Stockholm Course/Lectures/9_Mixed_Models/durocher.csv")
head(durocher, n = 1)

durocher.sem <- psem(
  lme(CR ~ Std.Temp, random = ~1|Pond.ID, na.action = na.omit, data = durocher),
  lme(Prich ~ Std.Temp, random = ~1|Pond.ID, na.action = na.omit, data = durocher),
  lme(Pbio ~ Prich, random = ~1|Pond.ID, na.action = na.omit, data = durocher),
  lme(GPP ~ Pbio, random = ~1|Pond.ID, na.action = na.omit, data = durocher),
  data = durocher
)
summary(durocher.sem)
```


## piecewiseSEM on IRX data

![First model, including all measured parameters that are implicated by logic or the literature](/home/leonard/Dropbox/SEM course/causal_model_1.pdf)

### Loading the Data

I start with data averaged by genotype, before hopefully coming back to it with data paired by vessel.

```{r load data, message = FALSE, include=TRUE}
library(dplyr)
library(tidyr)
library(ggplot2)
sem.data <- read.csv("file:///home/leonard/R/Output/PhD/SEM_data.csv") %>%
  select(-X, -X1599...cm....1.1119..cm....1) %>%
  drop_na()

# sem.gather <- sem.data %>%
#   gather(key = "variable",
#          value = "value",
#          - genotype,
#          - replicate,
#          - cell.type,
#          - technical)
# head(sem.gather)
ggplot(sem.data, aes(x = genotype, y = Integrated.lignin.peak)) +
  geom_violin() +
  geom_jitter() +
  facet_wrap(~ cell.type, ncol = 3)
```

### Overall Correlations

```{r plot correlations, message = FALSE}
library(PerformanceAnalytics)
chart.Correlation(select(ungroup(sem.data), 
                         -genotype, -cell.type, -replicate, -technical), 
                  histogram=TRUE, pch=21)
```

![Second model, now simplified and focused on the variables that are central to the hypothesis or implicated by direct correlation](/home/leonard/Dropbox/SEM course/causal_model_2.pdf)

### Fit First pSEM Model

```{r first psem}
first.model <- psem(
  lm(Circ. ~ n_v + 
       Integrated.cellulose.peak +
       Integrated.lignin.peak,
     # family = "binomial",
     data = sem.data),
  lm(Plant.height ~ Circ., data = sem.data),
  data = sem.data
)
summary(first.model, .progressBar = FALSE)
```

![Third model, now including the composite variable "Cellular environment"](/home/leonard/Dropbox/SEM course/causal_model_3.pdf)


```{r second psem}
#introduce neighbourhood composite
betas <- summary(lm(Circ. ~ n_f + n_v,
     data = sem.data))$coefficients[-1,1]

sem.data$composite <- rowSums(sem.data[, c("n_f", "n_v")] * betas)

second.model <- psem(
  lme(Circ. ~ 
        composite +
        # n_v +
        # n_f +
        Integrated.cellulose.peak +
        Integrated.lignin.peak +
        Lignin.peak.position,
      random = ~1|cell.type/replicate,
     # family = "binomial",
     data = sem.data),
  lme(Plant.height ~ Circ. +
        Integrated.lignin.peak +
        Lignin.peak.position,
     random = ~1|replicate,
     data = sem.data
     ),
  data = sem.data
)

summary(second.model, .progressBar = FALSE)

plot(second.model)
```

### Pointers for follow-ups

* multigroup analysis for cell type comparison
* categorical cell type for qualitative comparisons
* paths that can be constrained will be estimated from all data in the final diagram
* std. estimates are not comparable across multigroup models because they're dependent on the variance within the respective group

```{r multigroup}
# second.model.multigroup <- multigroup(second.model,
#                                       group = "cell.type"
#                                       )
# second.model.multigroup
```